\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{longtable}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,decorations.pathreplacing}
\usepackage{mathtools}

% ---------- Cyrillic Sha Symbol ----------
\DeclareFontFamily{U}{wncy}{}
\DeclareFontShape{U}{wncy}{m}{n}{<->wncyr10}{}
\DeclareSymbolFont{mcy}{U}{wncy}{m}{n}
\DeclareMathSymbol{\Sha}{\mathord}{mcy}{"58}
\DeclareMathOperator{\GL}{GL}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!70!black,
  urlcolor=blue!60!black,
  citecolor=blue!70!black
}

\pagestyle{fancy}
\fancyhead[L]{\small Constructive Reverse Mathematics Series}
\fancyhead[R]{\small Paul Chun-Kit Lee}
\fancyfoot[C]{\thepage}

\titleformat{\section}{\large\bfseries}{Paper \thesection:}{0.5em}{}
\titleformat{\subsection}{\normalsize\itshape}{}{0em}{}

\setcounter{section}{0}

\title{\LARGE\bfseries Constructive Reverse Mathematics Series\\[0.3em]
\Large Series Guide: Papers 1--86 with Abstracts}
\author{Paul Chun-Kit Lee\\New York University\\
\texttt{dr.paul.c.lee@gmail.com}}
\date{February 2026\\[0.5em]
\normalsize Series DOI: \href{https://doi.org/10.5281/zenodo.17054050}{10.5281/zenodo.17054050}}

\begin{document}
\maketitle
\thispagestyle{empty}

\begin{abstract}
\noindent This booklet collects the titles and abstracts of all 98 active papers (of 100 numbered) in the Constructive Reverse Mathematics (CRM) series. The programme calibrates the logical strength of theorems across mathematical physics and arithmetic geometry against the constructive hierarchy: BISH $\subset$ BISH+LLPO $\subset$ BISH+WLPO $\subset$ BISH+LPO $\subset$ CLASS, with independent axes for Markov's Principle (MP), the Fan Theorem (FT), and Dependent Choice (DC). The main finding---proved as a biconditional for motivic cycle-search (Paper~72)---is that the logical cost of mathematics is the logical cost of~$\mathbb{R}$. Papers 45--72 establish the Decidable Polarized Tannakian (DPT) framework, prove its characterization theorem (the DPT axioms are minimal and the Archimedean Principle is a biconditional), and identify the DPT/Fargues--Scholze partition as logically forced. Papers 76--100 apply the CRMLint Squeeze protocol to Hodge theory, BSD, FLT, and K3 surfaces, culminating in Paper~98's Motivic CRM Architecture and Paper~100's Kuga--Satake bifurcation. Approximately 99,000 lines of Lean~4 code formalize the results.
\end{abstract}

\tableofcontents
\newpage

%=============================================================================
% PROGRAMME OVERVIEW
%=============================================================================
\part*{Programme Overview}
\addcontentsline{toc}{part}{Programme Overview}

\subsection*{What is Constructive Reverse Mathematics?}

Constructive Reverse Mathematics (CRM) calibrates the logical strength of mathematical theorems against a hierarchy of constructive principles.  Rather than asking ``is this theorem true?'', CRM asks ``what non-constructive resources does this theorem \emph{require}?''  The answer is a position in the hierarchy
\[
\mathrm{BISH} \;\subset\; \mathrm{BISH{+}LLPO} \;\subset\; \mathrm{BISH{+}WLPO} \;\subset\; \mathrm{BISH{+}LPO} \;\subset\; \mathrm{CLASS},
\]
with independent axes for Markov's Principle (MP), the Fan Theorem (FT), and Dependent Choice (DC).

\begin{itemize}[nosep]
\item \textbf{BISH} (Bishop's constructive mathematics): the base system, requiring explicit witnesses for all existence claims.
\item \textbf{LLPO} (Lesser Limited Principle of Omniscience): decides disjunctions over binary sequences.
\item \textbf{WLPO} (Weak Limited Principle of Omniscience): decides whether a binary sequence is identically zero.
\item \textbf{LPO} (Limited Principle of Omniscience): decides whether a binary sequence contains a 1.
\item \textbf{MP} (Markov's Principle): if a computation cannot fail to halt, then it halts. Orthogonal to the omniscience spine.
\item \textbf{FT} (Fan Theorem): every bar on Cantor space is uniform.  Independent of the omniscience spine.
\item \textbf{DC} (Dependent Choice): sequential choice along a total relation.  Independent of omniscience and FT.
\end{itemize}

\subsection*{The programme in three phases}

\textbf{Phase I: Mathematical Physics (Papers 1--44).}  Systematic calibration of theorems across functional analysis, quantum mechanics, general relativity, statistical mechanics, QFT, quantum information, classical mechanics, cosmology, and quantum foundations.  The main finding: \emph{all empirically accessible predictions in known physics require exactly $\mathrm{BISH{+}LPO}$}.  The Fan Theorem and Dependent Choice are physically dispensable (Papers~30--31).  The LPO cost is genuine and instantiated by Fekete's Subadditive Lemma (Paper~29).

\textbf{Phase II: Arithmetic Geometry (Papers 45--66).}  Extension to the Hodge, Tate, BSD, Fontaine--Mazur, and Weight-Monodromy conjectures.  Each conjecture exhibits a \emph{de-omniscientizing descent}: geometric origin converts LPO-dependent claims to BISH-decidable ones.  The Decidable Polarized Tannakian (DPT) framework (Paper~50) axiomatizes this pattern.  Three invariants---analytic rank, Hodge level, Lang constant---classify the logical cost of cycle-search for any motive.

\textbf{Phase III: Synthesis (Papers 67--72).}  The arithmetic geometry monograph (Paper~67), the CRM audit of Fermat's Last Theorem (Paper~68, result: BISH), the function field Langlands audit (Paper~69, result: BISH), the capstone Archimedean Principle (Paper~70), engineering applications to lattice cryptography (Paper~71), and the DPT Characterization Theorem (Paper~72, result: Archimedean Principle is a biconditional).  The capstone finding: \emph{the logical cost of mathematics is the logical cost of~$\mathbb{R}$}, proved as a biconditional for motivic cycle-search.  DPT and Fargues--Scholze are a logically forced partition: discrete sector ($\mathbb{R}$, BISH) vs.\ continuous sector ($\mathbb{Q}_p$, CLASS).

%=============================================================================
% LOGICAL HIERARCHY DIAGRAM
%=============================================================================
\newpage
\subsection*{The Logical Hierarchy}
\addcontentsline{toc}{section}{The Logical Hierarchy}

\begin{figure}[ht]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=1.4cm and 2.8cm,
  principle/.style={draw, rounded corners, minimum width=2.2cm, minimum height=0.7cm, font=\small\bfseries},
  physlabel/.style={font=\scriptsize, text width=4.5cm, align=center},
  >=stealth
]
% Main spine
\node[principle] (bish) {BISH};
\node[principle, above=1.8cm of bish] (llpo) {LLPO};
\node[principle, above=of llpo] (wlpo) {WLPO};
\node[principle, above=of wlpo] (lpo) {LPO};
\node[principle, above=of lpo] (undec) {Undecidable};

% Choice hierarchy
\node[principle, above left=1.0cm and 2.2cm of bish] (cc) {CC};
\node[principle, above=0.8cm of cc] (dc) {DC};
% Orthogonal branch
\node[principle, above right=1.2cm and 2.2cm of bish] (mp) {MP};

% Fan Theorem: independent branch
\node[principle, below right=0.4cm and 3.5cm of bish] (ft) {FT};

% Edges (Hasse diagram)
\draw[thick, ->] (bish) -- (cc);
\draw[thick, ->] (cc) -- (dc);
\draw[thick, ->] (bish) -- (mp);
\draw[thick, ->] (bish) -- (llpo);
\draw[thick, ->] (llpo) -- (wlpo);
\draw[thick, ->] (wlpo) -- (lpo);
\draw[thick, ->] (lpo) -- (undec);
\draw[thick, ->] (bish) -- (ft);

% Physical labels
\node[physlabel, below=0.15cm of bish] {Finite physics, prep.\ HUP,\\Bell nonlocality, entanglement,\\decoherence bounds, local conservation,\\WKB amplitudes, discrete EL eqs.,\\RG flow, \textbf{Fermat's Last Thm},\\\textbf{function field Langlands}};
\node[physlabel, left=0.1cm of cc] {Mean ergodic ($\equiv$),\\weak LLN};
\node[physlabel, left=0.1cm of dc] {Birkhoff ergodic ($\equiv$),\\measurement HUP, SLLN};
\node[physlabel, right=0.1cm of mp] {Exact spectral\\membership,\\radioactive decay,\\\textbf{cycle search (rank $\geq2$)}};
\node[physlabel, right=0.4cm of llpo] {WKB turning points,\\Bell disjunction,\\KS sign decision,\\Bell angle/IVT};
\node[physlabel, right=0.4cm of wlpo] {Singular states, bidual gap,\\phase classification,\\threshold decoupling,\\\textbf{Arthur--Selberg trace formula}};
\node[physlabel, right=0.4cm of lpo] {Fekete's lemma,\\thermodynamic limit,\\geodesic incompleteness,\\exact decoherence, global energy,\\\textbf{Hodge conj.\ (dim $\geq4$)}};
\node[physlabel, right=0.4cm of undec] {Spectral gap decidability};
\node[physlabel, below=0.1cm of ft] {Compact optimization\\(EVT),\\action minimizer};

% Independence annotations
\draw[dashed, gray] (dc) -- node[above, font=\tiny, gray] {independent} (mp);
\draw[dashed, gray] (ft.north west) -- node[below, font=\tiny, gray, sloped] {independent} (mp.south east);

\end{tikzpicture}%
}
\caption{The logical geography of the CRM programme (Papers~1--72).  Arrows indicate strict implication over BISH.  The omniscience spine (BISH $<$ LLPO $<$ WLPO $<$ LPO) is the dominant vertical chain.  DC and MP occupy orthogonal positions.  FT is a third independent branch.  \textbf{Bold entries} are from the arithmetic geometry and synthesis phases.}
\label{fig:hierarchy}
\end{figure}

%=============================================================================
% MASTER CALIBRATION TABLE
%=============================================================================
\newpage
\subsection*{Master Calibration Table}
\addcontentsline{toc}{section}{Master Calibration Table}

The table below collects calibrated entries from 72 papers spanning mathematical physics, arithmetic geometry, and the Langlands programme.

\begin{center}
\begin{longtable}{@{}p{5.8cm}ll@{}}
\toprule
\textbf{Theorem / Result} & \textbf{Level} & \textbf{Source} \\
\midrule
\endfirsthead
\toprule
\textbf{Theorem / Result} & \textbf{Level} & \textbf{Source} \\
\midrule
\endhead
\multicolumn{3}{l}{\textit{BISH (fully constructive):}} \\
Finite-volume physics, finite-size bounds & BISH & Papers 8, 9 \\
Preparation uncertainty (HUP) & BISH & Paper 6 \\
Bell nonlocality (CHSH/Tsirelson) & BISH & Paper 11 \\
Entanglement entropy & BISH & Paper 11 \\
Schwarzschild interior (finite-time) & BISH & Paper 13 \\
Decoherence bounds (finite-step) & BISH & Paper 14 \\
Local conservation (Noether) & BISH & Paper 15 \\
Born probability (finite-dim) & BISH & Paper 16 \\
BH entropy (algebraic correction) & BISH & Paper 17 \\
WKB amplitude (specific barriers) & BISH & Paper 19 \\
Discrete Euler--Lagrange equations & BISH & Paper 28 \\
Yukawa RG one-loop flow & BISH & Paper 18 \\
QED Landau pole, Ward--Takahashi & BISH & Paper 32 \\
Fixed-order scattering (Bhabha) & BISH & Paper 34 \\
Hodge polarization degeneration & BISH & Paper 45 \\
Algebraic cycle verification & BISH & Paper 49 \\
CM elliptic motives (decidable) & BISH & Paper 53 \\
Fermat's Last Theorem & BISH & Paper 68 \\
Function field Langlands ($\mathrm{GL}_n$, gen.\ $G$) & BISH & Paper 69 \\
Lattice cryptography (LLL, exp.\ approx.) & BISH & Paper 71 \\
Cycle-search (positive-definite height) & BISH & Paper 72 \\
\midrule
\multicolumn{3}{l}{\textit{Choice hierarchy (CC $<$ DC):}} \\
Mean ergodic theorem & $\equiv$ CC & Paper 25 \\
Birkhoff's ergodic theorem & $\equiv$ DC & Paper 25 \\
Frequentist convergence (SLLN) & $\leq$ DC & Papers 16, 25 \\
\midrule
\multicolumn{3}{l}{\textit{LLPO:}} \\
WKB turning-point decision (IVT) & $\equiv$ LLPO & Paper 19 \\
Bell disjunctive conclusion & $\equiv$ LLPO & Paper 21 \\
KS sign decision (contextuality) & $\equiv$ LLPO & Paper 24 \\
Bell angle optimisation (IVT) & $\equiv$ LLPO & Paper 27 \\
\midrule
\multicolumn{3}{l}{\textit{MP (Markov's Principle):}} \\
Exact spectral membership & MP & Paper 4 \\
Radioactive decay (eventual detection) & $\equiv$ MP & Paper 22 \\
Cycle search, rank $\geq 2$ (without Lang) & MP & Paper 61 \\
SVP polynomial approx.\ (BKZ) & BISH+MP & Paper 71 \\
\midrule
\multicolumn{3}{l}{\textit{WLPO:}} \\
Bidual-gap / singular states & $\equiv$ WLPO & Papers 2, 7 \\
Phase classification (magnetization) & $\equiv$ WLPO & Paper 20 \\
Step-function threshold decoupling & WLPO & Paper 18 \\
Arthur--Selberg trace formula & WLPO & Paper 68 \\
\midrule
\multicolumn{3}{l}{\textit{FT (Fan Theorem) --- physically dispensable:}} \\
Compact optimization (EVT on $[a,b]$) & $\equiv$ FT & Paper 23 \\
Action minimizer existence (variational) & $\equiv$ FT & Paper 28 \\
\midrule
\multicolumn{3}{l}{\textit{LPO:}} \\
Thermodynamic limit existence & $\equiv$ LPO & Papers 8, 9, 29 \\
Fekete's lemma (subadditive sequences) & $\equiv$ LPO & Paper 29 \\
Geodesic incompleteness (completed limit) & $\equiv$ LPO & Paper 13 \\
Exact decoherence (wave fn collapse) & $\equiv$ LPO & Paper 14 \\
Global energy (infinite-volume) & $\equiv$ LPO & Paper 15 \\
BH entropy density convergence & $\equiv$ LPO & Paper 17 \\
Galois-invariance decidability ($\mathbb{Q}_\ell$) & $\equiv$ LPO & Paper 46 \\
$L(E,1)=0$ decision & $\equiv$ LPO & Paper 48 \\
Hodge type $(r,r)$ decidability ($\mathbb{C}$) & $\equiv$ LPO & Paper 49 \\
Cycle-search (indefinite height) & $\equiv$ LPO & Paper 72 \\
\midrule
\multicolumn{3}{l}{\textit{Undecidable:}} \\
Spectral gap decidability & $\Sigma_1^0$-complete & Cubitt et al. \\
Generic spectral gap (no promise) & $\Sigma_2^0$-complete & Paper 39 \\
\bottomrule
\end{longtable}
\end{center}

%=============================================================================
% DPT FRAMEWORK
%=============================================================================
\subsection*{The DPT Framework (Paper 50)}
\addcontentsline{toc}{section}{The DPT Framework}

A \emph{Decidable Polarized Tannakian} (DPT) category over $\mathbb{Q}$ is a $\mathbb{Q}$-linear abelian symmetric monoidal category $\mathcal{C}$ equipped with three axioms:

\begin{description}[style=unboxed,leftmargin=1em]
\item[Axiom 1 (Decidable morphisms --- Standard Conjecture D).] For all $X, Y \in \mathcal{C}$, the morphism space $\mathrm{Hom}(X,Y)$ has decidable equality: $\forall f,g\colon X \to Y,\; f = g \lor f \neq g$.

\item[Axiom 2 (Algebraic spectrum).] Every endomorphism $f \in \mathrm{End}(X)$ satisfies a monic polynomial $p \in \mathbb{Z}[t]$, forcing eigenvalues into $\overline{\mathbb{Q}}$.

\item[Axiom 3 (Archimedean polarization).] A faithful functor to real vector spaces equipped with a positive-definite bilinear form: $\langle x, x \rangle > 0$ for all $x \neq 0$.
\end{description}

\noindent Five conjectures (Hodge, Tate, BSD, Fontaine--Mazur, Weight-Monodromy) exhibit a uniform \emph{de-omniscientizing descent}: geometric origin converts LPO-dependent data to BISH-decidable numerical equivalence, mediated by the DPT axioms.

%=============================================================================
% ARCHIMEDEAN PRINCIPLE
%=============================================================================
\subsection*{The Archimedean Principle (Paper 70)}
\addcontentsline{toc}{section}{The Archimedean Principle}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    box/.style={draw, rounded corners=3pt, minimum width=4.8cm,
                minimum height=1.8cm, align=center, font=\small},
    arr/.style={-{Stealth[length=5pt]}, thick},
    ann/.style={font=\footnotesize\itshape, text=black!70}
  ]
  % Column headers
  \node[font=\bfseries] at (0, 2.6) {Archimedean};
  \node[font=\bfseries] at (6.2, 2.6) {Non-Archimedean};
  % Row headers
  \node[font=\bfseries, rotate=90, anchor=south] at (-3.3, 1.0) {Physics};
  \node[font=\bfseries, rotate=90, anchor=south] at (-3.3,-1.8) {Arithmetic};
  % Four boxes
  \node[box] (CP) at (0, 1.0)
    {Continuum physics\\[2pt] $\mathrm{LPO} \;\xrightarrow{\;\text{projection}\;}\; \mathrm{BISH}$};
  \node[box] (LP) at (6.2, 1.0)
    {Lattice physics\\[2pt] $\mathrm{BISH}$};
  \node[box] (NA) at (0,-1.8)
    {Number field arithmetic\\[2pt] $\mathrm{LPO} \;\xrightarrow{\;\text{search}\;}\; \mathrm{BISH} + \mathrm{MP}$};
  \node[box] (FA) at (6.2,-1.8)
    {Function field arithmetic\\[2pt] $\mathrm{BISH}$};
  % Horizontal arrows
  \draw[arr] (CP.east) -- (LP.west) node[midway,above,ann] {remove $\mathbb{R}$};
  \draw[arr] (NA.east) -- (FA.west) node[midway,above,ann] {remove $\mathbb{R}$};
  % u-invariant annotations
  \node[ann] at (0, -3.2) {$u(\mathbb{R}) = \infty$};
  \node[ann] at (6.2, -3.2) {$u(\mathbb{Q}_p) = 4$};
  % MP gap brace
  \draw[decorate, decoration={brace, amplitude=5pt, mirror}, thick]
    (-2.6, 0.0) -- (-2.6, -1.0)
    node[midway, left=6pt, font=\footnotesize] {MP gap};
\end{tikzpicture}
\caption{The four-domain parameterisation (Paper~70).  The CRM level is determined by one parameter: presence of the Archimedean place.  The descent type (projection vs.\ search) determines the residual.}
\label{fig:archimedean}
\end{figure}

\noindent The mechanism is $u(\mathbb{R}) = \infty$: the real numbers are the only completion of~$\mathbb{Q}$ where positive-definite forms exist in every dimension.  Three fields independently exploit this via the same architecture: \emph{Hilbert space inner product} (physics), \emph{Rosati involution} (arithmetic geometry), \emph{Petersson inner product} (automorphic forms).  Remove $\mathbb{R}$ and both physics and arithmetic collapse to BISH.  The difference between them is how they descend: physics uses \emph{projection} (eliminating MP), number theory uses \emph{search} (preserving MP as Diophantine hardness).

%=============================================================================
% FORMALIZATION STATISTICS
%=============================================================================
\subsection*{Formalization Statistics}
\addcontentsline{toc}{section}{Formalization Statistics}

\begin{center}
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Total papers (numbered) & 72 \\
Active papers & 70 \\
Withdrawn & 2 (Papers 1, 3) \\
Retired (merged) & 2 (Papers 60, 62) \\
Papers with Zenodo DOI & 67 \\
Papers with Lean~4 formalization & $\sim$62 \\
Total Lean~4 lines & $\sim$88,350+ \\
Papers with zero \texttt{sorry} & $\geq$22 \\
Lean toolchain & v4.28.0-rc1 \\
Mathlib dependency & v4.28.0-rc1 \\
\midrule
\textbf{Domains covered} & \\
\midrule
\multicolumn{2}{l}{Functional Analysis, Quantum Mechanics, General Relativity,} \\
\multicolumn{2}{l}{Statistical Mechanics, QFT (QED/QCD), Quantum Information,} \\
\multicolumn{2}{l}{Classical Mechanics, Foundations, Undecidability, AdS/CFT,} \\
\multicolumn{2}{l}{Cosmology, Arithmetic Geometry, Motives, Number Theory,} \\
\multicolumn{2}{l}{Langlands Programme, Cryptography} \\
\bottomrule
\end{tabular}
\end{center}

\newpage

%=============================================================================
% PART I: FOUNDATIONS (Papers 1--6)
%=============================================================================
\part*{Part I: Foundations (Papers 1--6)}
\addcontentsline{toc}{part}{Part I: Foundations}

%-----------------------------------------------------------------------------
\setcounter{section}{0}
\section{Rank-One Toggle Kernel}
\label{sec:p1}
\textbf{Status:} Withdrawn.\\
\textbf{DOI:} ---

\bigskip
\noindent\textit{This paper has been withdrawn from the series.}

%-----------------------------------------------------------------------------
\section{The Bidual Gap and WLPO}
\label{sec:p2}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18501689}{10.5281/zenodo.18501689}

\bigskip
\noindent\textbf{Abstract.}
We formalize the bidual gap of Banach spaces and prove it is equivalent to WLPO (the Weak Limited Principle of Omniscience). Gap detection requires exactly WLPO; the witness space is $c_0$ (sequences vanishing at infinity). The formalization establishes that Banach space non-reflexivity detection is a constructively calibrated problem with precise logical cost.

%-----------------------------------------------------------------------------
\section{[Withdrawn]}
\label{sec:p3}
\textbf{Status:} Withdrawn.\\
\textbf{DOI:} ---

\bigskip
\noindent\textit{This paper has been withdrawn from the series.}

%-----------------------------------------------------------------------------
\section{Axiom Calibration for Quantum Spectra}
\label{sec:p4}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.17118223}{10.5281/zenodo.17118223}

\bigskip
\noindent\textbf{Abstract.}
Five spectral properties---from approximate eigenvalues to exact membership---are stratified from BISH to WLPO+MP. This provides a complete calibration of quantum spectral theory against the constructive hierarchy, showing that different spectral questions about the same operator carry different logical costs.

%-----------------------------------------------------------------------------
\section{Schwarzschild Curvature Verification}
\label{sec:p5}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18489703}{10.5281/zenodo.18489703}

\bigskip
\noindent\textbf{Abstract.}
We make Axiom Calibration the organizing principle for a foundations-first study of General Relativity. The paper contributes three AxCal instruments for GR: (I)~witness families pinned to a fixed $\Sigma_0^{\mathrm{GR}}$ signature; (II)~proof-route flags and portal theorems; and (III)~HeightCertificates. We calibrate five loci: G1 (explicit vacuum checks, Height~0), G2 (Cauchy/MGHD), G3 (singularity theorems), G4 (maximal extensions), and G5 (computable evolution).

%-----------------------------------------------------------------------------
\section{Heisenberg Uncertainty Principle (v2, CRM over Mathlib)}
\label{sec:p6}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18519836}{10.5281/zenodo.18519836}

\bigskip
\noindent\textbf{Abstract.}
We formalize the Robertson--Schr\"odinger and Schr\"odinger uncertainty inequalities in Lean~4 using Mathlib's InnerProductSpace API, within the framework of Constructive Reverse Mathematics. Both preparation-uncertainty inequalities are proved at Height~0 (fully constructive) using Cauchy--Schwarz, centered-vector decomposition, and elementary complex-number identities---all drawn from Mathlib with zero custom axioms. Measurement uncertainty, which requires constructing infinite measurement histories, is calibrated at $\mathrm{DC}_\omega$-height.

%=============================================================================
% PART II: PHYSICAL CALIBRATIONS (Papers 7--28)
%=============================================================================
\newpage
\part*{Part II: Physical Calibrations (Papers 7--28)}
\addcontentsline{toc}{part}{Part II: Physical Calibrations}

%-----------------------------------------------------------------------------
\section{Physical Bidual Gap, Trace-Class Operators}
\label{sec:p7}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18527559}{10.5281/zenodo.18527559}

\bigskip
\noindent\textbf{Abstract.}
We present a Lean~4 formalization of the Physical Bidual Gap Theorem, establishing Banach space non-reflexivity of the trace-class operators $\mathcal{S}^1$ on a separable Hilbert space and showing that any constructive witness of this non-reflexivity implies WLPO. In the algebraic formulation of quantum mechanics, $\mathcal{S}^1$ is the natural state space; our result shows that the gap between physical density matrices and the ``generalized states'' in $(\mathcal{S}^1)^{**}$ is constructively inaccessible. The formalization comprises 754 lines of Lean~4 code across 8 files.

%-----------------------------------------------------------------------------
\section{1D Ising Model and LPO}
\label{sec:p8}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18516813}{10.5281/zenodo.18516813}

\bigskip
\noindent\textbf{Abstract.}
We prove two complementary results about the thermodynamic limit of the one-dimensional Ising model. (A)~The finite-size error bound $|f_N(\beta)-f_\infty(\beta)|\le(1/N)\tanh(\beta)^N$ is provable in BISH without any omniscience principle. (B)~The existence of the thermodynamic limit as a completed real number is equivalent over BISH to LPO, via bounded monotone convergence instantiated through the Ising free energy. Together: the LPO cost is genuine and dispensable. The combined formalization comprises 1374 lines of Lean across 18 modules with zero sorries.

%-----------------------------------------------------------------------------
\section{Ising Formulation-Invariance}
\label{sec:p9}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18517570}{10.5281/zenodo.18517570}

\bigskip
\noindent\textbf{Abstract.}
We prove that the logical cost of the thermodynamic limit for the 1D Ising model is formulation-invariant: the same axiom profile arises from a purely combinatorial derivation as from the transfer-matrix approach of Paper~8. Specifically, we re-derive (A)~BISH dispensability and (B)~LPO equivalence using only finite sums over $\{-1,+1\}^N$, the binomial parity sieve, and elementary hyperbolic function arithmetic. No transfer matrices, eigenvalues, or linear algebra are used. The formalization comprises 1319 lines across 18 modules with zero sorries.

%-----------------------------------------------------------------------------
\section{Logical Geography of Mathematical Physics}
\label{sec:p10}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18636180}{10.5281/zenodo.18636180}

\bigskip
\noindent\textbf{Abstract.}
This paper synthesizes the machine-verified results of the constructive calibration programme---spanning functional analysis, quantum spectra, general relativity, uncertainty relations, statistical mechanics, quantum information, decoherence, conservation laws, measurement statistics, quantum gravity, particle physics, semiclassical mechanics, quantum foundations, radioactive decay, ergodic theory, classical mechanics, and subadditive convergence---into a single interpretive framework. The calibration table contains approximately fifty entries spanning BISH to undecidable, populated by Lean~4 formalizations totalling approximately 25,400 lines of code across eleven domains.

%-----------------------------------------------------------------------------
\section{Entanglement, CHSH, Tsirelson Bound}
\label{sec:p11}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18527676}{10.5281/zenodo.18527676}

\bigskip
\noindent\textbf{Abstract.}
We provide a complete Lean~4 formalization of two foundational results in quantum information theory: (A)~the Tsirelson bound on the CHSH operator ($|\langle\psi,\mathrm{CHSH}\,\psi\rangle|\le 2\sqrt{2}$); and (B)~the entanglement entropy of the Bell singlet state ($S(\rho_A)=\log 2$). The formalization comprises 639 lines across 8 modules with zero sorry.

%-----------------------------------------------------------------------------
\section{Constructive History of Mathematical Physics}
\label{sec:p12}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18636250}{10.5281/zenodo.18636250}

\bigskip
\noindent\textbf{Abstract.}
This essay tells the story of 150 years of mathematical physics from an unfamiliar angle. A machine-verified research programme provides preliminary evidence that for every exactly solvable model and system with explicit finite-size bounds examined so far, empirical predictions can be derived using only the most elementary constructive logic. Paper~29 establishes a fundamental qualification: Fekete's Subadditive Lemma is equivalent to LPO, making the logical cost ineliminable for systems near phase transitions.

%-----------------------------------------------------------------------------
\section{Event Horizon as Logical Boundary}
\label{sec:p13}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18529007}{10.5281/zenodo.18529007}

\bigskip
\noindent\textbf{Abstract.}
The Schwarzschild interior decomposes constructively: the cycloid geodesic reaching $r=0$ at finite proper time is BISH; the completed-limit singularity assertion requires LPO. The event horizon demarcates a logical boundary; interior geometry is BISH, the singularity is LPO. Formalization: 1,021 lines, 8 modules.

%-----------------------------------------------------------------------------
\section{Quantum Decoherence}
\label{sec:p14}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18569068}{10.5281/zenodo.18569068}

\bigskip
\noindent\textbf{Abstract.}
A single qubit coupled to $N$ environmental qubits via controlled-rotation unitaries undergoes BISH-level decoherence; the completed-limit exact decoherence costs LPO. This is the third domain exhibiting the BMC $\leftrightarrow$ LPO pattern (after Ising and Schwarzschild).

%-----------------------------------------------------------------------------
\section{Noether's Theorem}
\label{sec:p15}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18572494}{10.5281/zenodo.18572494}

\bigskip
\noindent\textbf{Abstract.}
The local conservation law ($\partial_\mu J^\mu=0$) is BISH; global energy existence costs LPO via BMC. This is the fourth domain exhibiting the BISH/LPO boundary at BMC, and the first calibration of a structural (as opposed to phenomenological) physical law.  520 lines.

%-----------------------------------------------------------------------------
\section{Born Rule (Technical Note)}
\label{sec:p16}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18575377}{10.5281/zenodo.18575377}

\bigskip
\noindent\textbf{Abstract.}
Single-trial probability and the weak law of large numbers are BISH; frequentist convergence (the strong law) requires $\mathrm{DC}_\omega$. This fills the DC axis of the calibration table with physical content. 564 lines.

%-----------------------------------------------------------------------------
\section{Bekenstein--Hawking Formula}
\label{sec:p17}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18597306}{10.5281/zenodo.18597306}

\bigskip
\noindent\textbf{Abstract.}
Finite entropy computation is BISH; convergence of entropy densities via BMC costs LPO; the subleading logarithmic correction ($-3/2$) is BISH. This is the fifth independent physics domain exhibiting the BISH/LPO boundary, and the first CRM application to quantum gravity. 1,804 lines, 20 files.

%-----------------------------------------------------------------------------
\section{Yukawa RG Constructive Stratification}
\label{sec:p18}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18626839}{10.5281/zenodo.18626839}

\bigskip
\noindent\textbf{Abstract.}
Standard Model one-loop stratification. The discrete RG step and fixed-precision coupling are BISH; threshold crossings cost WLPO; the global coupling across all thresholds costs LPO via BMC.

%-----------------------------------------------------------------------------
\section{WKB Tunneling and LLPO}
\label{sec:p19}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18602596}{10.5281/zenodo.18602596}

\bigskip
\noindent\textbf{Abstract.}
Three tiers: (BISH)~WKB amplitude for algebraic barriers; (LLPO)~classical turning points via IVT; (LPO)~full semiclassical limit. This is the first physical LLPO calibration in the series---quantum tunneling sits strictly between BISH and WLPO. 1,081 lines, 15 files.

%-----------------------------------------------------------------------------
\section{Observable-Dependent Logical Cost: 1D Ising Magnetization}
\label{sec:p20}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18603079}{10.5281/zenodo.18603079}

\bigskip
\noindent\textbf{Abstract.}
Phase classification via magnetization zero-test costs WLPO; the same 1D Ising model requires BISH (finite), WLPO (phase), LPO (thermodynamic limit). Observable-dependent cost established: the logical cost is a property of the \emph{question}, not the system. 494 lines, 12 files.

%-----------------------------------------------------------------------------
\section{Bell Nonlocality and LLPO}
\label{sec:p21}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18603251}{10.5281/zenodo.18603251}

\bigskip
\noindent\textbf{Abstract.}
The step from ``local realism refuted'' (BISH via CHSH) to ``disjunctive conclusion'' (LLPO). Three-level stratification: BISH (CHSH bound), LLPO (disjunctive decision), WLPO (hierarchy). First CRM calibration of quantum foundations at LLPO level. 751 lines, 14 files.

%-----------------------------------------------------------------------------
\section{Markov's Principle and Radioactive Decay}
\label{sec:p22}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18603503}{10.5281/zenodo.18603503}

\bigskip
\noindent\textbf{Abstract.}
``Non-zero decay rate $\Rightarrow$ eventual detection'' is equivalent to Markov's Principle. First CRM calibration at MP level; the hierarchy becomes a partial order, not a linear chain. 814 lines, 12 files.

%-----------------------------------------------------------------------------
\section{Fan Theorem and Optimization}
\label{sec:p23}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18604312}{10.5281/zenodo.18604312}

\bigskip
\noindent\textbf{Abstract.}
The Extreme Value Theorem on $[a,b]$ requires exactly the Fan Theorem; instantiated via 1D Ising free energy. First FT calibration; third independent branch. The 1D Ising now has four distinct costs: BISH (finite), LPO (thermodynamic), WLPO (phase), FT (optimization). $\sim$680 lines, 14 files.

%-----------------------------------------------------------------------------
\section{Kochen--Specker Contextuality and LLPO}
\label{sec:p24}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18604317}{10.5281/zenodo.18604317}

\bigskip
\noindent\textbf{Abstract.}
KS uncolorability is BISH; sign decision costs LLPO. Structural identity: Bell sign decision $\equiv$ KS sign decision (both LLPO). Physically distinct no-go theorems share identical logical cost. 887 lines, 16 files.

%-----------------------------------------------------------------------------
\section{Choice Axis: Ergodic Theorems}
\label{sec:p25}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18615453}{10.5281/zenodo.18615453}

\bigskip
\noindent\textbf{Abstract.}
Mean ergodic theorem $\leftrightarrow$ Countable Choice; Birkhoff pointwise $\leftrightarrow$ Dependent Choice. Clean separation: ensemble average requires CC, individual trajectory requires DC. DC Ceiling Thesis: no calibratable physical theorem requires more than DC. 1,805 lines, 12 modules.

%-----------------------------------------------------------------------------
\section{Bidual Gap Arithmetic Route, WLPO}
\label{sec:p26}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18615457}{10.5281/zenodo.18615457}

\bigskip
\noindent\textbf{Abstract.}
Two independent proofs of WLPO-completeness for bidual gap detection: functional-analytic (Paper~2) and arithmetic (via G\"odel sequences, Lindenbaum algebra of $\Pi_1^0$ sentences). Robustness evidence that the classification is intrinsic. 1,213 lines, 30 axioms.

%-----------------------------------------------------------------------------
\section{Bell Angle Optimisation via IVT, LLPO}
\label{sec:p27}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18615459}{10.5281/zenodo.18615459}

\bigskip
\noindent\textbf{Abstract.}
$\text{LLPO}\leftrightarrow\text{Exact IVT}$; Bell angle optimization reduces to IVT instances. Identifies the mechanism (IVT) explaining why LLPO appears in Bell physics. Bell angle-finding is strictly below gap detection (WLPO). Zero sorries, six axioms.

%-----------------------------------------------------------------------------
\section{Newton vs.\ Lagrange vs.\ Hamilton}
\label{sec:p28}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18616620}{10.5281/zenodo.18616620}

\bigskip
\noindent\textbf{Abstract.}
Three formulations of classical mechanics are constructively stratified: Newtonian/Lagrangian equations (BISH), Hamiltonian discrete form (BISH), action minimization (FT). The variational interpretation is logically dispensable. First formal proof that variational mechanics is not logically necessary. 621 lines, zero custom axioms.


%=============================================================================
% PART III: CEILING AND DISPENSABILITY (Papers 29--35)
%=============================================================================
\newpage
\part*{Part III: Ceiling and Dispensability (Papers 29--35)}
\addcontentsline{toc}{part}{Part III: Ceiling and Dispensability}

%-----------------------------------------------------------------------------
\section{Fekete's Subadditive Lemma and LPO}
\label{sec:p29}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18643617}{10.5281/zenodo.18643617}

\bigskip
\noindent\textbf{Abstract.}
Fekete's Subadditive Lemma is equivalent over BISH to LPO. The backward direction encodes a binary sequence into a mock free energy $F_n=-n\cdot x_n$; applying Fekete yields a limit deciding the sequence. The forward direction composes $\text{LPO}\to\text{BMC}$ with the classical Fekete proof. This resolves Problem~1 of the calibration programme and establishes a three-tier hierarchy for thermodynamic-limit convergence: exact solvability (BISH), cluster expansions (BISH), generic subadditivity (LPO). 549 lines, 6 modules, zero sorries.

%-----------------------------------------------------------------------------
\section{Physical Dispensability of the Fan Theorem}
\label{sec:p30}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18638394}{10.5281/zenodo.18638394}

\bigskip
\noindent\textbf{Abstract.}
Every empirically accessible prediction from FT-calibrated results is recoverable in BISH+LPO without the Fan Theorem. Three pillars: (1)~LPO implies BMC yielding $\varepsilon$-approximate witnesses; (2)~equations of motion are BISH-valid without any minimizer; (3)~no finite experiment distinguishes exact from $\varepsilon$-approximate minimizers. FT captures exact optimizer existence; LPO captures convergent approximation. FT is physically dispensable.

%-----------------------------------------------------------------------------
\section{Physical Dispensability of Dependent Choice}
\label{sec:p31}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18645578}{10.5281/zenodo.18645578}

\bigskip
\noindent\textbf{Abstract.}
Every empirical prediction from DC-calibrated results is recoverable in BISH+LPO without DC. DC's content is a quantifier swap (outside measure vs.\ inside measure); the experimenter observes after quantifier choice. Together with Papers~29--30: the logical constitution of empirically accessible physics is BISH+LPO.

%-----------------------------------------------------------------------------
\section{QED One-Loop Renormalization: The Landau Pole}
\label{sec:p32}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642598}{10.5281/zenodo.18642598}

\bigskip
\noindent\textbf{Abstract.}
Complete CRM calibration of QED one-loop renormalization. The discrete RG step, finite-precision coupling, Ward--Takahashi identity, and---surprisingly---the Landau pole divergence itself are all BISH. Threshold crossings require WLPO; global coupling requires LPO. The Landau pole, naively most ``non-constructive,'' is fully BISH: the closed-form solution provides an explicit Cauchy modulus requiring no omniscience.

%-----------------------------------------------------------------------------
\section{QCD One-Loop Renormalization and Confinement}
\label{sec:p33}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642610}{10.5281/zenodo.18642610}

\bigskip
\noindent\textbf{Abstract.}
Asymptotic freedom ($\beta<0$); IR divergence at $\Lambda_{\mathrm{QCD}}$ is BISH; continuum limit costs LPO; mass gap decision costs WLPO; extracting positivity costs MP. Since LPO strictly implies both WLPO and MP, confinement is \emph{free}: the LPO already paid for the continuum limit automatically subsidizes the mass gap.

%-----------------------------------------------------------------------------
\section{Scattering Amplitudes Are Constructively Computable}
\label{sec:p34}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642612}{10.5281/zenodo.18642612}

\bigskip
\noindent\textbf{Abstract.}
The fixed-order inclusive cross section (Bhabha scattering $e^+e^-\to e^+e^-$) is pure BISH: finite composition of computable functions with UV divergences removed by $\overline{\mathrm{MS}}$ subtraction and IR cancellation by the Bloch--Nordsieck theorem. Only the all-orders series sum requires LPO.

%-----------------------------------------------------------------------------
\section{The Logical Constitution of Empirical Physics: A Conservation Metatheorem}
\label{sec:p35}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642616}{10.5281/zenodo.18642616}

\bigskip
\noindent\textbf{Abstract.}
Four components: (A)~BISH Conservation: finite compositions of computable functions are BISH; (B)~LPO Boundary: bounded monotone limits without modulus cost LPO; (C)~Exhaustiveness: all 38 calibration entries are $\le$ LPO; (D)~Three Mechanisms (BMC, Cauchy completeness, supremum existence) are mutually equivalent. Establishes the BISH+LPO ceiling.


%=============================================================================
% PART IV: UNDECIDABILITY AND BEYOND (Papers 36--44)
%=============================================================================
\newpage
\part*{Part IV: Undecidability and Beyond (Papers 36--44)}
\addcontentsline{toc}{part}{Part IV: Undecidability and Beyond}

%-----------------------------------------------------------------------------
\section{Stratifying Spectral Gap Undecidability: Cubitt's Theorem Is LPO}
\label{sec:p36}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642620}{10.5281/zenodo.18642620}

\bigskip
\noindent\textbf{Abstract.}
Cubitt's spectral gap undecidability is Turing--Weihrauch equivalent to LPO. Stratification: finite-volume gap (BISH), thermodynamic limit (LPO), each instance (LPO-decidable), physical zero-test (WLPO), uniform function (LPO-computable). Quantum undecidability introduces zero additional resources beyond thermodynamic limits.

%-----------------------------------------------------------------------------
\section{The Undecidability Landscape Is LPO}
\label{sec:p37}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642802}{10.5281/zenodo.18642802}

\bigskip
\noindent\textbf{Abstract.}
Extension to three further undecidability results: phase diagram uncomputability, 1D spectral gap, uncomputable RG flows. Meta-theorem: any physical undecidability from computable many-one reduction from the halting problem is LPO. Watson--Cubitt ground state energy density hardness is BISH (computational complexity, not logical undecidability). 660 lines.

%-----------------------------------------------------------------------------
\section{Wang Tiling and the Origin of Physical Undecidability}
\label{sec:p38}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642804}{10.5281/zenodo.18642804}

\bigskip
\noindent\textbf{Abstract.}
Every undecidability result in quantum many-body physics descends from the Wang tiling problem (Berger 1966), which is LPO. All descendants---from Kanter (1990) through Cubitt (2015)---inherit exactly LPO. The $\Sigma_1^0$ ceiling: no $\Sigma_1^0$-complete reduction can exceed LPO. 573 lines.

%-----------------------------------------------------------------------------
\section{Physics Reaches $\Sigma_2^0$: The Physical Stratification of Undecidability}
\label{sec:p39}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18642806}{10.5281/zenodo.18642806}

\bigskip
\noindent\textbf{Abstract.}
The ceiling is not $\Sigma_1^0$. The generic spectral gap (without promise gap) is $\Sigma_2^0/\Pi_2^0$-complete, requiring $\text{LPO}'$ (the Turing jump of LPO). The Physical Stratification Theorem: arithmetic complexity is determined by epistemic mode (exact vs promise-gapped vs empirical), not thermodynamic scaling. Platonic exact observables reach $\text{LPO}'$ ($\Sigma_2^0$); promise-gapped and empirical observables cap at LPO ($\Sigma_1^0$). 873 lines.

%-----------------------------------------------------------------------------
\section{The Logical Constitution of Physical Reality (Monograph)}
\label{sec:p40}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18654773}{10.5281/zenodo.18654773}

\bigskip
\noindent\textbf{Abstract.}
We prove that the logical resources for all empirical predictions in known physics are exactly BISH+LPO. Established by systematic axiom calibration across 42 papers spanning the Standard Model, general relativity, statistical mechanics, quantum information, AdS/CFT, and the cosmological constant problem. Three foundational results: (1)~Fekete's lemma $\leftrightarrow$ LPO (LPO is physically instantiated); (2)~the Fan Theorem is dispensable; (3)~Dependent Choice is dispensable. All calibrations formally verified in $\sim$35,000 lines of Lean~4 code.

%-----------------------------------------------------------------------------
\section{AdS/CFT Diagnostic}
\label{sec:p41}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18654780}{10.5281/zenodo.18654780}

\bigskip
\noindent\textbf{Abstract.}
The holographic dictionary is an axiom-preserving map: bulk and boundary carry identical axiom cost at every level. Vacuum AdS$_3$ (BISH), thermal BTZ (BISH entropy, LLPO phase), FLM correction (BISH), QES entropy (LPO), Page curve (BISH). Holography projects away the Fan Theorem: boundary entropy is computable at BISH+LPO without constructing the bulk surface. 955 lines.

%-----------------------------------------------------------------------------
\section{The Cosmological Constant Problem}
\label{sec:p42}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18654789}{10.5281/zenodo.18654789}

\bigskip
\noindent\textbf{Abstract.}
The $10^{120}$ discrepancy decomposes into three claims: (I)~the UV discrepancy is a regulator artifact (vanishes under dimensional regularization); (II)~``naturalness'' is a Bayesian prior, outside the hierarchy; (III)~the genuine constraint (55-decimal cancellation) requires the thermodynamic limit, hence LPO. The cosmological constant problem introduces no new logical resources. $\sim$830 lines, 10 modules.

%-----------------------------------------------------------------------------
\section{What the Ceiling Means: Constructive Schools}
\label{sec:p43}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18665418}{10.5281/zenodo.18665418}

\bigskip
\noindent\textbf{Abstract.}
The BISH+LPO ceiling unifies three constructive schools (Bishop, Brouwer, Markov); their disagreement localizes to Markov's Principle. LPO strictly implies MP. Physical actualisation requires Cournot's Principle + MP. Three-step chain: BISH computation $\to$ Cournot exclusion $\to$ MP witness extraction. $\sim$770 lines, 12 files.

%-----------------------------------------------------------------------------
\section{The Measurement Problem Dissolved: Quantum Interpretations}
\label{sec:p44}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18671162}{10.5281/zenodo.18671162}

\bigskip
\noindent\textbf{Abstract.}
Three interpretations stratified: Copenhagen (WLPO minimal, LPO strong), Many-Worlds (DC), Bohmian mechanics (LPO). Since WLPO $<$ LPO and DC is incomparable with both, the interpretations sit at provably distinct positions. The measurement problem is not one problem but three logically distinct commitments.


%=============================================================================
% PART V: ARITHMETIC GEOMETRY AND MOTIVES (Papers 45--60)
%=============================================================================
\newpage
\part*{Part V: Arithmetic Geometry and Motives (Papers 45--60)}
\addcontentsline{toc}{part}{Part V: Arithmetic Geometry and Motives (Papers 45--60)}

%-----------------------------------------------------------------------------
\section{Weight-Monodromy Conjecture and LPO}
\label{sec:p45}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18676170}{10.5281/zenodo.18676170}

\bigskip
\noindent\textbf{Abstract.}
Four theorems: (C1)~Hodge polarization forces degeneration in BISH; (C2)~abstract degeneration decidability $\leftrightarrow$ LPO for the coefficient field; (C3)~positive-definite Hermitian forms blocked over $p$-adic fields (dimension $\ge3$); (C4)~geometric perverse sheaves degenerate in BISH. The gap between C2 and C4 is the \emph{de-omniscientizing descent}: geometric origin replaces LPO with decidable equality.

%-----------------------------------------------------------------------------
\section{Tate Conjecture and LPO}
\label{sec:p46}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18682285}{10.5281/zenodo.18682285}

\bigskip
\noindent\textbf{Abstract.}
Four theorems: (T1)~Galois-invariance decidability $\leftrightarrow$ LPO($\mathbb{Q}_\ell$); (T2)~numerical equivalence decidable in BISH; (T3)~Poincar\'e pairing cannot be anisotropic over $\mathbb{Q}_\ell$ in dimension $\ge5$; (T4)~Standard Conjecture D is precisely the axiom that de-omniscientizes motivic morphism spaces. 21 custom axioms.

%-----------------------------------------------------------------------------
\section{Fontaine--Mazur Conjecture and LPO}
\label{sec:p47}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18682788}{10.5281/zenodo.18682788}

\bigskip
\noindent\textbf{Abstract.}
Five theorems: (FM1)~unramifiedness $\leftrightarrow$ LPO($\mathbb{Q}_p$); (FM2)~de Rham condition $\leftrightarrow$ LPO($\mathbb{Q}_p$); (FM3)~Faltings comparison descends to $\mathbb{Q}$, making BISH-decidable; (FM4)~geometric Frobenius traces decidable in BISH; (FM5)~$u(\mathbb{Q}_p)=4$ blocks positive-definite forms in dimension $\ge3$. De-omniscientizing descent via Faltings and the Weil conjectures.

%-----------------------------------------------------------------------------
\section{Birch and Swinnerton-Dyer Conjecture and LPO}
\label{sec:p48}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18683400}{10.5281/zenodo.18683400}

\bigskip
\noindent\textbf{Abstract.}
Four theorems: (B1)~$L(E,1)=0$ decision $\leftrightarrow$ LPO($\mathbb{R}$); (B2)~N\'eron--Tate height gives positive-definite Archimedean polarization; (B3)~regulator $\mathrm{Reg}>0$ in BISH; (B4)~$p$-adic height cannot be positive-definite for rank $\ge5$. BSD is the first conjecture where Archimedean polarization is available---Papers~45--47 proved it blocked at every finite prime. 9 custom axioms.

%-----------------------------------------------------------------------------
\section{Hodge Conjecture and Constructive Omniscience}
\label{sec:p49}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18683802}{10.5281/zenodo.18683802}

\bigskip
\noindent\textbf{Abstract.}
Five theorems: (H1)~Hodge type $(r,r)$ decidability $\leftrightarrow$ LPO($\mathbb{C}$); (H2)~rationality testing requires LPO+MP; (H3)~Hodge--Riemann polarization available ($u(\mathbb{R})=1$) but blind to the rational lattice; (H4)~algebraic cycle verification BISH via integer intersection numbers; (H5)~the Hodge Conjecture reduces LPO to BISH+MP. Unique phenomenon: polarization available but insufficient.

%-----------------------------------------------------------------------------
\section{Three Axioms for the Motive}
\label{sec:p50}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18705837}{10.5281/zenodo.18705837}

\bigskip
\noindent\textbf{Abstract.}
Five calibrations converge on a three-axiom specification of Grothendieck's category of numerical motives: the \emph{Decidable Polarized Tannakian} (DPT) category. Three axioms: (1)~decidable morphism equality (Standard Conjecture D), (2)~algebraic spectrum, (3)~Archimedean polarization. Theorem~A: Weil RH reduces to cancellation. Theorem~B: Honda--Tate inhabitant over $\mathbb{F}_q$. Theorem~C: D is the decidability axiom. Theorem~D: conjectures are $\Pi_2^0$ mandates with the motive as $-1$ shift operator. Theorem~E: CM elliptic motives unconditionally BISH-decidable. 8 files, 46 custom axioms.

%-----------------------------------------------------------------------------
\section{Constructive Archimedean Rescue in BSD}
\label{sec:p51}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18732168}{10.5281/zenodo.18732168}

\bigskip
\noindent\textbf{Abstract.}
The positive-definite Archimedean metric converts the rank-1 BSD generator search from MP (unbounded) to BISH (bounded exhaustive search) in an explicit Finset of size $O(B^2)$ where $B=\lceil\exp(2\hat{h}(y_K)+2\mu(E))\rceil$. The $p$-adic analogue fails: without positive-definiteness ($u=4$), the canonical height can vanish on non-torsion points. Zero sorries, zero custom axiom declarations.

%-----------------------------------------------------------------------------
\section{Decidability Transfer via Specialization: Standard Conjecture D for Abelian Threefolds}
\label{sec:p52}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18732559}{10.5281/zenodo.18732559}

\bigskip
\noindent\textbf{Abstract.}
Standard Conjecture D for abelian varieties of dimension $g\le3$ follows from the Tate conjecture for divisors (unconditional by Tate 1966) via decidability transfer from characteristic $p$ to characteristic~0. Three components: smooth proper base change, unconditionally definite Lefschetz ring, Fourier--Mukai stability of the liftable subspace. Fails sharply at dimension~4 (exotic Tate classes appear outside the Lefschetz ring). Zero custom axioms.

%-----------------------------------------------------------------------------
\section{The CM Decidability Oracle}
\label{sec:p53}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18713089}{10.5281/zenodo.18713089}

\bigskip
\noindent\textbf{Abstract.}
A verified decision procedure for numerical equivalence on products of the 13 CM elliptic curves over $\mathbb{Q}$ with class number~1. The decider returns a Boolean with correctness guaranteed by five principled axioms. All 13 curves pass all three DPT axioms computationally. We extend to the dimension-4 boundary: for Milne's CM abelian fourfold, $\deg(w\cdot w)=7>0$, confirming Hodge--Riemann for the exotic class. 15 files, $\sim$1500 lines.

%-----------------------------------------------------------------------------
\section{The Bloch--Kato Calibration: Out-of-Sample DPT Test}
\label{sec:p54}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18732964}{10.5281/zenodo.18732964}

\bigskip
\noindent\textbf{Abstract.}
First out-of-sample test of the DPT framework. Key results: (A)~LPO isolation for zero-testing; (B)~Axiom~2 realized by Deligne Weil~I; (C)~Axiom~3 partial (Hodge--Riemann unconditional, Beilinson conditional); (D)~Axiom~1 fails for mixed motives ($\mathrm{Ext}^1$ decidability unavailable); (E)~$p$-adic Tamagawa obstruction ($u(\mathbb{Q}_p)=4$); (F)~descent diagram with explicit fracture points. 8 files, 1,139 lines, 7 principled axioms.

%-----------------------------------------------------------------------------
\section{K3 Surfaces, Kuga--Satake Construction, and the DPT Framework}
\label{sec:p55}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18733731}{10.5281/zenodo.18733731}

\bigskip
\noindent\textbf{Abstract.}
Second out-of-sample test. Full DPT success for K3 surfaces. Axiom~1 transfers via Andr\'e (1996). Axiom~2 independent via Deligne Weil~I. Axiom~3 via the Kuga--Satake Clifford algebra and Rosati involution. Supersingular bypass at $\rho=22$. No Picard boundary: decidability does not degrade. Codimension principle confirmed. Calabi--Yau threefold correction: weight-3 Hodge--Riemann restores positive-definiteness. 8 files, 1,131 lines, 9 principled axioms.

%-----------------------------------------------------------------------------
\section{Exotic Weil Class Self-Intersection on CM Abelian Fourfolds}
\label{sec:p56}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18734021}{10.5281/zenodo.18734021}

\bigskip
\noindent\textbf{Abstract.}
We compute $\deg(w_0\cdot w_0)=\sqrt{\mathrm{disc}(F)}$ for exotic Weil classes on three CM abelian fourfolds: (1)~$K=\mathbb{Q}(\sqrt{-3})$, $\mathrm{disc}(F)=49$, degree~7; (2)~$K=\mathbb{Q}(i)$, $\mathrm{disc}(F)=81$, degree~9; (3)~$K=\mathbb{Q}(\sqrt{-7})$, $\mathrm{disc}(F)=169$, degree~13. All degrees positive (Hodge--Riemann), all exotic classes algebraic (Schoen). Determinant equation verified by \texttt{native\_decide}. $\sim$1,666 lines, 10 principled axioms, zero sorry.

%-----------------------------------------------------------------------------
\section{Exotic Weil Self-Intersection Across All Nine Heegner Fields}
\label{sec:p57}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18735172}{10.5281/zenodo.18735172}

\bigskip
\noindent\textbf{Abstract.}
Extension of Paper~56 to all nine class-number-1 imaginary quadratic fields (Baker--Heegner--Stark): $d\in\{1,2,3,7,11,19,43,67,163\}$. For each $K=\mathbb{Q}(\sqrt{-d})$, we compute the trace matrix, verify the discriminant via \texttt{native\_decide}, and derive the Weil class degree via the cyclic conductor theorem. Complete 9-row pattern table assembled. $\sim$1,257 lines, 1 principled axiom.

%-----------------------------------------------------------------------------
\section{Class Number Correction for Exotic Weil Classes}
\label{sec:p58}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18734718}{10.5281/zenodo.18734718}

\bigskip
\noindent\textbf{Abstract.}
Extension to $h_K>1$. The corrected formula is $h\cdot\mathrm{Nm}(\mathfrak{a})=f$, where $\mathfrak{a}$ is the Steinitz class. The topological volume $\det(G)=f^2|\Delta_K|$ is an absolute invariant; the class group redistributes it between metric and lattice density. Verified for $K=\mathbb{Q}(\sqrt{-5})$ ($h_K=2$) at all nine conductors. Norm obstruction is decidable in BISH. 6 files, 803 lines, zero sorry.

%-----------------------------------------------------------------------------
\section{De Rham Decidability and DPT Completeness}
\label{sec:p59}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18735931}{10.5281/zenodo.18735931}

\bigskip
\noindent\textbf{Abstract.}
We formalize the $p$-adic precision bound $N_M=v_p(\det(1-\varphi))=v_p(1-a_p+p)=v_p(\#E(\mathbb{F}_p))$ for elliptic curves with good reduction. The Hasse bound $a_p^2\le4p$ implies $\#E(\mathbb{F}_p)\ge1$. We verify $N_M$ for 24 entries across 4 curves; four anomalous ($N_M\ge1$), 20 generic ($N_M=0$). ``Axiom~5'' (de Rham decidability) is a \emph{theorem}: de Rham $\Rightarrow$ potentially semistable $\Rightarrow$ weakly admissible $\Rightarrow$ $N_M$ computable in BISH. All computation is pure integer arithmetic.

The DPT framework for numerical equivalence on pure motives is \emph{complete}: Axioms~1--3 plus automatic de Rham decidability suffice. No mixed motive axiom is needed. We initiate the extended framework for rational equivalence by proving an analytic rank stratification theorem: the logical complexity of computing $\mathrm{Ext}^1(\mathbb{Q}(0),M)$ is determined by $r=\mathrm{ord}_{s=s_0}L(M,s)$. For $r=0$, exact verification is BISH. For $r=1$, Bloch--Kato bounds the height and Northcott bounds the search, yielding BISH. For $r\ge2$, Minkowski's geometry of numbers forces Markov's Principle.

6 files, 762 lines, zero sorry, zero custom axioms. (Incorporates former Paper~60.)

%-----------------------------------------------------------------------------
\section{[Retired --- Merged into Paper 59]}
\label{sec:p60}
\textbf{Status:} Retired (2026-02-22).\\
\textbf{Former title:} Analytic Rank Stratification of Mixed Motives: Completing the DPT Framework\\
\textbf{Former DOI:} \href{https://doi.org/10.5281/zenodo.18728923}{10.5281/zenodo.18728923} (reserved, not published)

\bigskip
\noindent\textit{Content folded into Paper~59, \S\S9.4--9.6.}

%=============================================================================
% PART VI: MIXED MOTIVES, SELF-INTERSECTION, AND SYNTHESIS (Papers 61--69)
%=============================================================================
\newpage
\part*{Part VI: Mixed Motives, Self-Intersection, and Synthesis (Papers 61--70)}
\addcontentsline{toc}{part}{Part VI: Mixed Motives, Self-Intersection, and Synthesis}

%-----------------------------------------------------------------------------
\section{Lang's Conjecture as the MP$\to$BISH Gate}
\label{sec:p61}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18736959}{10.5281/zenodo.18736959}

\bigskip
\noindent\textbf{Abstract.}
We prove that an Effective Lang Height Lower Bound is the precise gate converting rank~$\ge2$ motives from MP to BISH, via inversion of Minkowski's Second Theorem. Five theorems: (A)~Lang + Minkowski inversion reduces rank-$r$ search to BISH; (B)~without Lang, Minkowski inversion fails in dimension~$\ge2$, requiring MP; (C)~Uniform Lang implies the $L$-function is a universal decidability certificate; (D)~explicit verification for $X_0(389)$ (rank~2). Complete Lean~4 formalization: 9 files, $\sim$900 lines, zero sorry.

%-----------------------------------------------------------------------------
\section{[Retired --- Merged into Paper 63]}
\label{sec:p62}
\textbf{Status:} Retired.\\
\textbf{Former title:} The Hodge Level Boundary\\
\textbf{Former DOI:} \href{https://doi.org/10.5281/zenodo.18736965}{10.5281/zenodo.18736965} (reserved, not published)

\bigskip
\noindent\textit{Content folded into Paper~63.}

%-----------------------------------------------------------------------------
\section{The Intermediate Jacobian Obstruction}
\label{sec:p63}
\textbf{DOI:} 10.5281/zenodo.18736965

\bigskip
\noindent\textbf{Abstract.}
Algebraicity of the intermediate Jacobian $J^p(X)$ determines cycle-search decidability. When $h^{n,0}=0$ (Hodge level $\ell\le1$): $J^p$ is an abelian variety with N\'eron--Tate height satisfying Northcott, and cycle search requires exactly MP. When $h^{n,0}\ge1$ (Hodge level $\ell\ge2$): $J^p$ is a non-algebraic complex torus with no height function, and cycle search requires LPO. Four-way equivalence: algebraic $\iff$ low Hodge $\iff$ Northcott $\iff$ MP. Five main theorems verified on the cubic threefold and quintic Calabi--Yau. Lean~4: 8 files, 1,136 lines, zero sorry. (Incorporates former Paper~62.)

%-----------------------------------------------------------------------------
\section{Uniform $p$-Adic Decidability for Elliptic Curves}
\label{sec:p64}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18737090}{10.5281/zenodo.18737090}

\bigskip
\noindent\textbf{Abstract.}
The crystalline precision bound $N_M=v_p(\#E(\mathbb{F}_p))$ for elliptic curves is uniformly bounded: $N_M\le2$ (with $N_M=2$ only at $p=2$, and $N_M\le1$ for all $p\ge3$). Verified computationally across 1,812 elliptic curves and 23,454 $(E,p)$ pairs. The $p$-adic side is uniformly BISH-decidable with no omniscience principles required, contrasting sharply with the Archimedean side's rank stratification. Computational paper, no Lean formalization.

%-----------------------------------------------------------------------------
\section{Self-Intersection Patterns Beyond Cyclic Cubics}
\label{sec:p65}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18743151}{10.5281/zenodo.18743151}

\bigskip
\noindent\textbf{Abstract.}
Verifies the Steinitz--conductor identity $h\cdot\mathrm{Nm}(\mathfrak{a})=f$ across all 1,220 pairs $(K,F)$ where $K=\mathbb{Q}(\sqrt{-d})$ with $d\le200$ and $F$ is cyclic cubic with conductor $f\le200$. Result: 738 pairs have free lattice ($h=f$), 482 pairs have Steinitz twist ($\mathrm{Nm}(\mathfrak{a})>1$), zero exceptions. For non-cyclic $S_3$ cubics: the scalar identity $h^2=\mathrm{disc}(F)$ never holds. Computational paper.

%-----------------------------------------------------------------------------
\section{Form-Class Resolution for Non-Cyclic Totally Real Cubics}
\label{sec:p66}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18745722}{10.5281/zenodo.18745722}

\bigskip
\noindent\textbf{Abstract.}
The trace-zero sublattice $\Lambda_0=\{x\in\mathcal{O}_F:\mathrm{Tr}_{F/\mathbb{Q}}(x)=0\}$ of a totally real cubic carries a positive-definite binary quadratic form of discriminant $-12\cdot\mathrm{disc}(F)$. For cyclic cubics, the form is $2f\cdot(1,1,1)$. For non-cyclic ($S_3$) cubics, the form is generically non-scalar. Computed for 51 non-cyclic totally real cubics with $\mathrm{disc}(F)\le2000$; the $\mathrm{GL}_2(\mathbb{Z})$-equivalence class is injective on discriminant. Computational paper.

%-----------------------------------------------------------------------------
\section{The Motive Is a Decidability Certificate (Monograph)}
\label{sec:p67}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18746343}{10.5281/zenodo.18746343}

\bigskip
\noindent\textbf{Abstract.}
Synthesis monograph of Papers~45--66 in the arithmetic geometry phase of the Constructive Reverse Mathematics programme. Working over BISH, the paper calibrates major conjectures (Hodge, Tate, BSD, Fontaine--Mazur, Weight-Monodromy) against the logical hierarchy BISH $\subset$ BISH+MP $\subset$ BISH+WLPO $\subset$ BISH+LPO $\subset$ CLASS. Three main outputs: (i)~a decidability classification showing three invariants (analytic rank~$r$, Hodge level~$\ell$, effective Lang constant~$c$) determine the logical strength of cycle-search for any motive; (ii)~the DPT (Decidable Polarized Tannakian) framework for computing motives; (iii)~the arithmetic identity $h\cdot\mathrm{Nm}(\mathfrak{a})=f$ with Steinitz generalization to non-cyclic $S_3$ cubics. Contains 53 Lean~4 formalizations (86,898+ lines total, 22 papers with zero \texttt{sorry}).

%-----------------------------------------------------------------------------
\section{Fermat's Last Theorem Is BISH}
\label{sec:p68}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18901011}{10.5281/zenodo.18901011}

\bigskip
\noindent\textbf{Abstract.}
Stage-by-stage constructive reverse mathematics audit of Wiles's proof of modularity of semistable elliptic curves and Fermat's Last Theorem, decomposing into five stages: (1)~residual modularity via Langlands--Tunnell, (2)~deformation ring, (3)~Hecke algebra, (4)~numerical criterion, (5)~Taylor--Wiles patching. Principal finding: asymmetry---Stages 2--5 are fully constructive (BISH), while Stage~1 requires WLPO. The Taylor--Wiles engine contributes zero logical cost. Two post-Wiles developments drive Stage~5 classification: Brochard's proof of de Smit's conjecture (2017, eliminates infinite inverse limit) and effective Chebotarev bounds (LMO 1979, Ahn--Kwon 2019, make prime search bounded). Stage~1 can be bypassed: the $p=2$ dihedral base case (Kisin 2009, Khare--Wintenberger 2009) replaces Wiles's $p=3$ octahedral case with Hecke theta series and Poisson summation (BISH). Overall: Fermat's Last Theorem is BISH. Lean~4: 493 lines across 3 files, zero sorry.

%-----------------------------------------------------------------------------
\section{The Logical Cost of the Archimedean Place: Function Field Langlands Is BISH}
\label{sec:p69}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18749757}{10.5281/zenodo.18749757}

\bigskip
\noindent\textbf{Abstract.}
CRM audit of the function field Langlands correspondence. Both Laurent Lafforgue's proof for $\GL_n$ (Inventiones, 2002) and Vincent Lafforgue's proof for general reductive groups (JAMS, 2018) are unconditionally BISH: every component operates within Bishop's constructive mathematics, with no omniscience principle required. The principal finding is the comparison with number fields. Paper~68 established that Wiles's proof route costs $\mathrm{BISH}+\mathrm{WLPO}$, with the WLPO entering solely through the Arthur--Selberg trace formula at the Archimedean place. The function field $\mathbb{F}_q(C)$ has no Archimedean place. Every component which costs WLPO over number fields has an algebraic counterpart over function fields that costs nothing: the Grothendieck--Lefschetz trace formula replaces the Arthur--Selberg trace formula; rational Plancherel measures replace transcendental ones; finite-dimensional spaces of cusp forms replace infinite-dimensional $L^2$ spaces. Structural discovery: the boundary between BISH and WLPO is not discrete-vs-continuous spectrum, but algebraic-vs-transcendental spectral parameters. The logical cost of the Langlands program is the logical cost of~$\mathbb{R}$. Lean~4: 236 lines, zero sorry, no \texttt{Classical.choice}.

%-----------------------------------------------------------------------------
\section{The Archimedean Principle: Why Physics and Number Theory Share a Logical Architecture}
\label{sec:p70}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18901192}{10.5281/zenodo.18901192}

\bigskip
\noindent\textbf{Abstract.}
Capstone paper identifying a single structural principle underlying 70~papers of Constructive Reverse Mathematics: the real numbers are the sole source of logical difficulty in mathematical physics and arithmetic geometry. The mechanism is $u(\mathbb{R})=\infty$---the real numbers are the only completion of~$\mathbb{Q}$ where positive-definite forms exist in every dimension---and three fields independently exploit it via the same architecture (Hilbert space inner product, Rosati involution, Petersson inner product). Three theorems: (A)~the Archimedean Principle---the CRM level of every domain is determined by whether it has an Archimedean place; (B)~the MP Gap---physics and arithmetic descend differently (projection vs.\ search), producing a strict separation $\mathrm{BISH}<\mathrm{BISH}+\mathrm{MP}$; (C)~Automorphic CRM Incompleteness---an integer witness proving the automorphic axioms alone cannot recover the Ramanujan bound. Paper~68 showed FLT is BISH; Paper~69 showed function field Langlands is BISH; this paper identifies what makes anything expensive: the Archimedean place and $u(\mathbb{R})=\infty$. Lean~4: 5 files, 420 lines, zero sorry, zero custom axioms.

%-----------------------------------------------------------------------------
\section{The Archimedean Principle in Cryptography and Numerical Computation}
\label{sec:p71}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18752015}{10.5281/zenodo.18752015}

\bigskip
\noindent\textbf{Abstract.}
Four engineering consequences of the Archimedean Principle (Paper~70). Theorem~A (Archimedean Security): lattice-based cryptography (SVP, LWE, Ring-LWE) is not amenable to Shor-type quantum attacks because solution targets are \emph{metric} (Archimedean norm bounds), not \emph{algebraic} (group-theoretic relations); metric targets delocalize in expectation under spectral projection by Fourier energy conservation; the function field control confirms SVP over~$\mathbb{F}_q[t]$ is BISH (polynomial-time). Theorem~B (SVP Phase Transition): exponential approximation is projection-descent (LLL, BISH); polynomial approximation is search-descent (BKZ, BISH+MP). Theorem~C (Conjugacy Design Principle): maximize Fourier conjugacy between algebraic operations and metric security assumptions; a conjugacy index quantifies structural security: Kyber $>$ NTRU $>$ RSA. Theorem~D (Eigendecomposition Integrality): any nontrivial eigendecomposition of a positive-definite integer matrix introduces irreducible transcendental contamination. All four applications follow from one mechanism: projection descent eliminates MP; search descent preserves it; the Archimedean metric is canonically conjugate to algebraic spectral decomposition. Lean~4: 5 files, zero sorry, zero custom axioms; sum-of-integer-squares lemma is a genuine Mathlib proof.


%-----------------------------------------------------------------------------
\section{The DPT Characterization Theorem: Archimedean Polarization Is Necessary for Cycle-Search}
\label{sec:p72}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18765393}{10.5281/zenodo.18765393}

\bigskip
\noindent\textbf{Abstract.}
We prove that the three DPT axioms (Paper~50) are the minimal axiom set for $\mathrm{BISH}$-decidable motivic cycle-search.  The new result: positive-definiteness of the height pairing (Axiom~3) is not merely sufficient but \emph{necessary}.  Without Northcott's finiteness guarantee, the search for Mordell--Weil generators is unbounded and the $L$-function zero-test encodes~$\mathrm{LPO}$ (Paper~48).  Combined with the forward direction (Papers~45--51), this gives a biconditional: Axiom~3 $\Leftrightarrow$ $\mathrm{BISH}$ cycle-search.  The Archimedean Principle (Paper~70) is thereby sharpened from a forward implication to an equivalence.  The $\mathbb{Z}$-density argument provides a second, independent obstruction: $\mathbb{Z}$ is discrete in~$\mathbb{R}$ but dense in~$\mathbb{Q}_p$, so bounded-height sets are infinite $p$-adically regardless of the form's algebraic signature.  DPT and Fargues--Scholze are thereby identified as a logically forced partition: discrete sector ($\mathbb{R}$, BISH, cycle-search) vs.\ continuous sector ($\mathbb{Q}_p$, CLASS, categorical).  Lean~4 formalization: ${\sim}350$ lines, zero \texttt{sorry}, no \texttt{Classical.choice}.

%-----------------------------------------------------------------------------
\section{Standard Conjecture D Is Necessary for Morphism Decidability}
\label{sec:p73}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18765700}{10.5281/zenodo.18765700}

\bigskip
\noindent\textbf{Abstract.}
We prove that Standard Conjecture~D is necessary for $\mathrm{BISH}$ morphism decidability in a realization-compatible motivic category: Conjecture~D $\Leftrightarrow$ $\mathrm{BISH}$ morphism decidability. Without Conjecture~D, the radical of the intersection pairing is non-detachable; morphism equality requires $\mathbb{Q}_\ell$ zero-testing $\to$ $\mathrm{LPO}$. Jannsen's theorem (1992) confirms the trade-off: numerical motives are $\mathrm{BISH}$ but lack faithful realization. The DPT axiom system's first axiom is thereby characterized as the unique condition for its domain (canonical, not merely minimal). Lean~4: ${\sim}250$ lines, zero \texttt{sorry}, 4~axioms.

%-----------------------------------------------------------------------------
\section{Algebraic Spectrum Is Necessary for Eigenvalue Decidability}
\label{sec:p74}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18773827}{10.5281/zenodo.18773827}

\bigskip
\noindent\textbf{Abstract.}
We prove that algebraic spectrum is necessary for $\mathrm{BISH}$ eigenvalue decidability: algebraic spectrum $\Leftrightarrow$ $\mathrm{BISH}$ eigenvalue decidability. Without geometric origin, the full analytic Langlands spectrum (Maass forms, unramified characters) has continuous spectral parameters; eigenvalue comparison costs $\mathrm{WLPO}$ (equality test, not existential search). The naive ``transcendental eigenvalues'' framing has three fatal flaws (Weil~II, linear algebra vacuum, $\ell$-adic $\mathrm{CLASS}$ trap); the correct domain of failure is the geometric-analytic boundary. The DPT axiom system is now fully characterized: each axiom is the unique condition for its domain (canonical, not merely minimal). Lean~4: ${\sim}200$ lines, zero \texttt{sorry}, zero \texttt{propext}, 4~axioms.

%-----------------------------------------------------------------------------
\section{Conservation Test: Genestier--Lafforgue LLC Calibration}
\label{sec:p75}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18773831}{10.5281/zenodo.18773831}

\bigskip
\noindent\textbf{Abstract.}
DPT passes its first external validation test. The Genestier--Lafforgue semisimple local Langlands correspondence for arbitrary reductive~$G$ decomposes into three independently calibrated layers: algebraic ($\mathrm{BISH}$, solidification via Mittag-Leffler), homological ($\mathrm{CLASS}$/Zorn, K-injective resolutions), geometric ($\mathrm{CLASS}$/BPI, v-topology). Statement costs $\mathrm{BISH}+\mathrm{WLPO}$; proof costs $\mathrm{CLASS}$; conservation gap of two levels. DPT Axiom~2 correctly predicts the statement cost. The conservation conjecture --- whether the $\mathrm{CLASS}$ scaffolding is eliminable --- remains open. Lean~4: ${\sim}180$ lines, zero \texttt{sorry}, zero \texttt{propext}, zero \texttt{Classical.choice}.

%=============================================================================
\newpage
\part*{Part VIII: Proof Automation (Paper 76)}
\addcontentsline{toc}{part}{Part VIII: Proof Automation}

%-----------------------------------------------------------------------------
\section{CRMLint: An Automated Logical Cost Analyzer for Proof Assistants}
\label{sec:p76}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18777597}{10.5281/zenodo.18777597}

\bigskip
\noindent\textbf{Abstract.}
We present \texttt{CRMLint}, a Lean~4 metaprogram that computes the CRM logical cost of any declaration in the Mathlib library. The tool traces transitive dependencies to classical axioms (\texttt{Classical.choice}, \texttt{Classical.em}, \texttt{propext}, \texttt{Quot.sound}), classifies each entry point by its CRM pattern (infrastructure artifact vs.\ essential classical content), and computes the CRM level as the join of essential costs over the hierarchy $\mathrm{BISH} \subset \mathrm{BISH{+}MP} \subset \mathrm{LLPO} \subset \mathrm{WLPO} \subset \mathrm{LPO} \subset \mathrm{CLASS}$. A critical v0.1 classification failure --- \texttt{Real.instField} read as $\mathrm{BISH}$ instead of $\mathrm{WLPO}$ --- is fixed by a type-aware classifier that inspects the type being decided. Calibration against 14~Lean~4 bundles from the CRM program yields correct classifications on all test cases. Namespace scans reveal conservation gaps across Mathlib: 70\% of \texttt{Nat.Prime} and 60\% of \texttt{Int.ModEq} are classified $\mathrm{CLASS}$ despite having $\mathrm{BISH}$ statements. Three such gaps are closed with constructive proofs. The conservation gap $\Delta(T) = c_p - c_s$ defines a heuristic logical cost function for AI theorem provers, converting constructive de-omniscientisation from a global rewrite into a localized DAG surgery problem. Lean~4: 1,069 lines (940 core + 129 tests), zero \texttt{sorry}.

%=============================================================================
\newpage
\part*{Part IX: CRMLint Squeeze Applications (Papers 77--78)}
\addcontentsline{toc}{part}{Part IX: CRMLint Squeeze Applications}

%-----------------------------------------------------------------------------
\section{Explicit Hodge Decompositions for $E^4$ via Automated De-Omniscientisation}
\label{sec:p77}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18777596}{10.5281/zenodo.18777596}

\bigskip
\noindent\textbf{Abstract.}
We introduce the CRMLint Squeeze protocol---a systematic method for extracting constructive content from classical existence theorems---and demonstrate it on the Hodge conjecture for~$E^4$ (a product of four copies of an elliptic curve with complex multiplication). The Classical Boundary Node is the Lefschetz~$(1,1)$ theorem: every rational $(1,1)$-class is algebraic. This is~$\mathsf{CLASS}$ (topological, uses Hodge theory over~$\mathbb{C}$). The constructive replacement: for~$E^4$, every Hodge $(2,2)$-class decomposes as an explicit $\mathbb{Q}$-linear combination of products of divisor classes, computed by a deterministic Python CAS and verified by \texttt{native\_decide} in Lean~4. All~36 Hodge $(2,2)$-classes admit explicit decompositions; the exotic Weil subspace is~$\{0\}$ (the Anderson obstruction does not apply to non-simple products). Total Lean: 798 lines (auto-generated), 0~sorry. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$.

%-----------------------------------------------------------------------------
\section{Explicit Local Langlands for Wildly Ramified Groups: A CRMLint Squeeze}
\label{sec:p78}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18789008}{10.5281/zenodo.18789008}

\bigskip
\noindent\textbf{Abstract.}
We apply the CRMLint Squeeze protocol to the Local Langlands Correspondence for wildly ramified supercuspidal representations of~$\mathrm{GL}_2(\mathbb{Q}_2)$. The Classical Boundary Node is the Harris--Taylor existence theorem: for every irreducible smooth representation~$\pi$, a Weil--Deligne parameter $\sigma(\pi)$ exists. This is~$\mathsf{CLASS}$ (Shimura varieties, \'etale cohomology, Lefschetz fixed-point). The constructive replacement: for supercuspidal representations built from Bushnell--Kutzko simple types, the Galois parameter is determined by finitely many character values computable from the type data. We carry out this program for~$\mathrm{GL}_2(\mathbb{Q}_2)$ at minimal wild conductor, using the totally ramified extension $E = \mathbb{Q}_2(i)$ with uniformizer $\pi_E = 1+i$ and a conductor-2 character~$\theta$. The character-trace matching identity is verified by \texttt{decide} in Lean~4 for~16 regular elliptic test elements; central elements are handled via a separate central character matching. The Frobenius eigenvalues $\alpha_1 = 1$, $\alpha_2 = -1$ are determined entirely by the Bushnell--Kutzko type data. Harris--Taylor's global existence theorem is present in the formalization as a~$\mathsf{CLASS}$ axiom but is never invoked. Total Lean: 354 lines, 0~sorry, 0~axioms on constructive path. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$.

%-----------------------------------------------------------------------------
\section{Standard Conjecture D for CM Abelian Fourfolds}
\label{sec:p79}
\textbf{DOI:} (Unassigned --- number skipped)

\bigskip
\noindent\textbf{Abstract.}
(Planned.) Homological equivalence = numerical equivalence for simple CM abelian fourfolds, where exotic Weil classes genuinely exist (Anderson obstruction applies). The CRMLint Squeeze targets the positive-definiteness failure on the exotic subspace. First target where the MCTS search mode may be necessary.

%-----------------------------------------------------------------------------
\section{Algebraic Gauss--Manin Connections over Function Fields: A CRMLint Squeeze}
\label{sec:p80}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18791985}{10.5281/zenodo.18791985}

\bigskip
\noindent\textbf{Abstract.}
We apply the CRMLint Squeeze protocol to the Gauss--Manin connection of the Legendre family of elliptic curves $y^2 = x(x-1)(x-t)$ over the function field~$\mathbb{Q}(t)$. The Classical Boundary Node is Ehresmann's fibration theorem: for smooth proper morphisms of complex manifolds, the higher direct images form a local system with regular singular flat connection. This is~$\mathsf{CLASS}$ (continuous topology, analytic continuation, Hodge theory). The constructive replacement: Griffiths pole reduction in algebraic de Rham cohomology over~$\mathbb{Q}[x,t]$, yielding the Picard--Fuchs operator $L = t(1-t)D^2 + (1-2t)D - \tfrac{1}{4}$, the traceless $2\times 2$ Gauss--Manin connection matrix, and the hypergeometric recurrence $4(n+1)^2 a_{n+1} = (2n+1)^2 a_n$ with closed form $a_n = \binom{2n}{n}^2 / 4^{2n}$. All identities are polynomial in~$t$ and verified by the \texttt{ring} tactic in Lean~4. The key infrastructure upgrade: Papers~77--78 used \texttt{native\_decide} on~$\mathbb{Q}$-arithmetic (0-dimensional); Paper~80 uses \texttt{ring} on~$\mathbb{Q}[x,t]$-algebra (1-dimensional moduli), enabling CRMLint to operate over families of varieties. Total Lean: 477 lines (244 CAS-emitted + 233 architecture), 0~sorry. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$.

%-----------------------------------------------------------------------------
\section{Algebraic Flat Sections and the Theorem of the Fixed Part: A CRMLint Squeeze}
\label{sec:p81}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18801814}{10.5281/zenodo.18801814}

\bigskip
\noindent\textbf{Abstract.}
We apply the CRMLint Squeeze protocol to the Theorem of the Fixed Part for the Legendre family $y^2 = x(x-1)(x-t)$. The Classical Boundary Node is Deligne's monodromy theorem: the local monodromy of a variation of Hodge structure is quasi-unipotent, and the global sections of the associated local system correspond to the monodromy-invariant classes. This is~$\mathsf{CLASS}$ (topological monodromy, fundamental groups, analytic continuation). The constructive replacement: the tensor Gauss--Manin connection $N(t) = M \otimes I_2 + I_2 \otimes M$ on $H^1 \otimes H^1$ is computed by Kronecker sum, and the constant flat sections are determined by polynomial linear algebra over~$\mathbb{Q}(t)$. The unique constant flat section is the symplectic form $W = (0,1,-1,0)$; the other kernel element $K(t) = (1, 2t, 0, t)$ is non-constant. Total Lean: 457 lines, 0~sorry. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$.

%-----------------------------------------------------------------------------
\section{Constructive Differential Galois Theory via Kovacic's Algorithm: A CRMLint Squeeze}
\label{sec:p82}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18801822}{10.5281/zenodo.18801822}

\bigskip
\noindent\textbf{Abstract.}
We apply the CRMLint Squeeze protocol to the differential Galois group of the Legendre Picard--Fuchs equation. The Classical Boundary Node is topological monodromy density: the image of the topological monodromy representation $\pi_1(\mathbb{P}^1 \setminus \{0,1,\infty\}) \to \mathrm{SL}_2(\mathbb{Z})$ is Zariski-dense in $\mathrm{SL}_2$. This is~$\mathsf{CLASS}$ (fundamental groups, path lifting, comparison theorem). The constructive replacement: Kovacic's algorithm determines $G_{\mathrm{gal}} = \mathrm{SL}_2$ by rational function analysis. All three Kovacic cases fail: Case~1 produces non-integer degree bounds, Case~2 and Case~3 produce negative degrees. The conclusion $G_{\mathrm{gal}} = \mathrm{SL}_2$ follows by elimination. Total Lean: 457 lines, 0~sorry. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$.

%-----------------------------------------------------------------------------
\section{Generic Picard Number of the Legendre Surface Family: A Function-Field Synthesis}
\label{sec:p83}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18801826}{10.5281/zenodo.18801826}

\bigskip
\noindent\textbf{Abstract.}
We synthesize the function-field pipeline (Papers~80--82) to compute the generic Picard number of $E_t \times E_t$, where $E_t : y^2 = x(x-1)(x-t)$ is the Legendre elliptic curve. The classical proof requires the Lefschetz~$(1,1)$ theorem (Baire category), the Noether--Lefschetz theorem (Zariski density), and topological monodromy density ($\mathsf{CLASS}$). The constructive proof replaces all of this with K\"unneth rank arithmetic, tensor flat section analysis (Paper~81: constant kernel dimension~$= 1$), Kovacic obstruction (Paper~82: $G_{\mathrm{gal}} = \mathrm{SL}_2$), and $\mathrm{SL}_2$ representation theory: $(V \otimes V)^{\mathrm{SL}_2}$ is 1-dimensional (the symplectic form). Generic Picard rank $= 2 + 1 = 3$. The ``Unbounded Degree Trap'' is resolved: differential Galois theory provides a finite certificate that no exotic algebraic cycles exist. Total Lean: 282 lines, 0~sorry, \emph{zero axioms} in the squeeze theorem. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$. Capstone of the four-paper function-field arc.

%-----------------------------------------------------------------------------
\section{Global Flatness of Weil Classes via Gauss--Manin Block Decomposition: A CRMLint Squeeze}
\label{sec:p84}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18802819}{10.5281/zenodo.18802819}

\bigskip
\noindent\textbf{Abstract.}
We compute the Gauss--Manin connection on the de~Rham cohomology of the genus-4 hyperelliptic family $C_t\colon y^2 = x^9 - tx^5 + x$ with $\mathbb{Q}(i)$-action, using Griffiths pole-order reduction with the coboundary subtraction $A_k = B_k - h_k'$. An order-8 automorphism $\tau(x,y) = (ix,\, e^{\pi i/4}\,y)$ decomposes the $8 \times 8$ connection matrix into four $2 \times 2$ blocks $M_k = \frac{2k+1}{4(t^2-4)} \bigl(\begin{smallmatrix} -t/2 & -1 \\ 1 & t/2 \end{smallmatrix}\bigr)$ for $k = 0,1,2,3$. Each block has $G_{\mathrm{gal}}(W_k) = \mathrm{SL}_2$ (nilpotent residues with non-proportional kernels). Since $\det(\mathrm{SL}_2) = 1$, the induced action on $\bigwedge^4(V_+) \cong \bigwedge^2(W_0) \otimes \bigwedge^2(W_2)$ is trivial: $G_{\mathrm{gal}}(\bigwedge^4 V_+) = \{1\}$. The exotic Weil class is therefore a global flat section---a constant, rational class across the entire family. By Deligne's Theorem of the Fixed Part, it is an absolute Hodge class on every fiber. This is a positive result: the function-field pipeline certifies that the monodromy obstruction to algebraicity is absent. Total Lean: 389 lines, 0~sorry, zero axioms. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$. Seventh CRMLint Squeeze.

%-----------------------------------------------------------------------------
\section{Universal Flatness of Exotic Weil Classes for Genus-4 Hyperelliptic Families: A CRMLint Squeeze}
\label{sec:p85}
\textbf{DOI:} \href{https://doi.org/10.5281/zenodo.18806608}{10.5281/zenodo.18806608}

\bigskip
\noindent\textbf{Abstract.}
For genus-$g$ hyperelliptic families $C_t\colon y^2 = f(x,t)$ with $\mathbb{Q}(i)$-action ($f$ odd in $x$), the $\sigma$-eigenspace $V_+$ is Lagrangian in $H^1_{\mathrm{dR}}(C_t)$: the cup-product pairing vanishes identically on $V_+ \times V_+$. The Lagrangian splitting $H^1_{\mathrm{dR}} = V_+ \oplus V_-$ gives the symplectic constraint $\mathrm{tr}(M_+) = -\mathrm{tr}(M_-)$, but does not force $\mathrm{tr}(M_+) = 0$ individually. We verify computationally that the exotic trace $\tau_+(t) = \mathrm{tr}(\nabla|_{V_+}) = 0$ for the test family $y^2 = x^9 + tx^7 + x$ (genus~4, no order-8 automorphism). The $V_+$~block is a dense $4 \times 4$ matrix with four nonzero diagonal entries whose numerator coefficients sum to zero. Multi-genus testing (genera 2--5, eight families) confirms the same vanishing universally. \textbf{Erratum.} Version~1 incorrectly claimed $V_+$ is symplectic; the eigenvalue argument proves $V_+$ is Lagrangian ($i^2 = -1 \neq 1$). Total Lean: 296 lines, 0~sorry. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$. Eighth CRMLint Squeeze.

%-----------------------------------------------------------------------------
\section{The Hodge Conjecture for Exotic Weil Classes via Kani--Rosen Splitting: A CRMLint Squeeze}
\label{sec:p86}
\textbf{DOI:} 10.5281/zenodo.18809903

\bigskip
\noindent\textbf{Abstract.}
We prove that the exotic Weil class on $J(C_t)$ is algebraic, where $C_t\colon y^2 = x^9 - tx^5 + x$ is the genus-4 hyperelliptic family from Papers~84--85. The proof exploits a hidden reciprocal involution $\mu(x,y) = (1/x,\, y/x^5)$ that, together with the order-4 automorphism~$\sigma$, generates a dihedral group~$D_8$ acting on~$C_t$. By the Kani--Rosen theorem, the Jacobian splits up to isogeny: $J(C_t) \sim J(Q_1) \times J(Q_2)$, where $Q_1$ and $Q_2$ are explicit genus-2 quotient curves. Over the complex numbers, $Q_1 \cong Q_2$, so $J(C_t) \sim A^2$ for an abelian surface $A = J(Q_1)$. For generic~$t$, $\mathrm{End}(A) = \mathbb{Z}$ forces $\mathrm{MT}(A) = \mathrm{GSp}_4$ (Zarhin), and all Hodge classes on products of such surfaces are algebraic by Weyl's invariant theory. The constructive content (BISH) identifies the splitting via polynomial identities verified by \texttt{ring} in Lean~4. The classical content (CLASS) provides the Hodge-theoretic conclusion. Total Lean: 0~sorry. CRM descent: $\mathsf{CLASS} \to \mathsf{BISH}$. Ninth CRMLint Squeeze.

%-----------------------------------------------------------------------------
\section{The Omniscience Cost of the Hodge Condition: A CRM Analysis}
\label{sec:p87}
\textbf{DOI:} 10.5281/zenodo.18809903

\bigskip
\noindent\textbf{Abstract.}
We perform the first Constructive Reverse Mathematics analysis of the Hodge Conjecture. The uniform Hodge test---deciding ``is $\alpha$ of type $(p,p)$?'' for an arbitrary period matrix presented as Cauchy sequences---requires $\mathsf{WLPO}$: it reduces to testing whether a real number (the $(p,p)$-projection norm) equals zero. When algebraic metadata is available (CM type, known Mumford--Tate group), the Shiga--Wolfart theorem gives algebraic period ratios, and the test drops to $\mathsf{BISH}$. We establish the biconditional: uniform $(p,p)$-decidability $= \mathsf{BISH}$ if and only if algebraic metadata is available. This explains structurally why all known results on the Hodge Conjecture (Lefschetz, Lieberman, Deligne, Shioda) require CM or equivalent rigidity: the omniscience cost of the generic Hodge test is $\mathsf{WLPO}$, and algebraic metadata is the unique mechanism for its elimination. Total Lean: 602 lines, 0~sorry. First CRM analysis of a Clay Millennium Problem.

%-----------------------------------------------------------------------------
\section{The Variational Squeeze: CM Anchors and Fermat Domination of Exotic Weil Classes}
\label{sec:p88}
\textbf{DOI:} 10.5281/zenodo.18809905

\bigskip
\noindent\textbf{Abstract.}
For the asymmetric genus-4 family $C_t\colon y^2 = x^9 + tx^7 + x$ from Paper~85, the Jacobian $J(C_t)$ is generically absolutely simple---the Kani--Rosen splitting of Paper~86 does not apply. The obstruction is algebraic: the core polynomial $g(x,t) = x^8 + tx^6 + 1$ is \emph{not} palindromic ($g - g_{\mathrm{rev}} = tx^2(x^4 - 1) \neq 0$), blocking the reciprocal involution that generated~$D_8$ in the split family. We prove the exotic Weil class is algebraic, conditionally on the Variational Hodge Conjecture, via CM specialization at $t = 0$. The curve $C_0\colon y^2 = x^9 + x$ is dominated by the Fermat curve $F_{16}$ through the rational map $\pi\colon (u,v) \mapsto (u^2/v^2,\, u/v^9)$, verified as the cleared-denominator identity $u^2 = u^{18} + u^2 v^{16}$ from the Fermat equation $u^{16} + v^{16} = 1$. Pullback differentials are holomorphic ($a + b = 8 \le 15$). Shioda's theorem (1982) gives algebraicity of all Hodge classes on~$J(C_0)$, and the Variational Hodge Conjecture spreads algebraicity from the CM fiber to generic~$t$. CRMLint bifurcation identified: split families (Paper~86) admit Kani--Rosen $\to$ unconditional; simple families (Paper~88) require CM anchor $+$ VHC $\to$ conditional. Total Lean: 411 lines, 0~sorry. CRM classification: $\mathsf{CLASS}$ (conditional on VHC). Tenth CRMLint Squeeze.

%-----------------------------------------------------------------------------
\section{The Universal Hyperelliptic Squeeze: Absolute Hodge Classes for the Full Weil Locus}
\label{sec:p89}
\textbf{DOI:} 10.5281/zenodo.18809909

\bigskip
\noindent\textbf{Abstract.}
We prove that the exotic Weil class on the Jacobian of the universal genus-4 hyperelliptic Weil family $C_{a,b,c}\colon y^2 = x^9 + ax^7 + bx^5 + cx^3 + x$ is a global flat section of the Gauss--Manin connection over the full 3-dimensional parameter space $\mathbb{Q}(a,b,c)$. The trace of the connection on the eigenvalue-$i$ eigenspace~$V_+$ vanishes identically---$\tau_+(a,b,c) = 0$---in all three parameter directions $\partial/\partial a$, $\partial/\partial b$, $\partial/\partial c$, verified as polynomial identities in $\mathbb{Z}[a,b,c]$ by Lean~4's \texttt{ring} tactic. At the Fermat anchor $(a,b,c) = (0,0,0)$, the curve $C_0\colon y^2 = x^9 + x$ is dominated by the Fermat hypersurface~$F_{16}$, where algebraicity follows from Shioda's 1982 theorem. Conditional on the Variational Hodge Conjecture, the exotic Weil class is algebraic across the entire hyperelliptic Weil locus. The palindromic locus $\{a = c\}$ is precisely characterized as the Kani--Rosen locus where unconditional algebraicity holds (Paper~86). This is the natural endpoint of the Griffiths pole reduction pipeline combined with the Fermat domination strategy. Total Lean: 434 lines, 0~sorry. CRM classification: $\mathsf{CLASS}$ (conditional on VHC). Eleventh CRMLint Squeeze.

%=============================================================================
\newpage

\section{The Squeeze Is a Microscope: Automated Constructivisation from CRMLint to the Hodge Horizon}
\label{sec:p90}
\textbf{DOI:} 10.5281/zenodo.18809911

\bigskip
\noindent\textbf{Abstract.}
This monograph synthesizes the generative phase (Papers~76--89) of the Constructive Reverse Mathematics program. Where the diagnostic phase (Papers~1--75) manually calibrated theorems against the omniscience hierarchy $\mathsf{BISH} \subset \mathsf{LLPO} \subset \mathsf{WLPO} \subset \mathsf{LPO} \subset \mathsf{CLASS}$, the generative phase automates the process through the CRMLint Squeeze protocol: a four-step procedure (Scaffold, X-Ray, Excise, Squeeze) that isolates the constructive core of a classical theorem from its irreducible classical boundary. Across eleven Squeeze applications spanning Hodge theory, the Langlands program, differential Galois theory, and algebraic geometry, the protocol produces a sharp decomposition: everything except ``an algebraic cycle exists'' is $\mathsf{BISH}$-verifiable via Lean~4's \texttt{ring} and \texttt{decide} tactics. The Hodge campaign (Papers~84--89) reveals a bifurcation: on the palindromic locus $\{a = c\}$, Kani--Rosen splitting gives unconditional algebraicity; on the generic complement, the Variational Hodge Conjecture is the irreducible $\mathsf{CLASS}$ boundary. Paper~87 proves the uniform Hodge test is equivalent to $\mathsf{WLPO}$, explaining why all known Hodge results require algebraic metadata. Total Lean across generative phase: ${\sim}5{,}300$ lines, 0~sorry. No new Lean bundle (synthesis paper). CRM classification: survey.

%=============================================================================
\newpage

\section{The Logical Cost of Unconditional Hodge: CRM Audit of the Markman--Floccari--Fu Theorem}
\label{sec:p91}
\textbf{DOI:} 10.5281/zenodo.18809917

\bigskip
\noindent\textbf{Abstract.}
We perform a CRM audit of Markman's February 2025 unconditional proof of the Hodge Conjecture for all abelian fourfolds of Weil type (arXiv:2502.03415). Four theorems: (A)~Markman's proof decomposes into $4$~$\mathsf{BISH}$ $+$ $5$~$\mathsf{CLASS}$ components (Fourier--Mukai, Orlov, Schoen, semi-regularity, secant existence); overall cost $\mathsf{CLASS}$. (B)~Comparison: the CRMLint Squeeze achieves 90\% $\mathsf{BISH}$ ($18/20$) conditional on VHC; Markman achieves 44\% $\mathsf{BISH}$ ($4/9$) unconditional. (C)~A posteriori VHC consistency: Markman's result shows VHC is not refuted for exotic Weil classes on $J(C_{a,b,c})$. (D)~The cycle class map (de Rham--Betti comparison) is irreducibly $\mathsf{CLASS}$; $20/0$ is impossible, $19/1$ is the best achievable ratio. Includes explicit palindromic cycle via Kani--Rosen (extending Paper~86). Total Lean: 668 lines, 0~sorry, 0~\texttt{Classical.choice}.

%-----------------------------------------------------------------------------
\section{Cohomological Flatness for Genera 5--8 via Zariski Grid Specialization: A CRMLint Squeeze}
\label{sec:p92}
\textbf{DOI:} 10.5281/zenodo.18816696

\bigskip
\noindent\textbf{Abstract.}
We extend the Gauss--Manin trace vanishing pipeline from genus~4 (Papers~84--89) to genera~5--8, proving that the exotic Weil class on the Jacobian of the universal odd hyperelliptic family is cohomologically flat in dimensions~10, 12, 14, and~16. For genus~5, the connection trace vanishes as an explicit polynomial identity in $\mathbb{Z}[a_1,a_2,a_3,a_4]$, verified by Lean~4's \texttt{ring} tactic. For genera~6--8, symbolic expansion triggers exponential intermediate expression swell; we bypass this via Zariski Grid Specialization: evaluate the exact rational Griffiths reduction at concrete integer fibers, verify cancellation by \texttt{norm\_num}, and invoke the Schwartz--Zippel lemma for generic vanishing. Discrete grid evaluation is a finite, terminating, decidable algorithm $\to$ $\mathsf{BISH}$. Total Lean: 333 lines, 98~theorems, 0~sorry, 0~\texttt{Classical.choice}. Twelfth CRMLint Squeeze.

%-----------------------------------------------------------------------------
\section{The Structural Vanishing Theorem: Why Exotic Weil Classes Are Universally Flat}
\label{sec:p93}
\textbf{DOI:} 10.5281/zenodo.18816989

\bigskip
\noindent\textbf{Abstract.}
Papers~84--92 verified computationally that the Gauss--Manin trace $\tau_+$ on $\bigwedge^g V_+$ vanishes identically for odd hyperelliptic Weil families in genera $g = 2,\ldots,8$. This paper provides two independent structural proofs. Proof~1 (hypergeometric): the substitution $u = x^2$ reduces $V_+$ to $H^1$ of a fractional local system; Aomoto--Varchenko exponents cancel to zero (discriminant drops out) while the root-origin exponent is frozen at $-1/4$ by Vieta's formula. Proof~2 (topological): Deligne regularity constrains $\tau_+$ to logarithmic poles; unipotent monodromy kills the discriminant term, and unit coefficients freeze the boundary. CRM decomposition: $13$~$\mathsf{BISH}$ $+$ $4$~$\mathsf{CLASS}$. Total Lean: 525 lines, 0~sorry.

%-----------------------------------------------------------------------------
\section{The Griffiths Group of the Mirror Quintic: A CRMLint Squeeze}
\label{sec:p94}
\textbf{DOI:} 10.5281/zenodo.18820969

\bigskip
\noindent\textbf{Abstract.}
We apply the CRMLint Squeeze protocol to the Griffiths group of the mirror quintic, the first CRM analysis of Calabi--Yau middle cohomology. Walcher (2007) computed the inhomogeneous Picard--Fuchs equation for the real Lagrangian $\mathbb{RP}^3 \subset X_5$: the source term $(15/8)\sqrt{z}$ is algebraic, and the expansion $T = \sqrt{z}\sum b_n z^n$ has $b_0 = 30$ (Walcher's disk count). Four theorems: (A)~PF differential algebra is $\mathsf{BISH}$ (13~sub-results). (B)~Walcher equation verified algebraically. (C)~Normal Function Squeeze: detection (source term $\delta \neq 0$) is $\mathsf{BISH}$; existence (Abel--Jacobi, Baire) is irreducibly $\mathsf{CLASS}$. (D)~Voisin decomposition: $11$~$\mathsf{BISH}$ $+$ $3$~$\mathsf{CLASS}$. Total Lean: 566 lines, 0~sorry. Thirteenth CRMLint Squeeze.

%=============================================================================
\newpage

\section{The BSD Squeeze: CRMLint Decomposition of the Gross--Zagier--Kolyvagin Proof}
\label{sec:p95}
\textbf{DOI:} 10.5281/zenodo.18821019

\bigskip
\noindent\textbf{Abstract.}
The Gross--Zagier--Kolyvagin proof of BSD for analytic rank $\leq 1$ is the only proven case among the open Clay Millennium Problems. We apply CRMLint to decompose the proof pipeline using the elliptic curve $\texttt{37a1}$ ($y^2 + y = x^3 - x$, rank~1, $w = -1$) as test case. Four theorems: (A)~Hecke eigenvalue arithmetic (28~sub-results) is strictly $\mathsf{BISH}$ (\texttt{native\_decide}). (B)~GZK Squeeze: $15$~$\mathsf{BISH}$ $+$ $6$~$\mathsf{CLASS}$ $=$ $21$ components ($71\%$ constructive). (C)~Detection/existence bifurcation: detecting the Heegner point via canonical height positivity is $\mathsf{BISH}$; bounding Mordell--Weil rank via the Euler system is $\mathsf{CLASS}$. (D)~BSD formula audit: algebraic side $\mathsf{BISH}$, analytic side $\mathsf{CLASS}$. Total Lean: 796 lines, 58~theorems, 0~sorry. Fourteenth CRMLint Squeeze. First CRM audit of a Clay Millennium Problem proof.

%-----------------------------------------------------------------------------
\section{The Root Number Bifurcation: CRM Cost of BSD Detection Controlled by the Functional Equation}
\label{sec:p96}
\textbf{DOI:} 10.5281/zenodo.18869924

\bigskip
\noindent\textbf{Abstract.}
We prove that the CRM cost of BSD detection is controlled by the root number. For rank-0 curves ($w = +1$), the ratio $L(E,1)/\Omega^+$ is a rational number computable by modular symbols --- if $L(E,1) \neq 0$, detection is strictly $\mathsf{BISH}$ (\texttt{norm\_num}). For rank-1 curves ($w = -1$), detection requires the Gross--Zagier formula, which depends on the Rankin--Selberg integral ($\mathsf{CLASS}$). Test cases: $\texttt{11a1}$ (rank~0, $L(E,1)/\Omega^+ = 1/5$ by \texttt{norm\_num}); $\texttt{37a1}$ (rank~1, $2$-descent excision gives rank bound $\mathsf{BISH}$, $\Sha$ finiteness $\mathsf{CLASS}$). Universal detection/existence table across Papers~84--96 confirms: detection is generically $\mathsf{BISH}$ ($3/4$ campaigns); the single exception is controlled by a computable invariant ($w = -1$). Existence is universally $\mathsf{CLASS}$. Total Lean: 1,033 lines, 86~theorems, 0~sorry. $10$~$\mathsf{BISH}$ $+$ $3$~$\mathsf{CLASS}$ ($77\%$). Fifteenth CRMLint Squeeze. BSD analogue of the palindromic bifurcation (Paper~89).

%-----------------------------------------------------------------------------
\section{BSD Landscape Survey --- Null Finding}
\label{sec:p97}

\bigskip
\noindent\textbf{Abstract.}
Internal note. We survey the BSD landscape beyond Papers~95--96 for additional CRMLint opportunities. Result: \emph{null finding}. The Gross--Zagier--Kolyvagin pipeline and the root number bifurcation exhaust the accessible CRM structure of BSD. No further excision or descent was found. This closes the BSD arc of the CRM programme.

%-----------------------------------------------------------------------------
\section{The Motivic CRM Architecture: Why the Six Domains of Mathematics Have Different Logical Signatures}
\label{sec:p98}
\textbf{DOI:} 10.5281/zenodo.18902037

\bigskip
\noindent\textbf{Abstract.}
We construct a motivic framework that unifies the CRM classifications of Papers~50--97. Three calibration theorems: (I)~Hodge bifurcates at $\mathsf{WLPO}$; (II)~BSD rank/Sha decouples into $\mathsf{BISH}$ detection and $\mathsf{CLASS}$ existence; (III)~Taylor--Wiles patching $= \mathsf{WKL}$. The Archimedean Obstruction Thesis (Theorem~5.1): $\mathsf{CLASS}$ enters arithmetic geometry through Betti realization only. De~Rham and \'etale realizations are algebraic ($\mathsf{BISH}$); Betti realization requires the complex topology ($\mathsf{CLASS}$). The Conservation Conjecture---that every $\mathsf{CLASS}$ theorem with a $\mathsf{BISH}$-statable conclusion admits a lower-level proof---is stated as the programme's principal open problem. Companion monograph: ``The Logical Cost of Everything.'' Total Lean: 607~lines, 0~sorry, 0~\texttt{Classical.choice}.

%-----------------------------------------------------------------------------
\section{The Hecke Theta Series Squeeze: Resolving the Dihedral Base Case of Modularity via Mumford's Algebraic Theta Functions}
\label{sec:p99}
\textbf{DOI:} 10.5281/zenodo.18821004

\bigskip
\noindent\textbf{Abstract.}
We prove $\mathrm{CRM}(\mathrm{FLT}) = \mathsf{WKL}$ by resolving the dihedral base case of modularity that blocked the $\mathsf{BISH}$ classification in Paper~68. Three $\mathsf{CLASS} \to \mathsf{BISH}$ excisions: (1)~Poisson summation $\to$ Mumford algebraic theta (de~Rham); (2)~Deligne--Serre $\to$ forward formulaic matching (\'etale); (3)~Chebotarev density $\to$ effective bounds (arithmetic). Validates Paper~98's Archimedean Obstruction Thesis empirically: the DPT $\to$ motives $\to$ proof structure chain is predictive. Ghost proof appendix (Lam\'e 1847). Sixteenth CRMLint application. Total Lean: 1,180~lines (7~files), 0~sorry, 0~\texttt{Classical.choice}.

%=============================================================================
\newpage

%-----------------------------------------------------------------------------
\section{The Kuga--Satake Bifurcation: Absolute Hodge Classes on K3 Surfaces via Shioda--Inose Descent}
\label{sec:p100}
\textbf{DOI:} 10.5281/zenodo.18821010

\bigskip
\noindent\textbf{Abstract.}
We analyse the constructive reverse mathematics of Deligne's theorem that Hodge cycles on K3 surfaces are Absolute Hodge---equivalently, that the Kuga--Satake correspondence is algebraic. (The Hodge conjecture itself, for degree-2 classes on a single K3, is just the Lefschetz $(1,1)$-theorem: unconditional and $\mathsf{BISH}$.) The CRM cost of the Absolute Hodge property bifurcates at Picard number $\rho = 20$: for singular K3 surfaces, the Shioda--Inose structure provides a degree-2 rational map to the Kummer surface of a product of CM elliptic curves, inducing an algebraic Kuga--Satake correspondence without Betti realization ($\mathsf{BISH}$). For generic $\rho < 20$, the Mumford--Tate group is maximal and Deligne's transcendental proof via the Betti comparison isomorphism is the only known route ($\mathsf{CLASS}$). CRMLint decomposition: $10$~$\mathsf{BISH}$ $+$ $5$~$\mathsf{CLASS}$ $=$ $15$ components ($67\%$ $\mathsf{BISH}$). The bifurcation parallels Paper~86's Kani--Rosen bifurcation exactly. First CRM analysis of K3 Hodge theory. Conservation Conjecture stated as the programme's principal open problem (0~counterexamples across 100~papers). Total Lean: 729~lines, 0~sorry, 0~\texttt{Classical.choice}.

%=============================================================================
\newpage

%-----------------------------------------------------------------------------
\section{Archimedean Dark Matter: CRMLint Audit of Scholze's Berkovich Motivic Proof of Independence of $\ell$}
\label{sec:p101}

\bigskip
\noindent\textbf{Abstract.}
We deploy the CRMLint LaTeX scanner against Scholze's Berkovich motivic proof of independence of~$\ell$ for \'etale cohomology---the first external application to an active research programme. The scanner produces 214~alerts across two papers: $62$~$\mathsf{BISH}$ (algebraic core), $65$~parasitic~$\mathsf{WLPO}$ (from $\R$-valued norms in Huber's adic spaces, excisable by switching to decidable $\Q$-valued groups), $26$~motivational, $61$~structural~$\mathsf{WKL}$ (profinite limits, perfectoid tilting). Overall CRM signature: $\mathsf{WKL}$. We discover \emph{Archimedean dark matter}: the v1.0 scanner missed 53\% of classical content in the Liquid Tensor Experiment blueprint; a v1.1 patch adding 11~keyword and 5~anti-marker patterns closes the gap. The methodology appendix establishes the scan-map-classify-search-verify loop with \emph{motivic descent} as the paradigmatic Step~4, and validates the Archimedean Obstruction Thesis of Paper~98: $\mathsf{CLASS}$ enters arithmetic geometry through Betti realization only; Scholze's motivic framework avoids it entirely. 42~pages; no Lean bundle (CRMLint scanner application).

%=============================================================================
\newpage

%-----------------------------------------------------------------------------
\section{Conservation Theorems for the Constructive Reverse Mathematics Programme}
\label{sec:p102}
\textbf{DOI:} 10.5281/zenodo.18821015

\bigskip
\noindent\textbf{Abstract.}
We establish conservation theorems for the CRM programme, formalizing the principal open problem: every $\mathsf{CLASS}$ theorem in arithmetic geometry with a $\mathsf{BISH}$-statable conclusion admits a proof at strictly lower CRM level. Coding lemma and case studies across the programme's 100-paper corpus provide positive evidence. Six Lean files formalize the conjecture and its test cases.

%=============================================================================
\newpage

%-----------------------------------------------------------------------------
\section{The Asymptotic Penumbra: Constructive Reverse Mathematics of the RCT Statistical Pipeline}
\label{sec:p103}
\textbf{DOI:} 10.5281/zenodo.18880102

\bigskip
\noindent\textbf{Abstract.}
We apply constructive reverse mathematics to the complete statistical pipeline of the randomised controlled trial (RCT). The central finding: every standard asymptotic test (two-sample $t$, $z$, logrank) passes through a transcendental gate---the normal CDF~$\Phi$---that converts a rational test statistic into an irrational $p$-value. Comparing this $p$-value with a rational significance threshold~$\alpha$ is not decidable in $\mathsf{BISH}$; it requires Markov's Principle ($\mathsf{MP}$). The \emph{Asymptotic Penumbra}---the band of width $2\varepsilon$ around~$\alpha$ where the Studentized Berry--Esseen bound cannot resolve significance---characterises the logical cost precisely. We prove five theorems: (A)~Penumbra Characterisation, (B)~$\mathsf{MP}$ Requirement, (C)~Subgroup Penalty, (D)~Firth Restoration (Cox MLE drops from $\mathsf{WKL}$ to $\mathsf{BISH}$ via Firth's penalised likelihood), (E)~Equivalence Barrier ($d_{\min} \approx 65{,}127$ events for the penumbra to fall below Cohen's $d = 0.2$). Three landmark trials (GUSTO, COURAGE, ISCHEMIA) are classified: GUSTO lies inside the penumbra ($\varepsilon / \alpha = 0.158$); COURAGE and ISCHEMIA are confirmed $\mathsf{BISH}$-decidable. CRM pipeline: $7$~$\mathsf{BISH}$ $+$ $1$~$\mathsf{BISH}{+}\mathsf{MP}$ $+$ $1$~$\mathsf{WKL}$ $+$ $1$~$\mathsf{LPO}$ $= 10$~stages ($70\%$ $\mathsf{BISH}$). First CRM paper in the medical domain. Total Lean: 956~lines (10~files), 58~theorems, 4~sorry (documented). First paper in the Clinical Sub-Series.

%=============================================================================
\newpage
\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

All mathematical content specified by the author. Lean~4 code generation and \LaTeX{} writing assisted by Claude (Anthropic, Opus 4.6) under human direction. Every theorem verified by the Lean~4 type checker or by the internal logic of the mathematical argument.

\end{document}
