\documentclass[11pt]{article}

% -------------------------------------------------
% Preamble (portable with fallbacks)
% -------------------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
% babel package removed for CI compatibility
\IfFileExists{lmodern.sty}{\usepackage{lmodern}}{} % Latin Modern fonts (optional)
\usepackage{geometry}
\geometry{margin=1in}
% microtype disabled for CI compatibility (requires scalable fonts)
% \IfFileExists{microtype.sty}{\usepackage{microtype}}{}

\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}
\usepackage{xcolor}

% mdframed with fallback
\IfFileExists{mdframed.sty}{
  \usepackage{mdframed}
  \mdfdefinestyle{status}{backgroundcolor=gray!10,linecolor=gray!60!black,linewidth=0.8pt,
    innerleftmargin=6pt,innerrightmargin=6pt,innertopmargin=4pt,innerbottommargin=4pt}
}{
  \newenvironment{mdframed}[1][]{\begin{quote}\itshape}{\end{quote}}
}

% -------------------------------------------------
% Theorem styles
% -------------------------------------------------
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% -------------------------------------------------
% AxCal + GR macros
% -------------------------------------------------
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}

\newcommand{\BISH}{\mathsf{BISH}}
\newcommand{\ZF}{\mathsf{ZF}}
\newcommand{\ZFC}{\mathsf{ZFC}}
\newcommand{\CZF}{\mathsf{CZF}}

\newcommand{\WLPO}{\mathrm{WLPO}}
\newcommand{\LEM}{\mathrm{LEM}}
\newcommand{\FT}{\mathrm{FT}}
\newcommand{\WKLz}{\mathrm{WKL}_0}
\newcommand{\RCA}{\mathrm{RCA}}
\newcommand{\MP}{\mathrm{MP}}
\newcommand{\AC}{\mathrm{AC}}
\newcommand{\ACw}{\mathrm{AC}_\omega}
\newcommand{\DCw}{\mathrm{DC}_\omega}
\newcommand{\DC}{\mathrm{DC}}

\newcommand{\Found}{\mathsf{Found}}
\newcommand{\SigmaZero}{\Sigma_{0}}
\newcommand{\AxCal}{\textsc{AxCal}}
\newcommand{\Frontierpos}{\partial^{+}}

\newcommand{\hChoice}{h_{\mathrm{Choice}}}    % AC/DC
\newcommand{\hComp}{h_{\mathrm{Comp}}}        % FT/WKL
\newcommand{\hLogic}{h_{\mathrm{Logic}}}      % LEM/WLPO/MP

\newcommand{\EFE}{\mathrm{EFE}} % Einstein Field Equations

% -------------------------------------------------
% Title
% -------------------------------------------------
\title{Axiom Calibration for General Relativity (Paper 5):\\
Portals, Profiles, and a Hybrid Plan for EPS and Schwarzschild}
\author{Paul Chun--Kit Lee\\
\texttt{dr.paul.c.lee@gmail.com}\\
New York University, NY}
\date{October 2025}

\begin{document}
\maketitle

\begin{abstract}
We make \emph{Axiom Calibration} (\AxCal) the organizing principle for a foundations‑first study of General Relativity (GR). The paper contributes three AxCal instruments for GR: (I) \emph{witness families} pinned to a fixed $\SigmaZero^{\mathrm{GR}}$ signature; (II) \emph{proof‑route flags} and \emph{portal theorems} that turn standard GR arguments into explicit frontier costs (Zorn$\Rightarrow\{\AC\}$, Limit‑Curve/Ascoli$\Rightarrow\{\FT/\WKLz\}$, Serial‑Chain$\Rightarrow\{\DCw\}$, Reductio$\Rightarrow\{\LEM\}$); and (III) \emph{HeightCertificates} that compose costs across results. On this basis we calibrate five loci: G1 (explicit vacuum checks: mathematically Height~0), G2 (Cauchy/MGHD: PDE core vs.\ Zorn portal), G3 (singularity theorems: compactness and contradiction portals), G4 (maximal extensions: Zorn portal), and G5 (computable evolution: negative template after Pour–El–Richards).

\medskip
\textbf{Methodological stance.} Our Lean 4 implementation uses \texttt{mathlib}, a classical library (ZFC). Accordingly, we position the artifact as a form of \emph{Structural Certification}: Lean checks the \AxCal bookkeeping, measuring the Strategic (Portal) Costs of GR arguments, while deep theorems may be imported as named axioms. We distinguish this from the Infrastructural Cost of the environment. For G1 (Schwarzschild vacuum), we establish a Mathematical Height of 0. The Lean formalization structurally certifies this claim by verifying the absence of portals, while executing the computation within the classical environment.
\end{abstract}

\begin{mdframed}[backgroundcolor=gray!10, linewidth=0pt]
\textbf{IMPORTANT DISCLAIMER}

\textbf{A Case Study: Using Multi-AI Agents to Tackle Formal Mathematics}

This entire Lean 4 formalization project was produced by multi-AI agents working under human direction. All proofs, definitions, and mathematical structures in this repository were AI-generated. This represents a case study in using multi-AI agent systems to tackle complex formal mathematics problems with human guidance on project direction.
\end{mdframed}

\begin{mdframed}[style=status]
\textbf{What is calibrated here (AxCal content).}
For each GR target we (a) define a \emph{witness family} at the pin, (b) mark \emph{proof‑route flags} indicating which portals are used, and (c) emit a \emph{HeightCertificate} with axiswise heights $(\hChoice,\hComp,\hLogic)$. Deep‑dive deliverables (EPS; Schwarzschild) produce mathematically Height~0 certificates verified via structural certification in classical \texttt{mathlib}; imported heavy theorems are recorded as \emph{named axioms} triggering portals and heights.
\end{mdframed}

\tableofcontents

% ===========================================================
\section{AxCal instrumentation for GR}
% ===========================================================
\subsection{Pinned signature \texorpdfstring{$\Sigma_0^{\mathrm{GR}}$}{Sigma0-GR}}
We fix the smooth category (second-countable, Hausdorff manifolds), tensor fields, Lorentzian metrics, Levi--Civita connection, curvature and Einstein tensors, $\EFE$, and pinned exemplars (Minkowski; a Schwarzschild-type vacuum metric). Interpretations must fix $\SigmaZero^{\mathrm{GR}}$.

\subsection{Tokens and witness families}
For foundations $F\in\Found$, we use tokens
\[
[{\rm HasAC}\,F],\ [{\rm HasDC}_{\omega}\,F],\ [{\rm HasFT}\,F],\ [{\rm HasWKL}_0\,F],\ [{\rm HasLEM}\,F],\ [{\rm HasWLPO}\,F].
\]
A \emph{witness family} $\mathcal{W}$ assigns to $F$ a groupoid of witnesses for the target statement over the pin.

\subsection{Proof-route flags and portals}
We make explicit route flags that, when present in a proof, trigger an \emph{AxCal portal}:
\begin{itemize}
\item \textsf{uses\_zorn}: applies Zorn on a $\SigmaZero$-definable poset of extensions $\Rightarrow$ \emph{Zorn portal} \textbf{(requiring AC, thus} $\Frontierpos\supseteq\{\AC\}$\textbf{)}.
\item \textsf{uses\_limit\_curve}: invokes Ascoli--Arzelà / compactness of causal curves $\Rightarrow$ \emph{Compactness portal} \textbf{(requiring FT constructively or WKL}$_0$ \textbf{classically, thus} $\Frontierpos\supseteq\{\FT/\WKLz\}$\textbf{)}.
\item \textsf{uses\_serial\_chain}: builds an infinite dependent chain (e.g. curve prolongation) $\Rightarrow$ \emph{Dependent-Choice portal} \textbf{(requiring DC}$_\omega$\textbf{, thus} $\Frontierpos\supseteq\{\DCw\}$\textbf{)}.
\item \textsf{uses\_reductio}: essential proof by contradiction on $\SigmaZero$ data $\Rightarrow$ \emph{LEM portal} \textbf{(requiring LEM, thus} $\Frontierpos\supseteq\{\LEM\}$\textbf{)}.
\end{itemize}

\begin{proposition}[Portal soundness]\label{prop:portals}
If a proof of a $\SigmaZero^{\mathrm{GR}}$-pinned statement uses a flagged route, the corresponding token is necessary along that route: Zorn $\Rightarrow$ HasAC; Limit-Curve $\Rightarrow$ HasFT/HasWKL$_0$ (depending on constructive/classical base); Serial-Chain $\Rightarrow$ HasDC$\omega$; Reductio $\Rightarrow$ HasLEM.
\end{proposition}

\begin{proof}[Sketch]
These are standard meta-implications: Zorn is equivalent to AC over $\ZF$; Ascoli-type compactness aligns with $\FT$ constructively (and with $\WKLz$ classically); building infinite dependent sequences is a canonical use of $\DCw$; essential reductio uses $\LEM$. The novelty is the \emph{proof-route} tagging that transports these meta-results to the $\SigmaZero^{\mathrm{GR}}$ pin. See Appendix~\ref{sec:portal-soundness} for detailed arguments.
\end{proof}

\subsection{Height certificates and composition}
Given portals triggered for a witness family $\mathcal{W}$, we record a HeightCertificate with $(\hChoice,\hComp,\hLogic)\in\{0,1,\omega\}^3$. Products of claims compose componentwise by the AxCal product law (Paper~3A).

\subsection{Scope of Calibration: Portals vs. Infrastructure}
Our hybrid implementation approach, utilizing the classical \texttt{mathlib}, requires a clear distinction regarding the scope of the AxCal measurements.

\begin{definition}[Axiomatic Cost Scopes]
We distinguish between:
\begin{itemize}
    \item \textbf{Infrastructural Cost:} The foundational axioms assumed by the formalization environment (e.g., classical logic within \texttt{mathlib}'s definitions of real numbers and calculus).
    \item \textbf{Strategic (Portal) Cost:} The axioms required by the high-level mathematical arguments specific to the GR proofs, tracked via the portals defined in Section 1.3.
\end{itemize}
\end{definition}

The AxCal framework, as deployed in this paper via \textbf{Structural Certification}, measures the \textbf{Strategic (Portal) Cost}. We refer to this measurement as the \textbf{Mathematical Height}.

\begin{remark}[Interpretation of Height 0]
A Height 0 certificate (profile $(0,0,0)$) signifies that the proof route does not pass through any high-level portals. It acknowledges the mathematical constructivity of the argument structure, even if the formal artifact relies on a classical infrastructure (Infrastructural Cost) to verify the correctness of the individual steps.
\end{remark}

% ===========================================================
\section{Literature anchors mapped to portals}
% ===========================================================
Robb and Reichenbach provide axiomatic scaffolding for kinematics; EPS derives Lorentz classes from light and free-fall~\cite{Robb1914,Reichenbach1969,EPS1972} (no Zorn; compactness may enter via curve families $\Rightarrow$ Compactness portal when maximizers are extracted). Pour--El--Richards show computable well-posed PDEs can yield non-computable evolutions~\cite{PourElRichards1989} (Logic/Computability axis). Bishop--Bridges and Hellman/Bridges guide which analytic steps are Height~0 and which align with choice or $\LEM$~\cite{BishopBridges1985,Hellman1998,BridgesReply1995}. Wald, Hawking--Ellis, and Choquet--Bruhat are used to \emph{locate} where standard GR proofs instantiate portals~\cite{Wald1984,HawkingEllis1973,ChoquetBruhat2009}.

% ===========================================================
\section{Calibration targets (G1--G5) with AxCal profiles}
% ===========================================================
\subsection*{G1. Explicit vacuum checks (Mathematical Height 0)}
\begin{definition}[G1 witness]
$\mathcal{C}^{\mathrm{G1}}$: the pinned Schwarzschild-type metric satisfies vacuum $\EFE$ at the pin.
\end{definition}
\begin{proposition}[G1 profile]\label{prop:G1}
The Mathematical Height (Strategic Cost) is $(\hChoice,\hComp,\hLogic)=(0,0,0)$.
\end{proposition}
\begin{proof}[Sketch and Certification Status]
The proof involves finite symbolic computation of $\Gamma^\alpha_{\mu\nu}$, $R_{\mu\nu}$, and $G_{\mu\nu}$ (cf.\ \cite[§B.4]{Wald1984}). This process is inherently constructive and does not trigger any high-level portals.

\emph{Certification Status:} Our Lean artifact provides Structural Certification for this $(0,0,0)$ profile by verifying the absence of portals in the proof route. As noted in Section 1.5, while the implementation (Deep Dive D2) utilizes the classical infrastructure of \texttt{mathlib} (Infrastructural Cost), the AxCal framework correctly certifies that the Strategic Cost is zero.
\end{proof}

\subsection*{G2. Cauchy problem (local well-posedness and MGHD)}
\begin{definition}[G2 witness]
$\mathcal{C}^{\mathrm{G2}}$: local well-posedness for $\EFE$ and existence/uniqueness of MGHD from data $(\Sigma,h,K)$.
\end{definition}
\begin{proposition}[G2 profile (route-separated)]\label{prop:G2}
The calibration splits:
\begin{itemize}
    \item \textbf{Local PDE:} In standard separable settings (harmonic gauge, energy estimates), the profile is typically $(0,0,0)$. It may rise to $(1,0,0)$ if the analysis requires selecting sequences of approximants without explicit witnesses (via $\ACw$).
    \item \textbf{MGHD:} The standard maximality step has \textsf{uses\_zorn}, resulting in a profile $(\hChoice,\hComp,\hLogic)\ge(1,0,0)$ for the global statement.
\end{itemize}
\end{proposition}
\begin{proof}[Sketch]
Local existence follows Choquet--Bruhat's PDE machinery \cite{ChoquetBruhat2009}, which can often be managed constructively or with $\ACw$ in separable spaces. MGHD standard proofs invoke Zorn on the poset of developments (Wald \cite[Thm.~10.1.2]{Wald1984}); by Prop.~\ref{prop:portals}, the Zorn portal yields the $\AC$ frontier.
\end{proof}

\subsection*{G3. Singularity theorems (Penrose/Hawking)}
\begin{definition}[G3 witness]
$\mathcal{C}^{\mathrm{G3}}$: under trapped surface + energy conditions in a globally hyperbolic spacetime, geodesic incompleteness holds.
\end{definition}
\begin{proposition}[G3 profile]\label{prop:G3}
Standard proofs trigger \textsf{uses\_limit\_curve} and \textsf{uses\_reductio}; hence $(\hChoice,\hComp,\hLogic)\ge(0,1,1)$ (the 1 on the second axis reads as FT/WKL\(_0\) compactness; the 1 on the third as LEM).
\end{proposition}
\begin{proof}[Sketch]
Raychaudhuri focusing plus limit-curve compactness (\cite[§8]{HawkingEllis1973}, \cite[§14]{Wald1984}) gives a maximizing geodesic; contradiction shows incompleteness. Portals: compactness and reductio.
\end{proof}

\subsection*{G4. Maximal extensions}
\begin{definition}[G4 witness]
$\mathcal{C}^{\mathrm{G4}}$: any local solution admits a maximal extension by isometric inclusion.
\end{definition}
\begin{proposition}[G4 profile]\label{prop:G4}
\textsf{uses\_zorn} holds; $(\hChoice,\hComp,\hLogic)\ge(1,0,0)$.
\end{proposition}
\begin{proof}
Chains of extensions admit upper bounds; Zorn yields a maximal element (portal to $\AC$).
\end{proof}

\subsection*{G5. Computable evolution}
\begin{definition}[G5 witness]
$\mathcal{C}^{\mathrm{G5}}$: computable initial data yield computable evolved fields (in a fixed representation) on a pinned globally hyperbolic class.
\end{definition}
\begin{proposition}[G5 negative template]\label{prop:G5}
Without added uniformity, $\mathcal{C}^{\mathrm{G5}}$ can fail by a Pour--El--Richards template \cite{PourElRichards1989}. The failure calibrates on the Logic/Computability axis; any attempt to extract definite infinite data sequences invokes \textsf{uses\_serial\_chain} (portal to $\{\DCw\}$).
\end{proposition}
\begin{proof}[Sketch]
Linear prototypes show computable$\to$non-computable evolution; for quasi-linear systems, similar non-uniformity can occur. If one insists on classical infinite sample paths from measurement-like procedures, the $\DCw$ portal is triggered.
\end{proof}

% ===========================================================
\section{Auxiliary lemmas (EPS; limit-curve) with sketches}
% ===========================================================
\begin{lemma}[EPS kinematics]\label{lem:EPS}
Under the Ehlers--Pirani--Schild axioms, light rays define a conformal structure and free fall a projective structure; compatibility yields a Weyl structure whose scale integrability produces a Lorentz metric class.
\end{lemma}
\begin{proof}[Sketch]
EPS axioms isolate the null cones and (unparameterized) timelike geodesics; the compatibility condition yields a torsion-free Weyl connection preserving the conformal class; vanishing length curvature selects a Levi--Civita metric representative \cite{EPS1972}. No Zorn; compactness may enter only if one extracts maximizing curves.
\end{proof}

\begin{lemma}[Limit-curve compactness]\label{lem:limitcurve}
In globally hyperbolic spacetimes, causal curves between compact sets form a compact set in the $C^0$ topology; maximizing causal geodesics exist.
\end{lemma}
\begin{proof}[Sketch]
Global hyperbolicity yields compact diamond sets; equicontinuity gives Ascoli--Arzelà; upper semicontinuity of length gives maximizers (cf.\ \cite[§14]{Wald1984}). This triggers the Compactness portal.
\end{proof}

% ===========================================================
\section{Hybrid plan: structured framework + selective deep dives}
% ===========================================================

\paragraph{Pragmatic Necessity.} Formalizing the differential geometry and PDE analysis required for General Relativity within a purely constructive framework is currently infeasible due to the lack of mature constructive libraries in Lean. Therefore, we adopt this hybrid approach, leveraging the infrastructure of the classical \texttt{mathlib} to manage domain complexity while strictly enforcing the AxCal methodology to track axiomatic dependencies via Structural Certification.

\subsection{Sprint 3 Milestone: Schwarzschild Vacuum Solution}
\textbf{Achievement:} Complete symbolic verification of the Schwarzschild vacuum solution, demonstrating that the Schwarzschild metric is a solution to Einstein's vacuum field equations $R_{\mu\nu} = 0$. The implementation includes:
\begin{enumerate}
\item \textbf{Metric components:} Diagonal metric with $g_{tt} = -f(r)$, $g_{rr} = f(r)^{-1}$, $g_{\theta\theta} = r^2$, $g_{\varphi\varphi} = r^2\sin^2\theta$, where $f(r) = 1 - 2M/r$.
\item \textbf{Christoffel symbols:} All 40 potentially non-zero symbols computed, with 9 non-vanishing components identified and verified.
\item \textbf{Ricci tensor:} Complete computation using the Ricci formula with systematic treatment of radial derivatives and angular components.
\item \textbf{Vacuum verification:} Explicit proof that all diagonal Ricci components vanish: $R_{tt} = R_{rr} = R_{\theta\theta} = R_{\varphi\varphi} = 0$.
\end{enumerate}
This represents a Height~0 achievement as no axiom portals are triggered—the entire computation proceeds through finite symbolic algebra.

\paragraph{Structured framework (implemented).} We have registered G1--G5 witness families as abstract propositions over the pinned signature $\Sigma_0^{\mathrm{GR}}$. Route flags are attached per standard proofs, and HeightCertificates are emitted using portal soundness (Prop.~\ref{prop:portals}). The framework maintains a verification ledger of \emph{named axioms} (MGHD existence, Penrose/Hawking singularity theorems, limit-curve compactness) with bibliographic citations. This structured approach provides machine-checkable height profiles while using abstract placeholders (\texttt{True}) for the detailed mathematical content---a deliberate design choice that prioritizes axiomatic calibration over full formalization.

\paragraph{Deep dive anchors (Height 0 demonstrations).}
\emph{We utilize \texttt{mathlib}'s infrastructure (e.g., calculus, real numbers) for these deep dives due to the pragmatic constraints mentioned above. This allows us to verify the mathematical steps classically while the AxCal framework certifies the absence of high-level axiom portals.}
\begin{itemize}
\item \textbf{D1 (EPS Core).} We implement the EPS kinematics framework showing how light rays and free fall determine metric structure. The implementation provides a mathematically Height~0 certificate by avoiding all portals, demonstrating that the EPS reconstruction can be done constructively. The Lean artifact verifies this in classical \texttt{mathlib}, providing structural certification of the portal‑free nature.
\item \textbf{D2 (Schwarzschild Engine).} We implement a tensor engine for the Schwarzschild vacuum check (detailed in \texttt{GR/Schwarzschild.lean}). This includes formalizing the metric components, Christoffel symbols, and the Ricci tensor computation using \texttt{mathlib}'s infrastructure for calculus and real analysis. While the underlying symbolic derivations $G_{\mu\nu}=0$ are mathematically Height~0 (no axiom portals), the Lean artifact verifies the computation using classical definitions, thus providing structural certification.
\end{itemize}

\begin{mdframed}[style=status]
\textbf{Success metrics.} (i) D1 and D2 compiled without \texttt{sorry} in the classical \texttt{mathlib} environment; (ii) HeightCertificates present for all G1--G5 tracking axiom dependencies; (iii) explicit portal flags in ledger showing frontier costs; (iv) CI and "no‑sorry" guards for deep‑dive directories. The artifact provides \emph{Structural Certification} of the AxCal bookkeeping, with mathematical Height~0 claims validated in a classical proof assistant.
\end{mdframed}

\begin{mdframed}[backgroundcolor=blue!5, linecolor=blue!30, linewidth=0.8pt]
\textbf{Reproducibility Box: Building and Verifying Paper 5}

\noindent\textbf{Prerequisites:}
\begin{verbatim}
# Install elan (Lean version manager)
curl https://raw.githubusercontent.com/leanprover/elan/master/elan-init.sh -sSf | sh

# Clone repository
git clone https://github.com/AICardiologist/FoundationRelativity.git
cd FoundationRelativity
\end{verbatim}

\noindent\textbf{Build Commands:}
\begin{verbatim}
# Toolchain is pinned by the repository; no manual override needed
# (elan will read ./lean-toolchain automatically)

# Clean and update dependencies
lake clean && lake update

# Build Paper 5 AxCal framework
lake build Papers.P5_GeneralRelativity.Main

# Run smoke test (verifies all components)
lake build Papers.P5_GeneralRelativity.Smoke

# Verify no sorries in implementation (checks all subdirectories, excluding comments)
grep -r "sorry" Papers/P5_GeneralRelativity/ | grep "\.lean:" | \
  grep -v "--.*sorry" | wc -l  # Should output: 0

# Run schematic/axiom audits (match CI)
bash scripts/SchematicAudit.sh
bash scripts/AxiomDeclAudit.sh

# Optional (informational): print transitive axiom dependencies
# (will not fail; used for human-readable auditing in CI logs)
lake env lean scripts/AxiomAudit.lean || true
# Note: This audit will show dependence on classical axioms (e.g., Classical.choice)
# imported via mathlib. This reflects the Infrastructural Cost and is consistent 
# with our Structural Certification methodology (see Section 1.5).
\end{verbatim}

\noindent\textbf{Verification Outputs:}
\begin{itemize}
\item All 7 height certificates compile: G1 (vacuum), G2 (local PDE + MGHD), G3 (Penrose), G4 (maximal extension), G5 (computable evolution + stream)
\item Portal theorems correctly map flags to heights
\item \textbf{Sprint 3 Complete}: Schwarzschild vacuum solution fully verified with explicit Christoffel symbols and Ricci tensor vanishing
\item EPS and Schwarzschild Height 0 anchors verified (structurally)
\item Profile computation: Zorn $\to$ $(1,0,0)$, Limit-Curve $\to$ $(0,1,0)$, Serial-Chain $\to$ $(1,0,0)$, Reductio $\to$ $(0,0,1)$
\end{itemize}
\end{mdframed}

% ===========================================================
\section{Calibration table (profiles at a glance)}
% ===========================================================
\begin{center}
\begin{tabular}{|l|c|l|l|}
\hline
\textbf{Target} & \textbf{Profile $(\hChoice,\hComp,\hLogic)$}${}^*$ & \textbf{Flags/Portals used} & \textbf{Notes} \\
\hline
G1: explicit vacuum & $(0,0,0)$ & none & Mathematical Height 0; Structurally Certified \\
\hline
G2: Cauchy/MGHD & $\ge(1,0,0)$ (global) & \textsf{uses\_zorn} & local PDE core can be $(0,0,0)$/$\ACw$ \\
\hline
G3: singularities & $\ge(0,1,1)$ & \textsf{uses\_limit\_curve}, \textsf{uses\_reductio} & compactness + contradiction \\
\hline
G4: maximal extension & $\ge(1,0,0)$ & \textsf{uses\_zorn} & Zorn portal \\
\hline
G5a: computability negative (PER) & $(0,0,0)$ & none & negative template; no portal cost tracked \\
\hline
G5b: measurement stream & $(1,0,0)$ & \textsf{uses\_serial\_chain} & DC$_\omega$ portal on dependent streams \\
\hline
\end{tabular}
\end{center}
${}^*$\footnotesize{Profiles indicate the Mathematical Height (Strategic Costs; see Section 1.5), not the Infrastructural Cost of the classical formalization environment.}

% ===========================================================
\section{Artifact Mapping: Paper Claims to Lean Implementation}
% ===========================================================

\subsection{Core AxCal Infrastructure}

\begin{center}
\small
\begin{tabular}{|p{5cm}|p{5.5cm}|p{4cm}|}
\hline
\textbf{Paper Concept} & \textbf{Lean Symbol} & \textbf{Module} \\
\hline
Height levels $(0, 1, \omega)$ & \texttt{Height.zero/.one/.omega} & \texttt{AxCalCore.Axis} \\
\hline
Axis profile $(\hChoice, \hComp, \hLogic)$ & \texttt{AxisProfile} & \texttt{AxCalCore.Axis} \\
\hline
Witness family type & \texttt{WitnessFamily} & \texttt{AxCalCore.Axis} \\
\hline
Height certificate structure & \texttt{HeightCertificate} & \texttt{GR.Certificates} \\
\hline
Portal flags (Zorn, Limit-Curve, etc.) & \texttt{PortalFlag} & \texttt{GR.Portals} \\
\hline
Route-to-profile mapping & \texttt{route\_to\_profile} & \texttt{GR.Portals} \\
\hline
\end{tabular}
\end{center}

\subsection{GR-Specific Components}

\begin{center}
\small
\begin{tabular}{|p{5cm}|p{5.5cm}|p{4cm}|}
\hline
\textbf{Paper Concept} & \textbf{Lean Symbol} & \textbf{Module} \\
\hline
Pinned signature $\Sigma_0^{\mathrm{GR}}$ & \texttt{Spacetime}, \texttt{LorentzMetric} & \texttt{GR.Interfaces} \\
\hline
Einstein Field Equations & \texttt{EFE}, \texttt{VacuumEFE} & \texttt{GR.Interfaces} \\
\hline
Schwarzschild pinning & \texttt{IsPinnedSchwarzschild} & \texttt{GR.Interfaces} \\
\hline
\end{tabular}
\end{center}

\subsection{Calibration Targets (G1--G5)}

\begin{center}
\small
\begin{tabular}{|p{3cm}|p{4.5cm}|p{4cm}|p{3cm}|}
\hline
\textbf{Target} & \textbf{Witness Family} & \textbf{Certificate} & \textbf{Verified Profile} \\
\hline
G1: Vacuum & \texttt{GR.G1\_Vacuum\_W} & \texttt{G1\_Vacuum\_Cert} & $(0,0,0)$ \checkmark$^S$ \\
\hline
G2: Local PDE & \texttt{GR.G2\_LocalPDE\_W} & \texttt{G2\_LocalPDE\_Cert} & $(0,0,0)$ \checkmark$^S$ \\
\hline
G2: MGHD & \texttt{GR.G2\_MGHD\_W} & \texttt{G2\_MGHD\_Cert} & $(1,0,0)$ \checkmark$^S$ \\
\hline
G3: Penrose & \texttt{GR.G3\_Penrose\_W} & \texttt{G3\_Penrose\_Cert} & $(0,1,1)$ \checkmark$^S$ \\
\hline
G4: MaxExt & \texttt{GR.G4\_MaxExt\_W} & \texttt{G4\_MaxExt\_Cert} & $(1,0,0)$ \checkmark$^S$ \\
\hline
G5: CompNeg & \texttt{GR.G5\_CompNeg\_W} & \texttt{G5\_CompNeg\_Cert} & $(0,0,0)$ \checkmark$^S$ \\
\hline
G5: Stream & \texttt{GR.G5\_MeasStream\_W} & \texttt{G5\_MeasStream\_Cert} & $(1,0,0)$ \checkmark$^S$ \\
\hline
\end{tabular}
\end{center}
\noindent\small{\checkmark$^S$: Structurally Certified (AxCal bookkeeping verified in Lean using \texttt{mathlib}).}

\subsection{Deep-Dive Deliverables (Height 0 Anchors)}

\begin{center}
\small
\begin{tabular}{|p{4cm}|p{5cm}|p{5.5cm}|}
\hline
\textbf{Deliverable} & \textbf{Main Theorem} & \textbf{Implementation Status} \\
\hline
D1: EPS Kinematics Core & \texttt{EPS\_Height\_Zero} & Complete framework \\
& \texttt{EPS\_Kinematics\_Height0} & \texttt{GR.EPSCore} \\
\hline
D2: Schwarzschild Vacuum & \texttt{Ricci\_vanishing} & \textbf{Sprint 3 Complete:} \\
& \texttt{Schwarzschild\_is\_vacuum} & Full symbolic verification \\
& & \texttt{GR.Schwarzschild} \\
\hline
\end{tabular}
\end{center}

\subsection{Portal Theorems}

\begin{center}
\small
\begin{tabular}{|p{3.5cm}|p{5cm}|p{6cm}|}
\hline
\textbf{Portal} & \textbf{Lean Implementation} & \textbf{Effect on Profile} \\
\hline
Zorn Portal & \texttt{Zorn\_portal} axiom & $\hChoice \gets 1$ \\
\hline
Limit-Curve Portal & \texttt{LimitCurve\_portal} axiom & $\hComp \gets 1$ \\
\hline
Serial-Chain Portal & \texttt{SerialChain\_portal} axiom & $\hChoice \gets 1$ (via DC$_\omega$) \\
\hline
Reductio Portal & \texttt{Reductio\_portal} axiom & $\hLogic \gets 1$ \\
\hline
\end{tabular}
\end{center}

\subsection{Verification Infrastructure}

\begin{center}
\small
\begin{tabular}{|p{4cm}|p{5cm}|p{5.5cm}|}
\hline
\textbf{Component} & \textbf{Lean Symbol} & \textbf{Purpose} \\
\hline
Main aggregator & \texttt{Paper5\_Main} & Verifies framework completeness \\
& & \texttt{Main} \\
\hline
Profile computation test & \texttt{Profile\_Computation\_Works} & Tests portal$\to$height mapping \\
& & \texttt{Main} \\
\hline
Smoke test & \texttt{Paper5\_Smoke\_Success} & CI aggregator, no-sorry guard \\
& & \texttt{Smoke} \\
\hline
Certificate registry & \texttt{Certificates.all\_certificates} & Lists all 7 height certificates \\
& & \texttt{GR.Certificates} \\
\hline
\end{tabular}
\end{center}

\subsection{Verification Ledger}

The AxCal framework maintains a structured ledger that tracks the provenance of each height assignment. Each certificate in \texttt{GR/Certificates.lean} includes: (i) the witness family defining the mathematical claim, (ii) the list of portal flags triggered by the standard proof route, (iii) the resulting height profile computed via \texttt{route\_to\_profile}, (iv) bibliographic citations to the source literature, and (v) a constructive upper bound proof or axiom import. This ledger design ensures that height costs are auditable and that alternative proof routes (avoiding certain portals) can be systematically explored. The framework correctly computes that Zorn's lemma triggers $\hChoice = 1$, limit‑curve compactness triggers $\hComp = 1$, and proof by contradiction triggers $\hLogic = 1$. All seven certificates (G1 vacuum, G2 local PDE, G2 MGHD, G3 Penrose, G4 maximal extension, G5 computability negative, G5 measurement stream) compile without \texttt{sorry} in classical \texttt{mathlib} and produce the expected height profiles as verified by \texttt{Paper5\_Main} and the smoke tests. This constitutes \emph{Structural Certification}—a machine-auditable verification of the AxCal bookkeeping and the logical composition of dependencies—rather than a formal constructive proof of the object-level GR theorems.

% ===========================================================
\section{Conclusion}
% ===========================================================
This paper is not a GR formalization for its own sake: it is an \emph{AxCal map} of GR. Portals, route flags, and HeightCertificates turn the folklore "this uses choice/compactness/LEM" into machine‑checkable artifacts that compose across the theory. The deep‑dive tasks (EPS; Schwarzschild) supply mathematically Height~0 anchors verified via \emph{Structural Certification} in classical \texttt{mathlib}, ensuring the project yields verifiable AxCal infrastructure while the schematic layer documents axiomatic cost with precision. 

\paragraph{Future Work.} To move from Structural Certification to full Foundational Verification for Height 0 targets (G1, D1, D2), future work will involve porting these implementations to a formal constructive library, should one become available for Lean 4.

% ===========================================================
\section*{Acknowledgments}
% ===========================================================
Development assistance provided by: Gemini 2.5 Deep Think (architecture exploration and theoretical framework design), GPT-5 Pro (Lean 4 scaffolding and implementation support), and Claude Code (repository management and development workflow).

% ===========================================================
\begin{thebibliography}{99}

\bibitem{Robb1914}
A.~A.~Robb.
\newblock {\em A Theory of Time and Space}.
\newblock Cambridge University Press, 1914.

\bibitem{Reichenbach1969}
H.~Reichenbach.
\newblock {\em Axiomatization of the Theory of Relativity}.
\newblock University of California Press, 1969 (transl.\ of 1924).

\bibitem{EPS1972}
J.~Ehlers, F.~A.~E.~Pirani, and A.~Schild.
\newblock The geometry of free fall and light propagation.
\newblock In {\em General Relativity: Papers in Honour of J.~L.~Synge}, 1972.

\bibitem{PourElRichards1989}
M.~B.~Pour-El and J.~I.~Richards.
\newblock {\em Computability in Analysis and Physics}.
\newblock Springer, 1989.

\bibitem{BishopBridges1985}
E.~Bishop and D.~S.~Bridges.
\newblock {\em Constructive Analysis}.
\newblock Springer, 1985.

\bibitem{Hellman1998}
G.~Hellman.
\newblock Mathematical constructivism in spacetime physics.
\newblock {\em Brit.~J.~Phil.~Sci.}, 49(3):425--450, 1998.

\bibitem{BridgesReply1995}
D.~S.~Bridges.
\newblock Constructive mathematics and unbounded operators: a reply to Hellman.
\newblock {\em J.~Philosophical Logic}, 24(5):549--561, 1995.

\bibitem{Wald1984}
R.~M.~Wald.
\newblock {\em General Relativity}.
\newblock University of Chicago Press, 1984.

\bibitem{HawkingEllis1973}
S.~W.~Hawking and G.~F.~R.~Ellis.
\newblock {\em The Large Scale Structure of Space-Time}.
\newblock Cambridge University Press, 1973.

\bibitem{ChoquetBruhat2009}
Y.~Choquet--Bruhat.
\newblock {\em General Relativity and the Einstein Equations}.
\newblock Oxford University Press, 2009.

\end{thebibliography}
\appendix


% ===========================
\section{Portals and Proof–Route Flags: Soundness}
\label{sec:portal-soundness}
% ===========================

We make explicit the mechanism that transports standard proof routes in GR into AxCal height costs.

\subsection{Route flags and their semantics}

\begin{definition}[Route flag]
A \emph{route flag} is a marker indicating that a derivation of a witness explicitly invokes a standard device:
\[
\mathrm{Flag} \in \{\mathrm{uses\_zorn},\ \mathrm{uses\_limit\_curve},\ \mathrm{uses\_serial\_chain},\ \mathrm{uses\_reductio}\}.
\]
\end{definition}

\begin{definition}[Usage predicate]
For a given derivation $\mathcal{D}$, $\mathrm{Uses}(\mathrm{Flag},\mathcal{D})$ is the proposition that the corresponding device is \emph{actually used as a step} in $\mathcal{D}$ (not merely available or admissible). Formally, in Lean we record this as a Prop argument to the witness:
\[
\mathsf{Uses}\ \mathrm{Flag} \quad\text{is a hypothesis to the witness family for that result.}
\]
\end{definition}

\begin{remark}[Route sensitivity]
Height costs attach to \emph{routes}, not just to statements. If the same theorem admits a proof avoiding a flagged device, then the certificate for that alternate route carries a lower profile.
\end{remark}

\subsection{Zorn's lemma portal (Choice axis)}

\begin{proposition}[Zorn portal: $\mathrm{uses\_zorn}\Rightarrow\{\mathrm{AC}\}$]
\label{prop:zorn-portal}
Over $\ZF$, any derivation $\mathcal{D}$ that uses Zorn's Lemma triggers a positive frontier on the Choice axis: if $\mathrm{Uses}(\mathrm{uses\_zorn},\mathcal{D})$, then $\mathrm{AC}$ is required.
\end{proposition}

\begin{proof}
In $\ZF$, the following are equivalent: the Axiom of Choice (AC), Zorn's Lemma, and the Hausdorff Maximal Principle; see any standard reference (e.g.\ Jech, \emph{Set Theory}, Thm.~8.1; Howard–Rubin, \emph{Consequences of the Axiom of Choice}). Hence invoking Zorn in $\mathcal{D}$ imports a principle equivalent to AC. In AxCal, we reflect this by a portal axiom
\[
\mathrm{Uses}(\mathrm{uses\_zorn},\mathcal{D})\ \Rightarrow\ \mathrm{HasAC}(F),
\]
which feeds the Choice coordinate of the height profile. \qedhere
\end{proof}

\subsection{Limit–curve portal (Compactness axis)}

\begin{proposition}[Limit–curve portal: $\mathrm{uses\_limit\_curve}\Rightarrow\{\FT\ \text{or}\ \WKLz\}$]
\label{prop:limit-curve-portal}
Suppose a GR derivation $\mathcal{D}$ invokes a \emph{limit–curve} argument: from a sequence of causal curves with uniform local bounds (e.g.\ equicontinuity and uniform speed control), $\mathcal{D}$ extracts a convergent subsequence (or a limit curve) without quantitative moduli of compactness. Then
\[
\mathrm{Uses}(\mathrm{uses\_limit\_curve},\mathcal{D}) \ \Rightarrow\ \big(\mathrm{HasFT}(F)\ \text{or}\ \mathrm{HasWKL}_0(F)\big).
\]
\end{proposition}

\begin{proof}[Proof sketch]
Limit–curve arguments in Lorentzian geometry are typically instances of Arzelà–Ascoli–type compactness on spaces of curves (or a diagonal Bolzano–Weierstraß selection). Over classical logic (e.g., within the framework of Reverse Mathematics using $\RCA_0$, a system for recursive comprehension), the necessary sequential compactness for $[0,1]$ and the Bolzano–Weierstraß theorem are equivalent to $\WKLz$ (Simpson, \emph{Subsystems of Second Order Arithmetic}, Chs.~III–IV). Over constructive bases, Heine–Borel/compactness of Cantor space (hence $[0,1]$ via coding) is calibrated by the Fan Theorem $\FT$ (Bishop–Bridges, \emph{Constructive Analysis}; Troelstra–van Dalen, \emph{Constructivism in Mathematics}). Thus a non-quantitative subsequence/limit extraction imports compactness strength: classically $\WKLz$; constructively $\FT$. We record this disjunctively, to be resolved by the chosen base. \qedhere
\end{proof}

\begin{remark}
If one supplies explicit moduli (e.g.\ an effective Arzelà–Ascoli hypothesis), the portal can be avoided; the certificate for that route will then carry a lower compactness height.
\end{remark}

\subsection{Serial–chain portal (Dependent Choice axis)}

\begin{proposition}[Serial–chain portal: $\mathrm{uses\_serial\_chain}\Rightarrow\{\DCw\}$]
\label{prop:dc-portal}
Let $R\subseteq X\times X$ be serial: $\forall x\in X\,\exists y\in X\,(xRy)$. If a derivation $\mathcal{D}$ requires the existence of an infinite $R$–chain $(x_n)_{n\in\mathbb{N}}$ with $x_n R x_{n+1}$, then
\[
\mathrm{Uses}(\mathrm{uses\_serial\_chain},\mathcal{D})\ \Rightarrow\ \mathrm{HasDC}_\omega(F).
\]
In the AxCal profile we record this on the Choice axis: $\hChoice \gets 1$.
\end{proposition}

\begin{proof}
This is exactly the axiom of Dependent Choice (for $\omega$) specialized to a serial relation. In $\ZF$, $\DCw$ is strictly weaker than AC and sufficient to construct such chains; in $\BISH$, the same scheme expresses the iteration of countably many dependent selections. The portal records this as a foundation‑scoped token. \qedhere
\end{proof}

\subsection{Reductio portal (Logic axis)}

\begin{proposition}[Reductio portal: $\mathrm{uses\_reductio}\Rightarrow\{\LEM\}$ (upper bound)]
\label{prop:reductio-portal}
If a derivation $\mathcal{D}$ obtains an \emph{existential} or a \emph{disjunctive} conclusion solely by contradiction (i.e.\ using $\neg\neg\varphi\Rightarrow\varphi$ at top level, not under stable predicates), then
\[
\mathrm{Uses}(\mathrm{uses\_reductio},\mathcal{D})\ \Rightarrow\ \mathrm{HasLEM}(F),
\]
giving an upper bound on the Logic axis.
\end{proposition}

\begin{proof}[Proof sketch]
In intuitionistic/constructive settings, double–negation elimination is not generally valid; it becomes available under $\LEM$ (or specific semi–classical schemes for restricted formula classes). The standard proofs of the singularity theorems—e.g.\ Penrose—often conclude by contradiction from a global completeness hypothesis without providing a constructed witness; see Hawking–Ellis, \emph{Large Scale Structure}, and Wald, \emph{General Relativity}. We therefore mark the route with $\mathrm{uses\_reductio}$ and import $\LEM$ as a conservative upper bound. If a route is reworked into a stable/existentially constructive form, the flag can be removed and the profile lowered. References: Troelstra–van Dalen; Bridges–Richman. \qedhere
\end{proof}

\subsection{Portal soundness summary}

Combining Propositions~\ref{prop:zorn-portal}–\ref{prop:reductio-portal} yields the meta–level transport principle used throughout:
\begin{quote}
\emph{Portal Soundness.} If a certificate includes a set of route flags, then the corresponding tokens on the AxCal axes are admissible in the foundation, yielding the advertised height profile as an \emph{upper bound}. When the route is replaced by one without a given flag, the corresponding coordinate can be lowered.
\end{quote}

\section{AxCal--Lean Ledger}\label{app:ledger}

This appendix records the Lean-facing names for tokens, portal axioms, witness families, and height certificates used in the calibration of G1--G5. It functions as a machine-checkable index aligning the paper's calibration table with repository artifacts.

\subsection{Axis tokens and portal axioms}\label{app:tokens-portals}

\begin{verbatim}
-- AxCal core tokens (foundation-scoped)
class HasAC   (F : Foundation) : Prop
class HasDCw  (F : Foundation) : Prop
class HasFT   (F : Foundation) : Prop
class HasWKL0 (F : Foundation) : Prop
class HasLEM  (F : Foundation) : Prop
class HasWLPO (F : Foundation) : Prop

/-- Proof-route flags (carried as hypotheses in witness families; see §\ref{app:witnesses}) -/
inductive PortalFlag
| uses_zorn
| uses_limit_curve
| uses_serial_chain
| uses_reductio

/-- Portal soundness axioms (paper Proposition~\ref{prop:portals}). 
    They are registered once per foundation F. -/
axiom Zorn_portal     : forall {F}, Uses PortalFlag.uses_zorn         -> HasAC   F
axiom LimitCurve_portal : forall {F}, Uses PortalFlag.uses_limit_curve -> (HasFT F or HasWKL0 F)
axiom SerialChain_portal : forall {F}, Uses PortalFlag.uses_serial_chain -> HasDCw F
axiom Reductio_portal  : forall {F}, Uses PortalFlag.uses_reductio     -> HasLEM F
\end{verbatim}

\paragraph{Notes.}
\begin{itemize}
\item The wrapper \verb|Uses flag| is a Propositional hypothesis attached to the witness (see Appendix B.2). It records that the specific derivation relies on the flagged device (e.g., Zorn's Lemma). This mechanism is what ties the proof \emph{route} to the axiomatic cost.
\item The compactness portal is recorded disjunctively (\verb|HasFT or HasWKL0|) to reflect constructive/classical bases; the certificate chooses the branch used in the imported argument.
\end{itemize}

\subsection{Witness families for G1--G5}\label{app:witnesses}

\begin{verbatim}
-- Pinned signature Sigma0^GR (interfaces only; no mathlib dependency)
structure Manifold := ...
structure LorentzMetric (M : Manifold) := ...
structure Spacetime := (M : Manifold) (g : LorentzMetric M)

-- Einstein tensor interface and EFE predicate
def EinsteinTensor (S : Spacetime) : Tensor := ...
def EFE (S : Spacetime) (T : Tensor) : Prop := ...

-- WitnessFamily type (from AxCal core)
--   WitnessFamily F  := Prop (witness existence over foundation F)

namespace GR

/-- G1: explicit vacuum check (Schwarzschild@pin) -/
def G1_Vacuum_W : WitnessFamily := fun F =>
  forall (Ssch : Spacetime), IsPinnedSchwarzschild Ssch -> EFE Ssch ZeroTensor

/-- G2: Cauchy problem split into local PDE and MGHD (global) -/
def G2_LocalPDE_W : WitnessFamily := fun F =>
  forall (ID : InitialData), LocalWellPosed ID         -- no portal flags
def G2_MGHD_W     : WitnessFamily := fun F =>
  forall (ID : InitialData), Uses PortalFlag.uses_zorn -> MGHD_Exists ID

/-- G3: Singularity theorem (schematic Penrose) -/
def G3_Penrose_W : WitnessFamily := fun F =>
  forall (S : Spacetime),
    (NullEnergyCondition S) →
    (HasTrappedSurface S)    →
    Uses PortalFlag.uses_limit_curve →
    Uses PortalFlag.uses_reductio     →
    ¬ GeodesicallyComplete S

/-- G4: Maximal extension existence -/
def G4_MaxExt_W : WitnessFamily := fun F =>
  forall (S : Spacetime),
    Uses PortalFlag.uses_zorn →
    exists Smax, IsMaximalExtension S Smax

/-- G5: Computable evolution (negative template and DC stream) -/
def G5_CompNeg_W : WitnessFamily := fun F =>
  exists (class : GHClass),
    ComputableInitialData class and
    NonComputableEvolution class  -- PER-style failure

def G5_MeasStream_W : WitnessFamily := fun F =>
  HasDCw F -> (forall proto : SerialProtocol, InfiniteHistory proto)

end GR
\end{verbatim}

\subsection{Height certificates (profiles and routes)}\label{app:certs}

\begin{verbatim}
-- Axis triple: (Choice, Compactness, Logic)
structure AxisProfile := (hChoice hComp hLogic : Height)   -- Height in {zero, one, omega}
structure HeightCertificate :=
{ W        : WitnessFamily
, profile  : AxisProfile
, flags    : List PortalFlag
, upper    : ProfileUpper profile W        -- constructive upper proof or portal imports
, cites    : List Citation                 -- paper-level references used
}

-- Concrete certificates (G1--G5)
def G1_Vacuum_Cert : HeightCertificate :=
{ W       := GR.G1_Vacuum_W
, profile := <zero, zero, zero>
, flags   := []
, upper   := by
    -- symbolic curvature computation at the pin (no portals)
    exact upper_height0_vacuum_check
, cites   := [cite "Wald §B.4"]
}

def G2_LocalPDE_Cert : HeightCertificate :=
{ W       := GR.G2_LocalPDE_W
, profile := <zero, zero, zero>  -- or ⟨one, zero, zero> if ACw is used in analysis
, flags   := []
, upper   := import_local_pde_result
, cites   := [cite "Choquet–Bruhat (2009)"]
}

def G2_MGHD_Cert : HeightCertificate :=
{ W       := GR.G2_MGHD_W
, profile := <one, zero, zero>
, flags   := [PortalFlag.uses_zorn]
, upper   := by
    intro F ID hzorn
    have hAC : HasAC F := Zorn_portal hzorn
    exact imported_mghd_existence hAC
, cites   := [cite "Wald Thm. 10.1.2"]
}

def G3_Penrose_Cert : HeightCertificate :=
{ W       := GR.G3_Penrose_W
, profile := <zero, one, one>
, flags   := [PortalFlag.uses_limit_curve, PortalFlag.uses_reductio]
, upper   := by
    intro F S nec trapped hlim hred
    have hComp : (HasFT F or HasWKL0 F) := LimitCurve_portal hlim
    have hLEM  : HasLEM F              := Reductio_portal hred
    exact imported_penrose hComp hLEM nec trapped
, cites   := [cite "Hawking–Ellis §8", cite "Wald §14"]
}

def G4_MaxExt_Cert : HeightCertificate :=
{ W       := GR.G4_MaxExt_W
, profile := <one, zero, zero>
, flags   := [PortalFlag.uses_zorn]
, upper   := by
    intro F S hz
    exact imported_maximal_extension (Zorn_portal hz)
, cites   := [cite "Wald §10.1"]
}

def G5_CompNeg_Cert : HeightCertificate :=
{ W       := GR.G5_CompNeg_W
, profile := <zero, zero, zero>  -- negative template; no portal cost tracked
, flags   := []
, upper   := imported_PER_negative_template
, cites   := [cite "Pour–El–Richards (1989)"]
}

def G5_MeasStream_Cert : HeightCertificate :=
{ W       := GR.G5_MeasStream_W
, profile := <one, zero, zero>  -- DCω sits on the Choice axis
, flags   := [PortalFlag.uses_serial_chain]
, upper   := by
    intro F hDC proto
    exact SerialChain_portal_elim hDC proto
, cites   := [cite "AxCal DCω eliminator"]
}
\end{verbatim}

\subsection{Verification table (names / profiles / portals)}\label{app:table}

\begin{center}
\begin{tabular}{@{}lllll@{}}
\hline
\textbf{Target} & \textbf{Witness (Lean)} & \textbf{Certificate (Lean)} & \textbf{Flags} & \textbf{Profile}\\
\hline
G1 & \verb|GR.G1_Vacuum_W| & \verb|G1_Vacuum_Cert| & -- & $(0,0,0)$ \\
G2 (local) & \verb|GR.G2_LocalPDE_W| & \verb|G2_LocalPDE_Cert| & -- & $(0,0,0)$ or $(1,0,0)$ \\
G2 (MGHD) & \verb|GR.G2_MGHD_W| & \verb|G2_MGHD_Cert| & Zorn & $(1,0,0)$ \\
G3 & \verb|GR.G3_Penrose_W| & \verb|G3_Penrose_Cert| & LimitCurve, Reductio & $(0,1,1)$ \\
G4 & \verb|GR.G4_MaxExt_W| & \verb|G4_MaxExt_Cert| & Zorn & $(1,0,0)$ \\
G5 (neg.) & \verb|GR.G5_CompNeg_W| & \verb|G5_CompNeg_Cert| & -- & $(0,0,0)$ \\
G5 (stream) & \verb|GR.G5_MeasStream_W| & \verb|G5_MeasStream_Cert| & SerialChain & $(1,0,0)$ \\
\hline
\end{tabular}
\end{center}

\subsection{File map (proposed layout)}\label{app:files}

\begin{verbatim}
Papers/P5_GR/
  AxCalCore/Axis.lean              -- Height, AxisProfile, ProfileUpper
  AxCalCore/Tokens.lean            -- HasAC, HasDCw, HasFT, HasWKL0, HasLEM, HasWLPO
  GR/Interfaces.lean               -- Sigma0^GR: manifolds, Lorentz metrics, EFE predicate
  GR/Portals.lean                  -- PortalFlag, Zorn_portal, LimitCurve_portal, ...
  GR/Witnesses.lean                -- G1_Vacuum_W, G2_*, G3_*, G4_*, G5_*
  GR/Certificates.lean             -- G*_Cert definitions (HeightCertificate)
  GR/EPSCore.lean                  -- (deep-dive) EPS kinematics proofs (Height 0)
  GR/Schwarzschild.lean            -- (deep-dive) vacuum check engine (Height 0)
  Ledger/Citations.lean            -- structured bibliography handles for certificates
  Smoke.lean                       -- CI aggregator; no-sorry guard for deep-dive dirs
\end{verbatim}

\subsection{Ledger policy}
\begin{itemize}
\item Every certificate includes \emph{flags} (route evidence) and \emph{cites} (bibliographic anchors).
\item Replacing an imported theorem by an internal proof that avoids a flagged portal \emph{automatically} lowers the certificate's profile (the AxCal algebra recomputes heights componentwise).
\item Disjunctive compactness (\verb|HasFT or HasWKL0|) must be resolved per foundation instance to produce a concrete $(\hComp)$ entry.
\item \textbf{Environment acknowledgment:} All certifications are structural, executed within a classical (\texttt{mathlib}) environment. Height 0 certificates (G1, D1, D2) certify the absence of high-level portals (like Zorn or LEM), not the foundational constructivity of the underlying calculus implementation used in the Lean artifact.
\end{itemize}

% ===========================================================
\section{Methodological Discussion: Structural Certification and Physical Insight}
\label{app:methodology}
% ===========================================================

This appendix addresses fundamental questions about the AxCal methodology, its validity, and its value to physics.

\subsection{Defending Structural Certification}

A natural critique of this project might argue: ``You cannot verify constructivity by working in a classical system. The claim of Height (0,0,0) for G1 is contradictory when the verification uses classical \texttt{mathlib}.''

This critique conflates two distinct activities:

\begin{itemize}
\item \textbf{Foundational Verification:} Proving a theorem \emph{within} a constructive system (e.g., BISH). This indeed requires a constructive proof assistant.
\item \textbf{Meta-Analysis (Structural Certification):} Reasoning \emph{about} the dependencies of a proof.
\end{itemize}

It is entirely valid to use a classical meta-theory to analyze proofs in a weaker object theory. For example, Reverse Mathematics practitioners typically use classical logic (ZFC) to prove that a theorem is equivalent to WKL$_0$ over RCA$_0$. They perform meta-analysis.

\textbf{Structural Certification formalizes this meta-analysis.} The claim that G1 has Mathematical Height (0,0,0) means the argument relies only on finite symbolic manipulation and does not invoke high-level non-constructive principles. The Lean artifact verifies:
\begin{enumerate}
\item The individual steps (calculus, algebra) are correct (classically).
\item The overall structure does not invoke the \texttt{uses\_zorn} flag or other portals.
\end{enumerate}

This combination provides a machine-checked guarantee that the high-level architecture of the proof is free from specific non-constructive principles.

\subsection{Physical Value of Axiom Calibration}

One might ask: ``Is this just math for math's sake? Are we gaining genuine physical insight?''

The AxCal framework offers substantial value to the foundations of physics by systematically mapping logical dependencies within GR:

\subsubsection{Clarifying Physics vs. Mathematical Idealization}

\begin{itemize}
\item \textbf{The Axiom of Choice and Physical Reality:} G4 (Maximal Extensions) and G2 (MGHD) rely on Zorn's Lemma. AC guarantees \emph{existence} but provides no algorithm to \emph{construct} these spacetimes. This forces physicists to question whether maximal extensions are physical necessities or mathematical artifacts.

\item \textbf{LEM and Singularities:} G3 (Singularity Theorems) uses proof by contradiction. Standard proofs demonstrate geodesic incompleteness must occur but provide no constructive method to locate singularities. AxCal highlights that singularity formation remains constructively elusive.
\end{itemize}

\subsubsection{The Crisis of Computability}

Physics strives for predictive theories. AxCal rigorously separates constructive (computable) parts of GR from non-constructive ones:

\begin{itemize}
\item \textbf{Limits of Prediction (G5):} The Pour-El-Richards result shows well-posed PDEs can have computable initial conditions evolving into \emph{non-computable} states. If GR exhibits this behavior, it imposes fundamental limits on predictability, independent of computational power.
\end{itemize}

\subsubsection{Grounding Theory in Observables}

\begin{itemize}
\item \textbf{The EPS Framework (D1):} The Ehlers-Pirani-Schild framework derives spacetime structure from light rays and free-falling particles. Formal verification that this reconstruction is constructive (Height 0) demonstrates GR can be grounded in idealized physical observations rather than abstract mathematical assumptions.
\end{itemize}

\subsubsection{Ensuring Correctness in Critical Computations}

\begin{itemize}
\item \textbf{Verified Tensor Calculus (G1/D2):} The Schwarzschild Engine provides verified tensor manipulation. The formal proof that $R_{\mu\nu}=0$ for the Schwarzschild metric guarantees algebraic correctness. This is vital for numerical relativity where simulations must faithfully represent the continuum theory.
\end{itemize}

\subsection{Pragmatic Choices: Classical Tools for Constructive Analysis}

Given the complexity of pure constructive formalization, why not abandon \texttt{mathlib}?

\subsubsection{The Complexity Challenge}

\begin{enumerate}
\item \textbf{Nature of Analysis:} Constructive analysis requires explicit tracking of witnesses and convergence moduli. Basic principles like real number trichotomy fail without decision algorithms.

\item \textbf{Library Support:} No constructive library matches \texttt{mathlib}'s scope for advanced mathematics. Building the necessary infrastructure would dwarf the GR formalization itself.

\item \textbf{Project Focus:} The primary goal is Axiom Calibration of GR, not developing constructive differential geometry libraries.
\end{enumerate}

\subsubsection{Strengthening Structural Certification}

While using a classical environment, we strengthen Height 0 claims through:

\begin{itemize}
\item \textbf{Axiom Auditing:} Using Lean's \texttt{\#print axioms} to inspect dependencies, ensuring Height 0 modules minimize classical principles.

\item \textbf{Linter Discipline:} Forbidding explicit classical mechanisms (like \texttt{by\_contra}) within Height 0 modules, maintaining direct proof style.
\end{itemize}

\subsection{Conclusion}

The AxCal project is not ``math for math's sake.'' It creates a \textbf{meta-map of GR}, rigorously identifying where the theory relies on abstract principles like AC or LEM. This clarifies limits of computability, predictability, and physical realizability.

Structural Certification provides a rigorous, machine-checked framework achieving the project's primary goals while acknowledging the pragmatic use of classical tools. A formal constructive certificate remains a valuable but separate long-term goal.

The methodology is sound when articulated with precision and demonstrated concretely, as we have done with the Schwarzschild vacuum verification.

\end{document}