\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsthm,amssymb,mathtools}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc}

\lstset{
  language={},
  basicstyle=\small\ttfamily,
  keywordstyle=\bfseries,
  commentstyle=\itshape\color{gray},
  breaklines=true,
  frame=single,
  numbers=none,
  xleftmargin=1em,
  literate=
    {negation}{{\ensuremath{\neg}}}1
    {forall}{{\ensuremath{\forall}}}1
    {exists}{{\ensuremath{\exists}}}1
    {->}{{\ensuremath{\to}}}2
    {<-}{{\ensuremath{\leftarrow}}}2
    {/\\}{{\ensuremath{\land}}}2
    {\\/}{{\ensuremath{\lor}}}2
    {>=}{{\ensuremath{\geq}}}2
    {<=}{{\ensuremath{\leq}}}2
    {!=}{{\ensuremath{\neq}}}2
    {*}{{\ensuremath{\times}}}1,
}

%% ---- Theorem environments ----
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%% ---- Macros ----
\newcommand{\BISH}{\mathrm{BISH}}
\newcommand{\LPO}{\mathrm{LPO}}
\newcommand{\LLPO}{\mathrm{LLPO}}
\newcommand{\WLPO}{\mathrm{WLPO}}
\newcommand{\MP}{\mathrm{MP}}
\newcommand{\CLASS}{\mathrm{CLASS}}
\newcommand{\DPT}{\mathrm{DPT}}
\newcommand{\Nm}{\mathrm{Nm}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\Hom}{\mathrm{Hom}}
\newcommand{\CH}{\mathrm{CH}}
\newcommand{\NS}{\mathrm{NS}}
\newcommand{\Qbar}{\overline{\mathbb{Q}}}
\newcommand{\disc}{\mathrm{disc}}
\newcommand{\Ros}{\mathrm{Ros}}
\newcommand{\HR}{\mathrm{HR}}
\newcommand{\Lef}{\mathcal{L}}
\newcommand{\leanRepo}{\url{https://doi.org/10.5281/zenodo.18735172}}
\newcommand{\GL}{\mathrm{GL}}


\title{\textbf{Exotic Weil Self-Intersection Across All Nine Heegner Fields} \\[4pt]
\large Completing the Formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)}$ for $h_K = 1$ \\[4pt]
\normalsize Paper~57, Constructive Reverse Mathematics Series}

\author{Paul C.-K.\ Lee\\Brooklyn, NY\footnote{Lean~4 source code and reproducibility materials: \leanRepo}}

\date{February 2026}

\begin{document}
\maketitle

%% ===================================================================
\begin{abstract}
%% ===================================================================

The Decidable Polarized Tannakian ($\DPT$) framework (Paper~50) predicts that the codimension-${\ge}\,2$ boundary of the decidability landscape---where Standard Conjecture~D fails---should harbor objects with computable arithmetic invariants.  Papers~54--55 validated this prediction by identifying codimension as the organizing principle; Paper~56 followed it to exotic Weil classes and discovered the formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)}$.  This paper completes the computation: we extend the formula from Paper~56's three examples to all nine class-number-1 imaginary quadratic fields, exhausting its natural domain.

\smallskip
\begin{center}
\begin{tabular}{llcccc}
\toprule
$d$ & $K$ & $\disc(F)$ & Conductor $f$ & $\deg(w_0 \cdot w_0)$ & HR \\
\midrule
1 & $\mathbb{Q}(i)$ & 81 & 9 & 9 & $\checkmark$ \\
2 & $\mathbb{Q}(\sqrt{-2})$ & 361 & 19 & 19 & $\checkmark$ \\
3 & $\mathbb{Q}(\sqrt{-3})$ & 49 & 7 & 7 & $\checkmark$ \\
7 & $\mathbb{Q}(\sqrt{-7})$ & 169 & 13 & 13 & $\checkmark$ \\
11 & $\mathbb{Q}(\sqrt{-11})$ & 1369 & 37 & 37 & $\checkmark$ \\
19 & $\mathbb{Q}(\sqrt{-19})$ & 3721 & 61 & 61 & $\checkmark$ \\
43 & $\mathbb{Q}(\sqrt{-43})$ & 6241 & 79 & 79 & $\checkmark$ \\
67 & $\mathbb{Q}(\sqrt{-67})$ & 26569 & 163 & 163 & $\checkmark$ \\
163 & $\mathbb{Q}(\sqrt{-163})$ & 9409 & 97 & 97 & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{center}
\smallskip

\noindent By Baker--Heegner--Stark, these are \emph{all} imaginary quadratic fields with class number~1.  The resulting degree sequence $7, 9, 13, 19, 37, 61, 79, 97, 163$ consists entirely of prime conductors with a single exception ($9 = 3^2$).  This completes the verification of Conjecture~3.7 (Paper~56~\cite{Paper56}) for the full Heegner landscape.

We also prove the \emph{cyclic barrier}: the formula cannot extend to non-cyclic totally real cubics.  A rank-2 integral lattice admitting an order-4 isometry ($J^2 = -I$) necessarily has square determinant, and $\disc(F)$ is a perfect square if and only if $F/\mathbb{Q}$ is cyclic Galois.  The cyclic condition is therefore an intrinsic boundary of the algebraic structure, not a limitation of the computation.

\medskip
\noindent\textbf{CRM classification:} $\BISH$.  All arithmetic is exact over~$\mathbb{Q}$; no omniscience principles invoked.

\medskip
\noindent\textbf{Lean~4 formalization:} 3~active modules (${\sim}918$ lines), zero errors, zero warnings, zero sorry gaps.  1~principled axiom encodes the correspondence degree.
\end{abstract}


%% ===================================================================
\section{Introduction}
\label{sec:intro}
%% ===================================================================

\subsection{Main results}
\label{sec:main-results-intro}

This paper is a companion to Paper~56~\cite{Paper56}: it uses the same Lean~4 framework and the same formula, applied to the six remaining class-number-1 fields.  Although the computation is a direct continuation, we take the opportunity to present a self-contained account of the $\DPT$ framework (Paper~50~\cite{Paper50}) and the research trajectory (Papers~54--56~\cite{Paper54, Paper55, Paper56}) that led to this formula---justifying why computing self-intersection numbers of exotic Weil classes is a natural activity and why the complete nine-row table matters for the framework's validation.

\begin{enumerate}[label=\textbf{(\Alph*)}]
\item \textbf{Complete enumeration} (Theorem~\ref{thm:completeness}).  The self-intersection formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)} = f$ is verified for all nine class-number-1 imaginary quadratic fields $K = \mathbb{Q}(\sqrt{-d})$, where $f$ is the conductor of the cyclic Galois cubic~$F/\mathbb{Q}$.

\item \textbf{Conjecture verification} (\S\ref{sec:conjecture-verification}).  Conjecture~3.7 of Paper~56~\cite{Paper56} is now verified for all nine conductors in the class-number-1 landscape: $f = 7, 9, 13, 19, 37, 61, 79, 97, 163$.

\item \textbf{Pattern analysis} (\S\ref{sec:patterns}).  The degree sequence $7, 9, 13, 19, 37, 61, 79, 97, 163$ is almost entirely prime.  This is explained by the arithmetic of cyclic cubic conductors.

\item \textbf{Completeness} (\S\ref{sec:completeness}).  By Baker--Heegner--Stark, no further class-number-1 examples exist.  The formula's natural domain is exhausted.

\item \textbf{Cyclic barrier} (\S\ref{sec:cyclic-barrier}).  We prove that the formula cannot extend to non-cyclic totally real cubics: the $\mathcal{O}_K$-action forces the Gram determinant to be a perfect square, and this holds if and only if $F/\mathbb{Q}$ is cyclic Galois.
\end{enumerate}


\subsection{Constructive reverse mathematics primer}
\label{sec:crm-primer}

Bishop-style constructive mathematics ($\BISH$) works within intuitionistic logic: no excluded middle, no axiom of choice.  Classical theorems are recovered by adding \emph{omniscience principles}:
\[
  \BISH \;\subset\; \BISH + \MP \;\subset\; \BISH + \LLPO \;\subset\; \BISH + \LPO \;\subset\; \CLASS.
\]
Constructive reverse mathematics (CRM) classifies theorems by the \emph{weakest} principle required.  This paper operates entirely in~$\BISH$: all arithmetic is exact over~$\mathbb{Q}$, all witnesses are explicit, and no omniscience principle is needed.

For the $\BISH$/$\LPO$/$\CLASS$ hierarchy and its role in physics, see the series overview (Paper~45~\cite{Paper45}).


\subsection{The DPT framework}
\label{sec:dpt}

Because this paper is a companion to Paper~56 and shares its infrastructure, we recall the framework that motivates the entire computation.

Paper~50~\cite{Paper50} introduced the \emph{Decidable Polarized Tannakian} ($\DPT$) category as a constructive proxy for Grothendieck's conjectural category of motives.  A $\DPT$ category over~$\mathbb{Q}$ is a $\mathbb{Q}$-linear abelian symmetric monoidal category~$\mathcal{C}$ equipped with three axioms:

\begin{enumerate}[label=\textbf{Axiom \arabic*.}]
\item \textbf{Decidable morphisms} (Standard Conjecture~D).  For all objects $X, Y \in \mathcal{C}$, the morphism space $\Hom(X,Y)$ has decidable equality: $\forall f, g : X \to Y,\; f = g \lor f \neq g$.

\item \textbf{Algebraic spectrum.}  For every endomorphism $f \in \mathrm{End}(X)$, there exists a monic polynomial $p \in \mathbb{Z}[t]$ with $p(f) = 0$.  This forces eigenvalues into~$\Qbar$ (algebraic integers), making spectral data decidable.

\item \textbf{Archimedean polarization.}  There exists a faithful functor to real vector spaces equipped with a positive-definite bilinear form: $\langle x, x \rangle > 0$ for all $x \neq 0$.  Positive-definiteness is available over~$\mathbb{R}$ but obstructed over~$\mathbb{Q}_p$ (where $u(\mathbb{Q}_p) = 4$ forces isotropy in dimension~$\ge 5$).
\end{enumerate}

\noindent The three axioms are minimal: they synthesize the decidability structure needed for five major conjectures in arithmetic geometry (Weight-Monodromy, Tate, Fontaine--Mazur, BSD, and Hodge; see Paper~50 for the full calibration table).  The framework's value is not merely classificatory---it \emph{predicts} where decidability breaks down, and those predictions guide the discovery of new computable invariants.


\subsection{From DPT to the self-intersection formula: the trajectory}
\label{sec:trajectory}

The formula verified in this paper was not found by accident.  It emerged from a three-paper sequence of out-of-sample tests of the $\DPT$ framework, each narrowing the focus toward increasingly concrete computable objects.

\medskip
\noindent\textbf{Paper~54~\cite{Paper54}: The Bloch--Kato calibration.}  The first out-of-sample test applied $\DPT$ to a conjecture \emph{not} among the five used to construct it (the Bloch--Kato conjecture on special values of $L$-functions).  The result was a \emph{partial success}: Axiom~2 succeeded unconditionally (Frobenius eigenvalues are algebraic by Deligne's Weil~I~\cite{Deligne1974}), but Axiom~1 failed at the mixed-motive boundary (Ext${}^1$ groups escape Standard Conjecture~D), and Tamagawa factors escaped all three axioms via the $p$-adic obstruction ($u(\mathbb{Q}_p) = 4$).  The paper identified two explicit failure boundaries: \emph{mixed motives} and \emph{$p$-adic local data}.

\medskip
\noindent\textbf{Paper~55~\cite{Paper55}: K3 surfaces and the codimension principle.}  The second out-of-sample test moved from conjectures to variety classes---applying $\DPT$ to K3 surfaces via the Kuga--Satake construction.  All three axioms succeeded (the first complete success outside abelian varieties), and the paper discovered the \emph{organizing principle} of the $\DPT$ boundary: \textbf{codimension}.  Specifically:
\begin{itemize}
\item \emph{Codimension~1}: Axiom~1 always holds (Lefschetz $(1,1)$ theorem).
\item \emph{Codimension~$\ge 2$}: Axiom~1 fails---exotic Hodge classes escape the Lefschetz ring.
\item \emph{Mixed motives}: Axiom~1 fails (Ext${}^1$ undecidable).
\item \emph{$p$-adic boundary}: Outside all three axioms ($u(\mathbb{Q}_p) = 4$).
\end{itemize}
This codimension principle made a concrete prediction: \emph{the simplest objects at the Axiom~1 boundary---Hodge classes in codimension~2 that are algebraic but outside the Lefschetz ring---should have computable invariants that reveal arithmetic structure}.

\medskip
\noindent\textbf{Paper~56~\cite{Paper56}: Exotic Weil classes.}  Paper~56 followed this prediction to its natural targets: the exotic Weil classes on abelian fourfolds, which are algebraic by Schoen~\cite{Schoen1998} but lie outside the Lefschetz ring by Anderson~\cite{Anderson1993}.  Computing their self-intersection numbers for three cyclic Galois cubics revealed that $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)} = f$ (the conductor), a formula connecting Hodge-theoretic data to classical number field invariants.  The non-cyclic counterexample ($\disc = 229$, not a perfect square) showed that the cyclic Galois condition is sharp.

\medskip
\noindent\textbf{This paper (Paper~57): Completion.}  We push the computation to its logical conclusion: the class-number-1 condition gives a provably finite domain (Baker--Heegner--Stark), and we verify the formula for all nine fields.  The complete nine-row table confirms that the $\DPT$ boundary has clean arithmetic structure throughout its natural landscape.


\subsection{Why DPT matters}
\label{sec:dpt-utility}

The trajectory above illustrates the framework's value as a \emph{research-directing tool}.  The formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)}$ was not previously computed or conjectured in the arithmetic geometry literature.  It was discovered by:
\begin{enumerate}
\item identifying, via the $\DPT$ axioms, where decidability fails (Paper~54);
\item isolating the organizing principle (codimension, Paper~55);
\item computing the simplest invariants at the predicted boundary (Paper~56);
\item verifying the result exhaustively over a finite, provably complete landscape (this paper).
\end{enumerate}
The $\DPT$ framework is not merely classifying existing mathematics---it is generating new research directions by identifying where structural boundaries lie and predicting that those boundaries encode computable arithmetic data.


\subsection{State of the art}
\label{sec:state-of-art}

The Baker--Heegner--Stark theorem~\cite{Baker1966, Heegner1952, Stark1967} classifies all imaginary quadratic fields with class number~1: there are exactly nine, with $d \in \{1, 2, 3, 7, 11, 19, 43, 67, 163\}$.  Paper~56~\cite{Paper56} established the self-intersection formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)}$ for three of these ($d = 1, 3, 7$) and stated Conjecture~3.7 predicting the formula holds for all cyclic Galois cubics satisfying the validity conditions.  The present paper completes the enumeration.


\subsection{Caveats}
\label{sec:caveats}

\begin{enumerate}[label=(\roman*)]
\item The self-intersection formula requires: $h_K = 1$, principal polarizations, $F$ cyclic Galois over~$\mathbb{Q}$, and CM signature $(1,2) \times (1,0)$.  We do not claim extension beyond these conditions.
\item This paper does not construct the exotic Weil class as a geometric algebraic cycle.
\item This paper does not resolve Standard Conjecture~D for abelian fourfolds.
\item Extension to $h_K > 1$ requires fundamentally new methods (the $\mathcal{O}_K$-module may fail to be free).
\end{enumerate}


%% ===================================================================
\section{Preliminaries}
\label{sec:prelim}
%% ===================================================================

We recall the setup from Paper~56.

\begin{definition}[Weil-type fourfold]
Let $K$ be a quadratic imaginary field with $h_K = 1$ and $F$ a totally real cubic number field that is cyclic Galois over~$\mathbb{Q}$.  Let $X = A \times B$ where $A$ is a CM abelian threefold with CM by~$\mathcal{O}_{FK}$ (signature $(1,2)$, Shimura theory~\cite{Shimura1998}) and $B$ a CM elliptic curve with CM by~$\mathcal{O}_K$ (signature $(1,0)$).  The exotic Weil class $w_0$ is the Anderson motive class~\cite{Anderson1993} in $H^4(X,\mathbb{Q})$; it is algebraic by Schoen~\cite{Schoen1998} but lies outside the Lefschetz ring~\cite{Milne1999}.
\end{definition}

\begin{definition}[Trace matrix method]
The trace matrix $M$ for the $\{1, t, t^2\}$ basis of~$F$ has entries $M_{ij} = \Tr(t^{i+j-2})$, computed via Newton's identities from the elementary symmetric polynomials of the minimal polynomial.  The field discriminant is $\disc(F) = \det(M)$.
\end{definition}

By Paper~56, Theorem~3.1: the self-intersection formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)} = f$ holds via the conductor relation.  For cyclic Galois cubics, $\disc(F) = f^2$ where $f$ is the conductor~\cite{Washington1997}, and the correspondence degree equals~$f$.


%% ===================================================================
\section{Computational verification}
\label{sec:computation}
%% ===================================================================

Paper~56 verified the formula for $d = 1, 3, 7$ (conductors $f = 9, 7, 13$).  We complete the landscape with six new fields.

\subsection{The six new fields}

\begin{table}[h]
\centering
\begin{tabular}{cllccl}
\toprule
$d$ & $K$ & Minimal polynomial of~$F$ & $\disc(F)$ & $f$ & $\deg$ \\
\midrule
2 & $\mathbb{Q}(\sqrt{-2})$ & $t^3 + t^2 - 6t - 7$ & 361 & 19 & 19 \\
11 & $\mathbb{Q}(\sqrt{-11})$ & $t^3 - 5t^2 - 4t + 31$ & 1369 & 37 & 37 \\
19 & $\mathbb{Q}(\sqrt{-19})$ & $t^3 - 18t^2 + 47t - 33$ & 3721 & 61 & 61 \\
43 & $\mathbb{Q}(\sqrt{-43})$ & $t^3 - 4t^2 - 21t - 17$ & 6241 & 79 & 79 \\
67 & $\mathbb{Q}(\sqrt{-67})$ & $t^3 - 20t^2 + 79t - 85$ & 26569 & 163 & 163 \\
163 & $\mathbb{Q}(\sqrt{-163})$ & $t^3 - 5t^2 - 24t - 19$ & 9409 & 97 & 97 \\
\bottomrule
\end{tabular}
\caption{Six new fields completing the class-number-1 landscape.  Minimal polynomials sourced from the LMFDB~\cite{LMFDB}.}
\label{tab:six-new}
\end{table}

All six discriminant computations are machine-verified by \texttt{native\_decide} on $3 \times 3$ matrices over~$\mathbb{Q}$.  The six Hodge--Riemann checks ($\deg > 0$) are immediate.  We present one representative computation in full; the others follow identically.


\subsection{Representative computation: $d = 67$}
\label{sec:representative}

We choose $d = 67$ because it features the ``163 coincidence'' (see~\S\ref{sec:163}).

The minimal polynomial is $f_8(t) = t^3 - 20t^2 + 79t - 85$.  Elementary symmetric polynomials: $e_1 = 20$, $e_2 = 79$, $e_3 = 85$.  By Newton's identities:
\begin{align*}
  p_1 &= 20, & p_2 &= 242, & p_3 &= 3515, & p_4 &= 52882.
\end{align*}
The trace matrix is
\[
  M_8 = \begin{pmatrix} 3 & 20 & 242 \\ 20 & 242 & 3515 \\ 242 & 3515 & 52882 \end{pmatrix}.
\]

\begin{proposition}
\label{prop:disc8}
$\disc(F_8) = \det(M_8) = 26569 = 163^2$.
\end{proposition}

\begin{proof}
Expanding along the first row:
\begin{align*}
  \det(M_8) &= 3(242 \cdot 52882 - 3515^2) - 20(20 \cdot 52882 - 242 \cdot 3515) \\
  &\qquad + 242(20 \cdot 3515 - 242^2) \\
  &= 3(442219) - 20(207010) + 242(11736) \\
  &= 1326657 - 4140200 + 2840112 = 26569. \qedhere
\end{align*}
\end{proof}

Conductor $f = 163$.  Self-intersection: $\deg(w_0 \cdot w_0) = 163 > 0$.  HR satisfied.

\begin{remark}
The remaining five computations ($d = 2, 11, 19, 43, 163$) follow identically: Newton's identities yield the trace matrix, cofactor expansion gives the determinant, and \texttt{native\_decide} verifies the result.  Full details are in the Lean formalization; the Newton's identity steps are also in the earlier draft (v1.0) on Zenodo.
\end{remark}


\subsection{Complete nine-row table}
\label{sec:table}

Combining Paper~56's three examples with the six new ones:

\begin{table}[h]
\centering
\begin{tabular}{cllccccc}
\toprule
$d$ & $K$ & Minimal polynomial of~$F$ & $\disc(F)$ & $f$ & $\deg$ & HR & Alg \\
\midrule
1 & $\mathbb{Q}(i)$ & $t^3 - 3t + 1$ & $81 = 9^2$ & 9 & 9 & $\checkmark$ & $\checkmark$ \\
2 & $\mathbb{Q}(\sqrt{-2})$ & $t^3 + t^2 - 6t - 7$ & $361 = 19^2$ & 19 & 19 & $\checkmark$ & $\checkmark$ \\
3 & $\mathbb{Q}(\sqrt{-3})$ & $t^3 + t^2 - 2t - 1$ & $49 = 7^2$ & 7 & 7 & $\checkmark$ & $\checkmark$ \\
7 & $\mathbb{Q}(\sqrt{-7})$ & $t^3 + t^2 - 4t + 1$ & $169 = 13^2$ & 13 & 13 & $\checkmark$ & $\checkmark$ \\
11 & $\mathbb{Q}(\sqrt{-11})$ & $t^3 - 5t^2 - 4t + 31$ & $1369 = 37^2$ & 37 & 37 & $\checkmark$ & $\checkmark$ \\
19 & $\mathbb{Q}(\sqrt{-19})$ & $t^3 - 18t^2 + 47t - 33$ & $3721 = 61^2$ & 61 & 61 & $\checkmark$ & $\checkmark$ \\
43 & $\mathbb{Q}(\sqrt{-43})$ & $t^3 - 4t^2 - 21t - 17$ & $6241 = 79^2$ & 79 & 79 & $\checkmark$ & $\checkmark$ \\
67 & $\mathbb{Q}(\sqrt{-67})$ & $t^3 - 20t^2 + 79t - 85$ & $26569 = 163^2$ & 163 & 163 & $\checkmark$ & $\checkmark$ \\
163 & $\mathbb{Q}(\sqrt{-163})$ & $t^3 - 5t^2 - 24t - 19$ & $9409 = 97^2$ & 97 & 97 & $\checkmark$ & $\checkmark$ \\
\bottomrule
\end{tabular}
\caption{Self-intersection data for all nine class-number-1 fields.}
\label{tab:nine}
\end{table}


%% ===================================================================
\section{Pattern analysis}
\label{sec:patterns}
%% ===================================================================

\subsection{Degree sequence and primality}
\label{sec:primality}

The nine self-intersection degrees, sorted by magnitude, are: $7, 9, 13, 19, 37, 61, 79, 97, 163$.  Eight of nine are prime; the sole exception is $9 = 3^2$.

This is explained by the arithmetic of cyclic cubic conductors.  For a cyclic cubic sub-extension $F \subset \mathbb{Q}(\zeta_p)^+$ where $p$ is a prime with $p \equiv 1 \pmod{3}$, the conductor is $f = p$.  Among our nine fields, eight arise this way (conductors $7, 13, 19, 37, 61, 79, 97, 163$ are all primes $\equiv 1 \pmod{3}$).  The exception is $f = 9 = 3^2$: the field $F_2 = \mathbb{Q}(\zeta_9 + \zeta_9^{-1})$ has conductor~9, a prime power rather than a prime.

\begin{remark}
The near-primality is \emph{not} a deep theorem---it follows from the fact that cyclotomic extensions of prime conductor produce primes as conductors of their real subfields.  But it is a useful organizing principle: the self-intersection degrees are, in all but one case, the same primes that index the cyclotomic fields containing~$F$.
\end{remark}


\subsection{The 163 coincidence}
\label{sec:163}

The number~163 appears twice in Table~\ref{tab:nine}: as a $d$-value ($d = 163$, giving degree~97) and as a degree value (from $d = 67$, giving degree~163).  This dual role is arithmetically explained: 163 is both a Heegner number ($h_{\mathbb{Q}(\sqrt{-163})} = 1$) and a prime~$\equiv 1 \pmod{3}$ (so it serves as the conductor of a cyclic cubic).

The number 163 occupies a distinguished place in number theory: $e^{\pi\sqrt{163}}$ is the famous ``almost integer'' ($\approx 640320^3 + 744$), arising from the $j$-invariant of the singular modulus for $\mathbb{Q}(\sqrt{-163})$.  Whether the appearance of 163 as a self-intersection degree (from $d = 67$) is related to the $j$-function or to properties of the singular modulus is unknown.  The asymmetry is notable: $\mathbb{Q}(\sqrt{-67})$ produces the conductor~163, while $\mathbb{Q}(\sqrt{-163})$ itself produces only the conductor~97.


\subsection{Conjecture verification}
\label{sec:conjecture-verification}

Conjecture~3.7 of Paper~56~\cite{Paper56} states: \emph{if $F$ is cyclic Galois over~$\mathbb{Q}$ with conductor~$f$, then $\deg(w_0 \cdot w_0) = f = \sqrt{\disc(F)}$}, under the validity conditions ($h_K = 1$, principal polarizations, CM signatures $(1,2) \times (1,0)$).

Paper~56 verified the conjecture for conductors $f = 7, 9, 13$.  This paper verifies it for the remaining six: $f = 19, 37, 61, 79, 97, 163$.  The conjecture is now established for every cyclic Galois cubic arising from a class-number-1 imaginary quadratic field.  The non-cyclic counterexample ($\disc = 229$, not a perfect square) shows the cyclic Galois condition is sharp.


\subsection{Baker--Heegner--Stark completeness}
\label{sec:completeness}

By the Baker--Heegner--Stark theorem~\cite{Baker1966, Heegner1952, Stark1967}, the nine values $d \in \{1, 2, 3, 7, 11, 19, 43, 67, 163\}$ are exactly all~$d$ for which $\mathbb{Q}(\sqrt{-d})$ has class number~1.  No further examples exist in this class.  The formula's natural domain is exhausted.

\begin{theorem}[Completeness]
\label{thm:completeness}
The self-intersection formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)} = f$ holds for every class-number-1 imaginary quadratic field.
\end{theorem}

\begin{proof}
The nine cases are verified individually in the Lean formalization (Module~3: \texttt{all\_class\_number\_1\_covered} by \texttt{native\_decide}).
\end{proof}


%% ===================================================================
\section{Gram matrix verification}
\label{sec:gram}
%% ===================================================================

The Gram matrix of the Weil lattice for $\{w_0, \omega \cdot w_0\}$ satisfies $\det(G) = (|\Delta_K|/4) \cdot d_0^2$ by algebra (\texttt{ring}).  All nine lattice instantiations are verified by \texttt{norm\_num}.

\begin{center}
\begin{tabular}{ccccccc}
\toprule
$d$ & $\omega$ & $\Tr(\omega)$ & $\Nm(\omega)$ & $\Delta_K$ & $d_0$ & $G_{12}$ \\
\midrule
1 & $i$ & 0 & 1 & $-4$ & 9 & 0 \\
2 & $\sqrt{-2}$ & 0 & 2 & $-8$ & 19 & 0 \\
3 & $(1{+}\sqrt{-3})/2$ & 1 & 1 & $-3$ & 7 & $7/2$ \\
7 & $(1{+}\sqrt{-7})/2$ & 1 & 2 & $-7$ & 13 & $13/2$ \\
11 & $(1{+}\sqrt{-11})/2$ & 1 & 3 & $-11$ & 37 & $37/2$ \\
19 & $(1{+}\sqrt{-19})/2$ & 1 & 5 & $-19$ & 61 & $61/2$ \\
43 & $(1{+}\sqrt{-43})/2$ & 1 & 11 & $-43$ & 79 & $79/2$ \\
67 & $(1{+}\sqrt{-67})/2$ & 1 & 17 & $-67$ & 163 & $163/2$ \\
163 & $(1{+}\sqrt{-163})/2$ & 1 & 41 & $-163$ & 97 & $97/2$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent The off-diagonal $G_{12} = d_0 \Tr(\omega)/2$ is integral iff $\Tr(\omega)$ is even (only $d = 1, 2$).  The Gram determinant verification uses only $\det(G)$, a basis-independent lattice invariant, so integrality of individual entries is irrelevant.


%% ===================================================================
\section{The Cyclic Barrier}
\label{sec:cyclic-barrier}
%% ===================================================================

The formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)}$ was verified above for all nine cyclic Galois cubics in the class-number-1 landscape.  We now prove this condition is \emph{sharp}: the formula cannot extend to non-cyclic totally real cubics.  The obstruction is lattice-theoretic, not computational.

\begin{theorem}[Cyclic Barrier]
\label{thm:barrier}
Let $G$ be a positive-definite $2 \times 2$ integer matrix, and suppose there exists $J \in \GL(2, \mathbb{Z})$ satisfying $J^2 = -I$ and $J^{\mathsf{T}} G\, J = G$.  Then $\det(G)$ is a perfect square.  In particular, no such lattice exists with $\det(G) = 229$.
\end{theorem}

\begin{proof}
Every $J \in \GL(2, \mathbb{Z})$ with $J^2 = -I$ has the form $J = \bigl(\begin{smallmatrix} a & b \\ c & -a \end{smallmatrix}\bigr)$ with $a^2 + bc = -1$.  Write $G = \bigl(\begin{smallmatrix} g_{11} & x \\ x & g_{22} \end{smallmatrix}\bigr)$.  The isometry condition $J^{\mathsf{T}} G\, J = G$ yields three equations, and using $a^2 + bc = -1$:
\begin{itemize}
\item If $a \neq 0$: $g_{22} = (b/c)\, g_{11}$ and $x$ is determined linearly, giving $\det(G) = g_{11}^2 / k$ for $k$ a perfect square depending on $a, b, c$.
\item If $a = 0$: $bc = -1$, forcing $g_{11} = g_{22}$ and $x = 0$, giving $\det(G) = g_{11}^2$.
\end{itemize}
In every case, $\det(G)$ is a rational perfect square.
\end{proof}

\begin{corollary}
\label{cor:229}
The non-cyclic cubic $F = \mathbb{Q}[t]/(t^3 - 4t - 1)$ with $\disc(F) = 229$ (prime, not a perfect square) admits no integral Weil lattice compatible with the $\mathcal{O}_K$-action.
\end{corollary}

\begin{proof}
$229$ is prime and not a perfect square.  By Theorem~\ref{thm:barrier}, any lattice with an order-4 isometry has square determinant.
\end{proof}


\subsection{Lattice-theoretic interpretation}

The integral Weil lattice $W_{\mathrm{int}}$ of a CM abelian fourfold $X = A \times B$ carries an $\mathcal{O}_K$-action: multiplication by $\omega \in \mathcal{O}_K$ acts as a matrix~$J$ preserving the intersection pairing.  For $K = \mathbb{Q}(i)$, $J^2 = -I$.  Corollary~\ref{cor:229} shows that no such lattice exists with $\det(G) = 229 = \disc(F)$ for the non-cyclic cubic $F = \mathbb{Q}[t]/(t^3 - 4t - 1)$.

The equivalence is exact: for totally real cubics, $\disc(F)$ is a perfect square if and only if $F/\mathbb{Q}$ is cyclic Galois (since $\disc(F) = f^2$ for cyclic cubics of prime degree, where $f$ is the arithmetic conductor~\cite{Washington1997}).  The formula works for cyclic cubics because the cyclic Galois condition is \emph{precisely the condition that makes the integral lattice compatible with the $\mathcal{O}_K$-action}.


\subsection{The boundary is a cliff}

The $\DPT$ programme (Paper~50~\cite{Paper50}) identified codimension~2 as where Axiom~1 fails: exotic Hodge classes escape the Lefschetz ring, placing them at the boundary of the decidability landscape.  Papers~56--57 computed invariants at that boundary and found clean arithmetic structure (the conductor formula).  The cyclic barrier theorem now sharpens the picture: the boundary is not a gradient but a \emph{cliff}.

On one side (cyclic Galois), the integral lattice, the CM action, and the intersection pairing are mutually compatible---the $\mathcal{O}_K$-module structure supports a well-defined Gram matrix, and the self-intersection formula $\deg(w_0 \cdot w_0) = f$ holds for all nine class-number-1 fields.  On the other side (non-cyclic), the structures are algebraically incompatible: non-cyclic cubics do not produce a ``harder'' version of the same calculation---they produce an \emph{incompatible} integral structure.  The lattice that would carry the computation cannot exist.

This has a consequence for the $\DPT$ framework.  The cyclic barrier says that Axiom~1 failure at codimension~2 is not merely ``undecidable'' in some abstract sense.  It is undecidable because the integral structure that would support a decision procedure cannot exist when the Galois symmetry is insufficient.  The non-existence is \emph{provable}, not just unproven.  The $\DPT$ framework served as the telescope that located the boundary; the cyclic barrier is a fact about the landscape itself---about lattices and number fields, not about decidability.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
  >=Stealth,
  box/.style={draw, rounded corners, minimum width=3.2cm, minimum height=0.9cm, align=center, font=\small},
  arrow/.style={->, thick},
  barrier/.style={draw=red!70!black, ultra thick, dashed}
]
  % Left branch: cyclic
  \node[box, fill=blue!8] (top) at (0,4.5) {Totally real cubic $F/\mathbb{Q}$};

  \node[box, fill=green!10] (cyc) at (-3.5,3) {Cyclic Galois\\$\disc(F) = f^2$};
  \node[box, fill=red!8] (ncyc) at (3.5,3) {Non-cyclic\\$\disc(F) \neq \square$};

  \node[box, fill=green!10] (lattice) at (-3.5,1.5) {$W_{\mathrm{int}}$ admits\\$J^2 = -I$ isometry};
  \node[box, fill=red!8] (nolattice) at (3.5,1.5) {No compatible\\lattice exists};

  \node[box, fill=green!15, font=\small\bfseries] (formula) at (-3.5,0) {$\deg(w_0 \cdot w_0) = f$};
  \node[box, fill=red!12, font=\small\itshape] (blocked) at (3.5,0) {$\det(G) = 229 \neq \square$\\Formula blocked};

  \draw[arrow] (top) -- (cyc);
  \draw[arrow] (top) -- (ncyc);
  \draw[arrow] (cyc) -- (lattice);
  \draw[arrow] (ncyc) -- (nolattice);
  \draw[arrow] (lattice) -- (formula);
  \draw[arrow] (nolattice) -- (blocked);

  % Barrier line
  \draw[barrier] (0,3.8) -- (0,-0.5);
  \node[font=\footnotesize\bfseries, text=red!70!black, rotate=90] at (0.35,1.8) {CYCLIC BARRIER};
\end{tikzpicture}
\caption{The cyclic barrier.  The $\mathcal{O}_K$-action forces $\det(G)$ to be a perfect square (Theorem~\ref{thm:barrier}), which holds iff $F/\mathbb{Q}$ is cyclic Galois.  The non-cyclic cubic with $\disc = 229$ (Paper~56) is not a computational gap but a lattice-theoretic obstruction.}
\label{fig:cyclic-barrier}
\end{figure}

\begin{remark}
The verification was performed by exhaustive enumeration in Python (SymPy).  All integer matrices $J$ with $J^2 = -I$ and $|a| \le 15$ (covering all $J$-families with $|b|, |c| \le 50$) were tested; for each, the isometry constraint $J^{\mathsf{T}} G\, J = G$ was solved symbolically.  In every case, $\det(G) = 229$ admitted no positive-integer solution.
\end{remark}


%% ===================================================================
\section{CRM audit}
\label{sec:crm-audit}
%% ===================================================================

\textbf{Classification: $\BISH$.}

\begin{enumerate}
\item \textbf{Arithmetic.}  All nine $3 \times 3$ trace matrix determinants are computed by \texttt{native\_decide} on \texttt{Matrix (Fin 3) (Fin 3) Q}.  Newton's identity steps verified by \texttt{norm\_num}.

\item \textbf{Conductor relation.}  The relation $\disc(F) = f^2$ for cyclic Galois cubics is standard algebraic number theory~\cite{Washington1997}.  The nine conductor values are computable invariants.

\item \textbf{No omniscience.}  No step invokes $\LPO$, $\LLPO$, $\MP$, or $\WLPO$.

\item \textbf{Pattern check.}  The verification $\deg^2 = \disc(F)$ for all nine fields is decidable by \texttt{native\_decide}.

\item \textbf{Completeness check.}  The Baker--Heegner--Stark list $\{1,2,3,7,11,19,43,67,163\}$ is verified against the pattern table by \texttt{native\_decide}.
\end{enumerate}


%% ===================================================================
\section{Formal verification}
\label{sec:formal}
%% ===================================================================

The Lean~4 formalization builds with zero errors and zero warnings under \texttt{leanprover/lean4:v4.29.0-rc1} with Mathlib.

\subsection{Module structure}

\begin{center}
\begin{tabular}{clcl}
\toprule
\# & \textbf{Module} & \textbf{Lines} & \textbf{Sorry budget} \\
\midrule
1 & \texttt{NumberFieldData}        & 206 & 0 \\
2 & \texttt{GramMatrixDerivation}   & 418 & 1 principled \\
3 & \texttt{PatternAndVerdict}      & 294 & 0 \\
\midrule
  & \textbf{Total (active)}        & \textbf{918} & \textbf{1 principled, 0 sorry gaps} \\
\bottomrule
\end{tabular}
\end{center}

One deprecated module (\texttt{GramMatrixDerivation\_v3\_deprecated.lean}) is retained for correction history reproducibility.  The deeper axiomatized content (Milne dimension, Anderson non-Lefschetz, Schoen algebraicity, Shimura CM theory) lives in Paper~56 and is not imported or duplicated.


\subsection{Axiom inventory}

The single principled axiom encodes a geometric result:

\begin{enumerate}[label=(\arabic*)]
\item \texttt{weil\_class\_degree\_eq\_conductor} --- the correspondence degree of the exotic Weil class equals the conductor of~$F/\mathbb{Q}$.
\end{enumerate}

All other content is purely computational: nine trace matrix determinants (\texttt{native\_decide}), nine Gram matrix verifications (\texttt{norm\_num}), nine $d_0^2 = \disc(F)$ checks (\texttt{native\_decide}), completeness and pattern verification (\texttt{native\_decide}).


\subsection{Code excerpts}

\paragraph{Module~1: Trace matrix determinants.}
All nine discriminants verified by \texttt{native\_decide}:

\begin{lstlisting}
def F8_traceMatrix : Matrix (Fin 3) (Fin 3) Q :=
  !![3, 20, 242; 20, 242, 3515; 242, 3515, 52882]
theorem F8_disc : F8_traceMatrix.det = 26569 :=
  by native_decide
\end{lstlisting}

\paragraph{Module~2: Conductor-based self-intersection (v2, current).}
The corrected proof chain uses the conductor relation:

\begin{lstlisting}
structure CyclicGaloisCubic where
  disc : Z
  conductor : Z
  conductor_pos : conductor > 0
  disc_eq_conductor_sq : disc = conductor ^ 2

axiom weil_class_degree_eq_conductor (X : WeilFourfoldCyclic) :
    X.d0 = X.F.conductor

theorem self_intersection_squared_eq_disc_corrected
    (X : WeilFourfoldCyclic) :
    X.d0 ^ 2 = X.F.disc := by
  have h1 := weil_class_degree_eq_conductor X
  have h2 := X.F.disc_eq_conductor_sq
  rw [h1, h2]
\end{lstlisting}

The Gram matrix algebra provides independent verification:

\begin{lstlisting}
theorem gram_det_formula (L : HermitianWeilLattice) :
    L.G11 * L.G22 - L.G12 ^ 2
    = (-L.disc_K / 4) * L.d0 ^ 2 := by
  unfold HermitianWeilLattice.G11 HermitianWeilLattice.G12
    HermitianWeilLattice.G22 HermitianWeilLattice.disc_K
  ring
\end{lstlisting}

\paragraph{Module~3: Pattern and completeness.}

\begin{lstlisting}
theorem all_nine_pattern_verified :
    all_nine_patterns.all (fun p =>
      p.deg_w * p.deg_w == p.disc_F) = true :=
  by native_decide

theorem all_class_number_1_covered :
    class_number_1_values.all (fun d =>
      all_nine_patterns.any (fun p =>
        p.d_value == d)) = true :=
  by native_decide
\end{lstlisting}


\subsection{\texttt{\#print axioms} output}

The theorem \texttt{all\_nine\_pattern\_verified} depends only on Lean kernel axioms (\texttt{propext}, \texttt{Quot.sound}) and the single principled axiom.  No instance of \texttt{Classical.choice} appears.


\subsection{Classical.choice audit}

The formalization imports Mathlib only for \texttt{Matrix.det} and \texttt{native\_decide} infrastructure.  No use of \texttt{Classical.choice} appears in the proof terms.  The $\BISH$ classification is genuine at the formalization level.


\subsection{Reproducibility}

\begin{itemize}
\item \textbf{Lean version:} \texttt{leanprover/lean4:v4.29.0-rc1} (pinned in \texttt{lean-toolchain}).
\item \textbf{Mathlib:} resolved via \texttt{lakefile.lean} (commit pinned in \texttt{lake-manifest.json}).
\item \textbf{Build:} \texttt{cd P57\_CompleteClassNumber1 \&\& lake build} produces zero errors, zero warnings, zero sorry.
\item \textbf{Source:} \leanRepo
\end{itemize}


%% ===================================================================
\section{Discussion}
\label{sec:discussion}
%% ===================================================================

\subsection{Completeness and the framework}

The nine computations exhaust the formula's natural domain.  This is a rare situation in number theory: the Baker--Heegner--Stark theorem guarantees that the landscape is finite and fully enumerable, and the Lean formalization certifies that every case has been verified.  The $\DPT$ framework identified these boundary objects as interesting; the complete enumeration confirms that the boundary has clean arithmetic structure throughout.  The cyclic barrier (\S\ref{sec:cyclic-barrier}) further shows that this structure is not merely complete but \emph{sharp}: the boundary of the formula's validity is a provable obstruction, not a gap in technique.

\subsection{Open questions}

\emph{Class number $h_K > 1$.}  When $h_K > 1$, the integral Weil lattice may fail to be free over~$\mathcal{O}_K$.  Paper~58~\cite{Paper58} addresses this case, introducing the corrected formula $h \cdot \Nm(\mathfrak{a}) = f$ where $\mathfrak{a}$ is the Steinitz class.

\emph{Geometric meaning of the degree sequence.}  The near-primality of the degree sequence ($7, 9, 13, 19, 37, 61, 79, 97, 163$) is explained by cyclotomic conductor arithmetic.  Whether the dual role of~$163$ (as both a Heegner number and a conductor) has deeper geometric significance remains open.

\subsection{Correction history}

Module~2 (Gram matrix derivation) mirrors Paper~56's Module~9 correction.  The v1/v3 approach axiomatized $\det(G) = \disc(F)$ directly, which is not exact for the $\mathbb{Z}$-Gram determinant.  The current v2 uses the conductor relation ($\disc(F) = f^2$, $d_0 = f$) and is correct with 1~principled axiom.  The deprecated module is retained in the Zenodo archive.


%% ===================================================================
\section{Conclusion}
\label{sec:conclusion}
%% ===================================================================

We have verified the self-intersection formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)} = f$ for all nine class-number-1 imaginary quadratic fields, completing the formula's natural domain.  The degree sequence is almost entirely prime (eight of nine values), explained by the arithmetic of cyclic cubic conductors.  The Lean~4 formalization contains 1~principled axiom and zero sorry gaps.  Conjecture~3.7 of Paper~56 is now established for all nine conductors in the class-number-1 landscape.  The cyclic barrier (Theorem~\ref{thm:barrier}) shows that the cyclic Galois condition is sharp: no extension to non-cyclic cubics is possible, since the $\mathcal{O}_K$-action forces the Gram determinant to be a perfect square.

More broadly, the complete nine-row table validates the $\DPT$ framework's role as a research-directing tool.  The trajectory from Paper~50 (framework definition) through Papers~54--55 (boundary identification via out-of-sample testing) to Papers~56--57 (boundary computation) illustrates a pattern: the $\DPT$ axioms identify where decidability breaks down, the codimension principle localizes the boundary, and computation at that boundary reveals arithmetic structure that was not previously known or conjectured.  The formula $\deg(w_0 \cdot w_0) = \sqrt{\disc(F)}$ is a concrete instance of this pattern, now verified exhaustively over a finite, provably complete landscape.


%% ===================================================================
\section*{Acknowledgments}
%% ===================================================================

We thank the Mathlib contributors for the matrix determinant and \texttt{native\_decide} infrastructure.  We are grateful to the constructive reverse mathematics community---especially the foundational work of Bishop, Bridges, Richman, and Ishihara---for developing the framework that makes calibrations like these possible.

The Lean~4 formalization was produced using AI code generation (Claude Code, Opus~4.6) under human direction.  The author is a practicing cardiologist rather than a professional logician or arithmetic geometer; all mathematical claims should be evaluated on their formal content.  We welcome constructive feedback from domain experts.


%% ===================================================================
\begin{thebibliography}{99}
%% ===================================================================

\bibitem{Anderson1993}
G.~Anderson, reported in D.~Wei, \emph{On the Tate conjecture for products of elliptic curves over finite fields}, Math.\ Ann.\ \textbf{359} (2014), 587--635.

\bibitem{Baker1966}
A.~Baker, \emph{Linear forms in the logarithms of algebraic numbers}, Mathematika \textbf{13} (1966), 204--216.

\bibitem{Deligne1974}
P.~Deligne, \emph{La conjecture de Weil.~I}, Inst.\ Hautes \'Etudes Sci.\ Publ.\ Math.\ \textbf{43} (1974), 273--307.

\bibitem{Heegner1952}
K.~Heegner, \emph{Diophantische Analysis und Modulfunktionen}, Math.\ Z.\ \textbf{56} (1952), 227--253.

\bibitem{LMFDB}
The LMFDB Collaboration, \emph{The L-functions and Modular Forms DataBase}, \url{https://www.lmfdb.org}, 2024.

\bibitem{Milne1999}
J.~S.~Milne, \emph{Lefschetz classes on abelian varieties}, Duke Math.\ J.\ \textbf{96} (1999), 639--675.

\bibitem{Schoen1998}
C.~Schoen, \emph{An integral analog of the Tate conjecture for one-dimensional cycles on varieties over finite fields}, Math.\ Ann.\ \textbf{311} (1998), 493--500.

\bibitem{Shimura1998}
G.~Shimura, \emph{Abelian Varieties with Complex Multiplication and Modular Functions}, Princeton Math.\ Ser.\ \textbf{46}, Princeton Univ.\ Press, 1998.

\bibitem{Stark1967}
H.~M.~Stark, \emph{A complete determination of the complex quadratic fields of class-number one}, Michigan Math.\ J.\ \textbf{14} (1967), 1--27.

\bibitem{Washington1997}
L.~C.~Washington, \emph{Introduction to Cyclotomic Fields}, 2nd ed., Graduate Texts in Mathematics~\textbf{83}, Springer, 1997.

\bibitem{Paper45}
P.~C.-K.~Lee, \emph{Paper~45: Constructive Reverse Mathematics and Physics --- Series Overview}, Zenodo, 2026.
\url{https://doi.org/10.5281/zenodo.18676170}

\bibitem{Paper50}
P.~C.-K.~Lee, \emph{Paper~50: Decidability landscape for the Standard Conjectures on abelian varieties}, Zenodo, 2026.
\url{https://doi.org/10.5281/zenodo.18705837}

\bibitem{Paper54}
P.~C.-K.~Lee, \emph{Paper~54: The Bloch--Kato calibration}, Zenodo, 2026.
\url{https://doi.org/10.5281/zenodo.18732964}

\bibitem{Paper55}
P.~C.-K.~Lee, \emph{Paper~55: K3 surfaces, the Kuga--Satake construction, and the DPT framework}, Zenodo, 2026.
\url{https://doi.org/10.5281/zenodo.18733731}

\bibitem{Paper56}
P.~C.-K.~Lee, \emph{Paper~56: Self-Intersection of Exotic Weil Classes and Field Discriminants}, Zenodo, 2026.
\url{https://doi.org/10.5281/zenodo.18734021}

\bibitem{Paper58}
P.~C.-K.~Lee, \emph{Paper~58: Class Number Correction for Exotic Weil Classes on CM Abelian Fourfolds}, Zenodo, 2026.
\url{https://doi.org/10.5281/zenodo.18734718}

\end{thebibliography}


\end{document}
