% NOTE: Supporting report only. The main paper is paper-v6-020526.tex.
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\title{WLPO Constructive Completion Status (v3)}
\author{P2\_BidualGap Audit}
\date{February 5, 2026}

\begin{document}
\maketitle

\section*{Summary}
\begin{itemize}
  \item \textbf{Lean proof status}: \texttt{HB/WLPO\_to\_Gap\_HB.lean} remains complete (no \texttt{sorry}s). The direct dual witness \texttt{G} now uses the constructive $\sigma_\varepsilon$ coefficients, avoiding sign case splits.
  \item \textbf{WLPO usage}: norm attainment for \(\ell^1\) now calls \texttt{exists\_coeff\_near\_sup\_WLPO}; it uses WLPO to decide the ``all false vs not all false'' case, but still uses classical extraction of a witness index.
  \item \textbf{Build status}: full \texttt{lake build} succeeded. There are still \texttt{sorry} declarations in \texttt{Papers/P1\_GBC/RankOneToggle/Spectrum.lean}; P2 HB/Slim files are sorry-free (script pass).
  \item \textbf{Small folder}: added \texttt{LeanProof/} as a lightweight wrapper that re-exports \texttt{HB/WLPO\_to\_Gap\_HB.lean} and provides \texttt{LeanProof/All.lean}.
\end{itemize}

\section*{Lean Proof: What Changed}
\begin{itemize}
  \item \texttt{HB/DirectDual.lean}: replaced sign-vector coefficients with $\sigma_\varepsilon$; finite-sum bounds now use \(2\,\|f\|\). This keeps the direct witness constructive in spirit.
  \item \texttt{HB/WLPO\_DualBanach.lean}: replaced the classical \texttt{exists\_coeff\_near\_sup} with \texttt{exists\_coeff\_near\_sup\_WLPO}. WLPO decides the bound; a classical step still extracts a witness.
\end{itemize}

\section*{CRM Constructivity Status}
The HB path is \emph{improved but not fully CRM-constructive}. Remaining classical dependencies include:
\begin{itemize}
  \item witness extraction from \texttt{\textasciitilde forall} in \texttt{exists\_coeff\_near\_sup\_WLPO};
  \item summability via \texttt{summable\_of\_sum\_le} (tsum existence in Lean's classical reals);
  \item Hahn--Banach style dual isometries in \texttt{HB/DualIsometriesComplete.lean}.
\end{itemize}

\section*{Latest LaTeX Draft: Incomplete Places}
\textbf{Latest draft:} \texttt{documentation/paper-v5.tex} (newest mtime).
\begin{itemize}
  \item Line ~757: \texttt{sorry} in the WLPO-to-gap sketch.
  \item Line ~802: \texttt{sorry} in the algebra structure on \(\ell^\infty / c_0\).
  \item Lines ~817--820: \texttt{sorry} placeholders in the Stone window equivalence construction.
\end{itemize}
\textbf{Other drafts:} \texttt{paper-final.tex} still contains explicit \texttt{sorry} blocks (e.g., lines ~899, ~905).

\section*{Next Actions for Full CRM Compliance}
\begin{itemize}
  \item Replace classical witness extraction in \texttt{exists\_coeff\_near\_sup\_WLPO} with a constructive witness principle (or explicitly assume LPO/Markov).
  \item Replace classical summability lemmas with constructive Cauchy bounds where possible.
  \item Isolate Hahn--Banach dependencies in a classical-only namespace and keep the WLPO path free of them.
\end{itemize}

\end{document}
