% WLPO Completion Report for Paper 2 (P2_BidualGap)
\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,mathtools}
\usepackage[american]{babel}
\usepackage{enumitem}
\usepackage{hyperref}

\title{WLPO Completion Report (Paper 2)}
\author{Codex Audit}
\date{\today}

\begin{document}
\maketitle

\section*{Scope}
This report summarizes:
\begin{itemize}
  \item the latest LaTeX draft in \texttt{P2\_BidualGap/documentation},
  \item explicit incompletenesses in that draft, and
  \item the Lean-side status for the WLPO $\Rightarrow$ bidual gap direction.
\end{itemize}

\section*{Latest LaTeX Draft}
By modification time, the latest draft is:
\begin{itemize}
  \item \texttt{documentation/paper-v5.tex}
\end{itemize}

\section*{Incomplete Parts in the Latest Draft}
The following are explicit placeholders in \texttt{paper-v5.tex}:
\begin{itemize}
  \item \textbf{WLPO $\to$ Gap (coding-based sketch)}: the Lean listing \texttt{gap\_from\_WLPO} ends with \texttt{sorry}.
  \item \textbf{Stone window construction}: the algebra instance on $\ell^\infty/c_0$ and the Stone window isomorphism contain multiple \texttt{sorry} placeholders (well-defined multiplication, inverse map, and proof of inverses).
\end{itemize}
These are the only explicit \texttt{sorry}-style gaps in the latest LaTeX draft.

\section*{Lean Completion Status}
In \texttt{Papers/P2\_BidualGap/HB/WLPO\_to\_Gap\_HB.lean}, the two previous axioms
\texttt{dual\_is\_banach\_c0\_from\_WLPO} and
\texttt{dual\_is\_banach\_c0\_dual\_from\_WLPO}
were replaced by explicit proofs. The file builds.

\paragraph{Note on Constructivity.}
The completion currently uses classical operator-norm machinery (via $\sup$ on closed balls)
for the DualIsBanach obligations. This removes all Lean-level sorries/axioms in the WLPO $\Rightarrow$ gap proof, but it is not yet the fully WLPO-only constructive route envisioned in the CRM plan. A fully WLPO-based locatedness/supremum argument for $\ell^\infty$ remains the remaining conceptual gap if strict constructivity is required.

\section*{Recommended Next Steps (if strict WLPO-only is required)}
\begin{itemize}
  \item Implement a WLPO-based locatedness/approximate supremum lemma for bounded sequences.
  \item Use that to show norm\_located and norm\_attained for $(\ell^1)^*$.
  \item Rewire the WLPO $\Rightarrow$ gap proof to rely only on WLPO (no classical sup arguments).
\end{itemize}

\end{document}
