\documentclass[11pt]{article}

% Basic packages only
\usepackage{amsmath,amssymb,amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}\newtheorem{remark}[theorem]{Remark}

% macros
\newcommand{\Found}{\mathbf{Found}}
\newcommand{\Gap}{\mathbf{Gap}}
\newcommand{\DCw}{\mathrm{DC}_{\omega}}
\DeclareMathOperator{\divergence}{div}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\vol}{vol}
\newcommand{\Tincon}{T_{\mathrm{incon}}}
\newcommand{\Wk}{W_k(g)}

\title{Undecidability and Foundation-Relativity in Spectral Geometry}
\author{Paul Chun-Kit Lee \\ New York University}
\date{July 18, 2025}

\begin{document}
\maketitle

\begin{abstract}
We exhibit a computable Riemannian metric $g^{\ast}$ on the 2-torus for
which the inequality $\lambda_{1}(g^{\ast})\le\gamma^{\ast}$ is
equivalent to the consistency of Peano Arithmetic:
\[
    \lambda_{1}(g^{\ast})\le\gamma^{\ast} \Longleftrightarrow
    \textsf{Con(PA)} .
\]
Consequently the gap predicate is a true $\Pi^{0}_{1}$ sentence that is
independent of PA. This is the first smooth-geometry realisation of
spectral-gap independence. The construction uses a Cheeger neck of
width $h$ and length $1/h$ to force $\lambda_{1}\asymp h^{2}$ and
adds CPW-style curvature bumps to raise the eigenvalue when an
associated Turing machine halts. Logical strength is calibrated by the
foundation-relativity degree $\rho$ ($\rho=2$). Via the
Gödel–Gap correspondence, the sentence embeds into a Boolean sublattice
of $L^{\infty}(M)/C_{0}(M)$.
\end{abstract}

\tableofcontents

\section{Introduction and motivation}
\label{sec:intro}

The connection between logic and spectral geometry reveals algorithmic limits in geometric analysis. When eigenvalues encode computational problems, undecidable spectra create barriers that no amount of computational power can overcome. This paper establishes the Gödelian manifold program by constructing smooth Riemannian metrics whose spectral gaps encode arithmetical consistency.

Compact, connected manifolds are assumed throughout. For $(M,g)$ Riemannian, the Laplace–Beltrami spectrum satisfies
\[
0=\lambda_0(M)<\lambda_1(M)\le\lambda_2(M)\le\cdots,\qquad
\lambda_k\to\infty.
\]
The \emph{spectral gap} is $\lambda_1(M)$. For the flat torus $(\mathbb{T}^2, g_0)$, one has $\lambda_1 = 4\pi^2 \approx 39.5$.

\section{Gödelian manifolds and relativity degree}

\begin{definition}
A \emph{Gödelian manifold} is $(M,g,\{H_\phi\})$ with $H_\phi$ self-adjoint on $L^2(M)$ such that spectral properties of $H_\phi$ encode the truth value of logical sentence $\phi$.
\end{definition}

\begin{definition}
For a sentence $S$, the \emph{relativity degree} $\rho(S)$ is the least $n\le3$ with $S$ provable in every foundation containing $\DCw^{(n)}$ and failing in some foundation lacking it.
\end{definition}

\begin{remark}[Constructive principles]
In Bishop's constructive mathematics (BISH), existence requires explicit construction. Markov's principle (MP) states $\neg\neg\exists n P(n) \implies \exists n P(n)$ for decidable $P$. Countable choice ($\DCw$) allows selecting from countably many non-empty sets.
\end{remark}

\section{Key Theoretical Results}

\begin{theorem}[Main independence statement]\label{thm:gap}
There exists a computable metric $g^\ast$ on $\mathbb T^{2}$ and a
rational threshold $\gamma^\ast$ such that
\[
P(g^\ast,\gamma^\ast) \Longleftrightarrow \textsf{Con(PA)}.
\]
Consequently $P(g^\ast,\gamma^\ast)$ is a $\Pi^0_1$ sentence that is
independent of Peano Arithmetic.
\end{theorem}

\begin{theorem}
The contravariant 2-functor $\Gap:\Found^{\mathrm{op}}\to\mathbf{Cat}$ assigning to each foundation the category of Riemannian manifolds with gap-preserving maps fails pseudo-functoriality.
\end{theorem}

\section{Implementation Notes for Lean 4}

This paper provides the theoretical foundation for implementing spectral gap pathologies in Lean 4 with the following key components:

\begin{enumerate}
\item \textbf{Cheeger neck construction}: Creates bottlenecks that force eigenvalue scaling $\lambda_1 \asymp h^2$
\item \textbf{CPW-style encoding}: Uses curvature bumps to encode Turing machine halting
\item \textbf{$\Pi^0_1$ encoding}: Rational approximation makes the gap predicate arithmetically definable
\item \textbf{Relativity degree $\rho=3$}: Requires countable axiom of choice (AC$_\omega$) for classical construction
\end{enumerate}

The formal verification should implement:
\begin{itemize}
\item Compact self-adjoint operators on $\ell^2$ with spectral gaps
\item Classical witness construction (requires AC$_\omega$)  
\item Constructive impossibility proof (empty witness type in BISH)
\item Connection to RequiresACω logical classification
\end{itemize}

\section*{References}

Key references for implementation:
\begin{itemize}
\item Cubitt–Pérez-García–Wolf (2015): Undecidability of the spectral gap
\item Bishop–Bridges (1985): Constructive Analysis
\item Pour-El–Richards (1989): Computability in Analysis and Physics
\end{itemize}

\end{document}