\documentclass[12pt, a4paper]{article}

% ============================================================
% Packages
% ============================================================
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, calc}

% ============================================================
% Theorem Environments
% ============================================================
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ============================================================
% Lean 4 Listings
% ============================================================
\lstdefinelanguage{Lean4}{
  morekeywords={theorem, def, lemma, axiom, opaque, noncomputable,
    import, open, where, instance, structure, inductive, deriving,
    if, then, else, match, with, fun, let, have, show, by,
    exact, rfl, intro, cases, constructor, unfold, rw, simp,
    contradiction, absurd, decide, refine, Prop, Type, Bool,
    true, false, Nat, sorry},
  sensitive=true,
  morecomment=[l]{--},
  morecomment=[s]{/-}{-/},
  morestring=[b]",
  literate={→}{$\to$}1 {←}{$\leftarrow$}1 {↔}{$\leftrightarrow$}1
           {∧}{$\wedge$}1 {∨}{$\vee$}1 {¬}{$\neg$}1
           {≤}{$\le$}1 {≥}{$\ge$}1 {≠}{$\ne$}1
           {⟨}{$\langle$}1 {⟩}{$\rangle$}1
           {∀}{$\forall$}1 {∃}{$\exists$}1
           {ℕ}{$\mathbb{N}$}1 {ℤ}{$\mathbb{Z}$}1
           {ℚ}{$\mathbb{Q}$}1 {ℝ}{$\mathbb{R}$}1
}

\lstset{
  language=Lean4,
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries\color{blue!60!black},
  commentstyle=\itshape\color{green!40!black},
  stringstyle=\color{red!60!black},
  breaklines=true,
  columns=flexible,
  frame=single,
  framerule=0.4pt,
  xleftmargin=1em,
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=5pt
}

% ============================================================
% Notation
% ============================================================
\newcommand{\BISH}{\mathrm{BISH}}
\newcommand{\BISHMP}{\mathrm{BISH{+}MP}}
\newcommand{\LLPO}{\mathrm{LLPO}}
\newcommand{\WLPO}{\mathrm{WLPO}}
\newcommand{\LPO}{\mathrm{LPO}}
\newcommand{\CLASS}{\mathrm{CLASS}}
\newcommand{\CRM}{\mathrm{CRM}}
\newcommand{\DPT}{\mathrm{DPT}}
\newcommand{\Qp}{\mathbb{Q}_p}
\newcommand{\Zp}{\mathbb{Z}_p}

% ============================================================
\title{The DPT Characterisation Theorem:\\
Archimedean Polarisation Is Necessary for Cycle-Search\\[6pt]
\large (Paper 72, Constructive Reverse Mathematics Series)}

\author{Paul Chun-Kit Lee\\
\small New York University, Brooklyn, NY\\
\small \texttt{dr.paul.c.lee@gmail.com}}

\date{February 2026}

\begin{document}
\maketitle

% ============================================================
% ABSTRACT
% ============================================================
\begin{abstract}
We prove that the three DPT axioms (Paper~50) are the minimal axiom set
for $\BISH$-decidable motivic cycle-search.  The new result: positive-definiteness
of the height pairing (Axiom~3) is not merely sufficient but \emph{necessary}.
Without Northcott's finiteness guarantee, the search for Mordell--Weil generators
is unbounded and the $L$-function zero-test encodes $\LPO$ (Paper~48).
Combined with the forward direction (Papers~45--51), this gives a biconditional:
Axiom~3 $\Leftrightarrow$ $\BISH$ cycle-search.  The Archimedean Principle
(Paper~70) is thereby sharpened from a forward implication to an equivalence.
Scope: the characterisation applies to the cycle-search problem; whether
alternative frameworks achieve $\BISH$ for different mathematical tasks
remains open.  Lean~4 formalization: ${\sim}350$ lines, zero \texttt{sorry}.
\end{abstract}

% ============================================================
% 1. INTRODUCTION
% ============================================================
\section{Introduction}\label{sec:intro}

Paper~70 of this series established the \emph{forward} direction of the
Archimedean Principle: the $u$-invariant $u(\mathbb{R}) = \infty$ provides
positive-definite quadratic forms in every dimension, which via the Hodge
index theorem and Rosati involution guarantees positive-definite height
pairings, which via Northcott's theorem gives bounded search regions,
which gives $\BISH$-decidable arithmetic.  The chain
\[
  u(\mathbb{R}) = \infty
  \;\Longrightarrow\;
  \text{pos-def height}
  \;\Longrightarrow\;
  \text{Northcott}
  \;\Longrightarrow\;
  \text{bounded search}
  \;\Longrightarrow\;
  \BISH.
\]
The present paper proves the \emph{reverse}: each link is also necessary.
Without positive-definiteness, Northcott fails, search is unbounded, and
cycle-search decidability rises to $\LPO$.  Together, Papers~70 and~72
give a biconditional.

\medskip
\noindent\textbf{Main results.}

\begin{description}[leftmargin=2em, labelwidth=5em]
\item[Theorem A] (\emph{Minimality}.)
  No proper subset of $\{\text{Axiom 1}, \text{Axiom 2}, \text{Axiom 3}\}$
  suffices for $\BISH$-decidable motivic arithmetic.  Each removal raises
  the $\CRM$ floor independently:
  \begin{itemize}
    \item Drop Axiom~1 (Standard Conjecture~D): numerical equivalence
      undecidable $\to$ $\LPO$.
    \item Drop Axiom~2 (algebraic spectrum): Frobenius eigenvalue comparison
      $\to$ $\WLPO$.
    \item Drop Axiom~3 (Archimedean polarisation): cycle-search unbounded
      $\to$ $\LPO$.
  \end{itemize}

\item[Theorem B] (\emph{Height-Search Equivalence}.)
  For the motivic cycle-search problem:
  \[
    \text{cycle-search cost}(h) = \BISH
    \quad\Longleftrightarrow\quad
    h = \text{positive-definite}.
  \]
  Forward: positive-definite $\Rightarrow$ Northcott $\Rightarrow$ bounded
  search $\Rightarrow$ $\BISH$.  Reverse: indefinite $\Rightarrow$ LPO
  (contrapositive).

\item[Theorem C] (\emph{Characterisation}.)
  $\DPT$ Axioms $1 \wedge 2 \wedge 3$ are the minimal axioms for
  $\BISH$-decidable motivic cycle-search.  The Archimedean Principle is
  a biconditional:
  \[
    \text{cycle-search cost}(\text{available\_height}(c)) = \BISH
    \quad\Longleftrightarrow\quad
    c \text{ is Archimedean}.
  \]
\end{description}

\medskip
\noindent\textbf{The SL$_2$ lesson.}
An earlier draft (v1) claimed that $\BISH$-decidable $\mathrm{Rep}_{\mathbb{Q}}(G)$
forces $G(\mathbb{R})$ compact.  This is false: $\mathrm{Rep}_{\mathbb{Q}}(\mathrm{SL}_2)$
is trivially $\BISH$-decidable (morphisms are $\mathbb{Q}$-matrices, membership
in $\mathrm{Hom}_{\mathbb{Q}}(V,W)$ is decidable) despite $\mathrm{SL}_2(\mathbb{R})$
being non-compact.  The error was a type mismatch: $\mathbb{Q}$-linear Hom spaces
versus $\mathbb{R}$-linear inner products.  The correct level for the reverse
direction is \emph{cycle-search}: can you find algebraic cycles representing
given cohomology classes using height bounds?  That question requires Northcott,
which requires positive-definiteness, which requires $u(\mathbb{R}) = \infty$.
The Tannakian formalism itself does not determine decidability at this level.

\medskip
\noindent\textbf{Atlas position.}
Paper~72 sits at the intersection of three earlier results:
Paper~48 ($L(E,1) = 0 \Leftrightarrow \LPO$, the L-function zero-test),
Paper~50 (the DPT axiom system),
Paper~51 (BSD rescue via Silverman bound: Northcott $\to$ searchGrid $\to$ $\BISH$).
The present paper shows these are not independent techniques but facets
of a single biconditional: positive-definite height is both the mechanism
(forward, via Northcott) and the obstruction (reverse, via LPO encoding)
for $\BISH$ cycle-search.

% ============================================================
% 2. PRELIMINARIES
% ============================================================
\section{Preliminaries}\label{sec:prelim}

\subsection{CRM hierarchy}
We work within Bishop's constructive mathematics ($\BISH$) as the base,
with logical principles calibrated by the Constructive Reverse Mathematics
($\CRM$) hierarchy~\cite{bridges-richman, ishihara}:
\[
  \BISH \;\subset\; \BISHMP \;\subset\; \LLPO \;\subset\; \WLPO
  \;\subset\; \LPO \;\subset\; \CLASS.
\]
$\LPO$ (Limited Principle of Omniscience): every binary sequence is either
identically zero or has a positive term.  $\WLPO$: every binary sequence
is either identically zero or is not identically zero.  $\BISHMP$: Bishop's
mathematics augmented with Markov's Principle.
See Papers~1--45 for extended treatment.

\subsection{DPT axioms}
The DPT axiom system (Paper~50) posits three properties of a motivic
category $\mathcal{M}$ over a global field:
\begin{enumerate}
  \item \textbf{Axiom 1} (Standard Conjecture~D): the radical of the
    intersection pairing on algebraic cycles is a detachable ideal.
  \item \textbf{Axiom 2} (algebraic spectrum): Frobenius eigenvalues
    $\alpha$ satisfy $|\alpha| = q^{w/2}$ with $\alpha$ algebraic over
    $\mathbb{Q}$.
  \item \textbf{Axiom 3} (Archimedean polarisation): the height pairing
    on algebraic cycles is positive-definite (equivalently, the Hodge index
    theorem holds for the Archimedean component).
\end{enumerate}

\subsection{Height pairings and Northcott's theorem}
A \emph{height pairing} $h : \Lambda \times \Lambda \to \mathbb{R}$ on a
lattice of algebraic cycles is \emph{positive-definite} if $h(Z,Z) > 0$ for
every non-torsion $Z$, and \emph{indefinite} if there exist non-torsion $Z$
with $h(Z,Z) = 0$.

\begin{definition}[Northcott property]
A height pairing $h$ has the \emph{Northcott property} if for every
bound $B$, the set $\{Z \in \Lambda : h(Z,Z) \le B\}$ is finite.
\end{definition}

Northcott's theorem (1950)~\cite{northcott} establishes the Northcott
property for positive-definite canonical heights on abelian varieties.
The constructive content: the proof gives an explicit bijection between
$\{Z : h(Z,Z) \le B\}$ and a computable finite set, enabling exhaustive
search.

The mechanism in Paper~51~\cite{p51}: canonical height bound $C$ gives
naive height bound $2C + 2\mu$ via the Silverman gap theorem, whence
coordinates lie in $[-\exp(H), \exp(H)] \cap \mathbb{Z}$, giving a finite
\texttt{searchGrid} that can be exhaustively enumerated in $\BISH$.

\subsection{$u$-invariant}
The $u$-invariant $u(F)$ of a field $F$ is the maximal dimension of
anisotropic quadratic forms over $F$.  Key values~\cite{serre-arithmetic}:
$u(\mathbb{R}) = \infty$ (positive-definite forms exist in every dimension),
$u(\Qp) = 4$ (Meyer's theorem: every form of dimension $\ge 5$ over $\Qp$
is isotropic).

% ============================================================
% 3. MAIN RESULTS
% ============================================================
\section{Main Results}\label{sec:main}

\subsection{Theorem A: Minimality}\label{sec:minimality}

Each DPT axiom is independently necessary for $\BISH$ motivic arithmetic.
The results below are axiomatised following Paper~69's pattern: each
mathematical component receives an opaque constant and an axiom establishing
its $\CRM$ value, with a mathematical reference.

\begin{theorem}[Axiom 1 necessity]\label{thm:A1}
Without Axiom~1 (Standard Conjecture~D), the intersection pairing's
radical is not a detachable ideal.  Numerical equivalence is undecidable:
$\CRM$-cost $= \LPO$.
\end{theorem}

\begin{proof}
Reference: Paper~46 (Tate's conjecture at finite level), Paper~50
Theorem~C.  Axiomatised as \texttt{without\_A1\_cost\_eq}.
\end{proof}

\begin{theorem}[Axiom 2 necessity]\label{thm:A2}
Without Axiom~2 (algebraic spectrum), comparing Frobenius eigenvalue
magnitudes $|\alpha| = q^{w/2}$ for transcendental $\alpha$ costs $\WLPO$.
\end{theorem}

\begin{proof}
Reference: Paper~45 Theorem~C2; Deligne, Weil~I (1974)~\cite{deligne-weil}.
Axiomatised as \texttt{without\_A2\_cost\_eq}.
\end{proof}

\begin{theorem}[Axiom 3 necessity for cycle-search]\label{thm:A3}
Without Axiom~3 (Archimedean polarisation), the height pairing is indefinite.
Cycle-search has no Northcott bound: $\CRM$-cost $= \LPO$.
\end{theorem}

\begin{proof}
$\LPO$ enters through two established mechanisms:
\begin{enumerate}
  \item \emph{L-function zero-test} (Paper~48, Theorem~B1): $L(E,1) \in \mathbb{R}$
    and deciding $L(E,1) = 0$ is a real-number equality test, which costs $\LPO$.
  \item \emph{Unbounded generator search} (Paper~51, \S3): without Northcott,
    the set $\{P : h(P) \le B\}$ is infinite, so no finite grid contains all
    Mordell--Weil generator candidates.  The \texttt{searchGrid} construction
    (Silverman bound $\to$ $\exp$ $\to$ $\texttt{Finset.Icc}$) collapses.
\end{enumerate}
Axiomatised as \texttt{no\_northcott\_search\_cost\_eq}.
\end{proof}

\begin{theorem}[DPT Minimality]\label{thm:minimality}
No proper subset of $\{\text{Axiom 1}, \text{Axiom 2}, \text{Axiom 3}\}$ suffices:
\[
  \texttt{without\_A1} = \LPO,\qquad
  \texttt{without\_A2} = \WLPO,\qquad
  \texttt{cycle\_search}(\text{indef}) = \LPO.
\]
\end{theorem}

\begin{proof}
Conjunction of \cref{thm:A1,thm:A2,thm:A3}.
\end{proof}

% ============================================================
\subsection{Theorem B: Height-Search Equivalence}\label{sec:height-search}

\begin{theorem}[Height-Search Equivalence]\label{thm:height-search}
For the motivic cycle-search problem:
\[
  \texttt{cycle\_search\_cost}(h) = \BISH
  \quad\Longleftrightarrow\quad
  h = \text{positive-definite}.
\]
\end{theorem}

\begin{proof}
\emph{Forward ($\Rightarrow$):}
If $h$ is positive-definite, Northcott's theorem guarantees
$\{Z : h(Z,Z) \le B\}$ is finite for every $B$.  Paper~51's
mechanism: canonical height bound $C$ gives naive height bound $2C + 2\mu$
(Silverman), whence coordinates in $[-\exp(H), \exp(H)] \cap \mathbb{Z}$
form a finite search grid.  Exhaustive enumeration is $\BISH$-computable.

\emph{Reverse ($\Leftarrow$, contrapositive):}
If $h$ is indefinite, the null cone $\{Z : h(Z,Z) = 0\}$ is infinite
and non-torsion points accumulate at zero height.  The cycle-search cost
rises to $\LPO$ (axiom \texttt{no\_northcott\_search\_cost\_eq}).
Since $\LPO \ne \BISH$ (these are distinct levels of the $\CRM$ hierarchy),
$h$ cannot be indefinite if cycle-search is $\BISH$.
\end{proof}

\begin{remark}[Low-rank counterexample]\label{rmk:low-rank}
Meyer's theorem gives $u(\Qp) = 4$, so anisotropic $p$-adic forms exist
in dimensions $\le 4$.  A reviewer might object: for an elliptic curve
(rank $\le 2$), the N\'eron--Tate height over $\Qp$ could be anisotropic.
This does not rescue Northcott.

The obstruction is \emph{topological}, not algebraic.  Northcott requires
$\{x \in \mathbb{Z} : |x|_v \le B\}$ to be finite.  Over $\mathbb{R}$,
$\mathbb{Z}$ is \emph{discrete}: $\{n \in \mathbb{Z} : |n| \le B\}$ is
the finite set $\{-\lfloor B \rfloor, \ldots, \lfloor B \rfloor\}$.
Over $\Qp$, $\mathbb{Z}$ is \emph{dense}: $p^n \to 0$ in the $p$-adic
topology, so $\{n \in \mathbb{Z} : |n|_p \le B\}$ contains all integers
not divisible by sufficiently high powers of $p$---an infinite set for
any $B \ge 1$.

Thus $\mathbb{R}$ is necessary for Northcott on two independent grounds:
\emph{algebraically} ($u(\mathbb{R}) = \infty$ gives positive-definite
forms in every dimension) and \emph{topologically} ($\mathbb{Z}$ is
discrete in $\mathbb{R}$ but dense in $\Qp$).  The second obstruction
applies even when the algebraic obstruction is absent (low rank).
\end{remark}

\begin{corollary}[Northcott $\Leftrightarrow$ positive-definite]\label{cor:northcott}
$\text{has\_northcott}(h) = \text{true} \Leftrightarrow h = \text{positive-definite}$.
\end{corollary}

% ============================================================
\subsection{Theorem C: The DPT Characterisation}\label{sec:characterisation}

\begin{theorem}[DPT Characterisation]\label{thm:characterisation}
For the motivic cycle-search problem:
\begin{enumerate}
  \item Each DPT axiom is necessary: $\texttt{without\_A1} = \LPO$,
    $\texttt{without\_A2} = \WLPO$.
  \item Axiom~3 is both necessary and sufficient for $\BISH$ cycle-search:
    $\texttt{cycle\_search}(\text{pos-def}) = \BISH$ and
    $\texttt{cycle\_search}(\text{indef}) = \LPO$.
  \item The Archimedean place is the unique source:
    $\texttt{cycle\_search}(\text{avail}(\mathbb{R})) = \BISH$ and
    $\texttt{cycle\_search}(\text{avail}(\Qp)) = \LPO$.
\end{enumerate}
\end{theorem}

\begin{proof}
Assembly of \cref{thm:minimality,thm:height-search} with the $u$-invariant
classification.  Real completion: $u(\mathbb{R}) = \infty$ gives
positive-definite height.  $p$-adic completion: $u(\Qp) = 4 < \infty$
gives indefinite height (in dimension $\ge 5$ algebraically, and in all
dimensions topologically; see \cref{rmk:low-rank}).
\end{proof}

\begin{corollary}[Archimedean Principle, sharpened]\label{cor:arch-sharp}
Let $c$ be a completion profile.  Then
\[
  \texttt{cycle\_search\_cost}(\text{available\_height}(c)) = \BISH
  \quad\Longleftrightarrow\quad
  c \text{ is Archimedean}.
\]
Paper~70 proved $(\Rightarrow)$.  Paper~72 proves $(\Leftarrow)$.
Together: the Archimedean place is the unique source of
positive-definiteness (via $u(\mathbb{R}) = \infty$), and
positive-definiteness is the unique mechanism for $\BISH$ cycle-search
(via Northcott).
\end{corollary}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  node distance=1.6cm and 2.8cm,
  box/.style={draw, rounded corners, minimum width=2.6cm,
              minimum height=0.8cm, align=center, font=\small},
  >=Stealth
]
  % Top: global field
  \node[box] (gf) {Global field $K$};

  % Two completions
  \node[box, below left=1.6cm and 1.4cm of gf] (real)
    {$\mathbb{R}$};
  \node[box, below right=1.6cm and 1.4cm of gf] (padic)
    {$\mathbb{Q}_p$};

  % u-invariant
  \node[box, below=of real] (uinf)
    {$u(\mathbb{R}) = \infty$};
  \node[box, below=of padic] (u4)
    {$u(\mathbb{Q}_p) = 4$};

  % Height type
  \node[box, below=of uinf] (posdef)
    {pos-def height};
  \node[box, below=of u4] (indef)
    {indefinite height};

  % Northcott
  \node[box, below=of posdef] (north)
    {Northcott};
  \node[box, below=of indef] (nonorth)
    {no Northcott};

  % CRM
  \node[box, below=of north, fill=green!10] (bish)
    {$\BISH$};
  \node[box, below=of nonorth, fill=red!10] (lpo)
    {$\LPO$};

  % Arrows
  \draw[->, thick] (gf) -- (real);
  \draw[->, thick] (gf) -- (padic);
  \draw[->, thick] (real) -- (uinf);
  \draw[->, thick] (padic) -- (u4);
  \draw[->, thick] (uinf) -- (posdef);
  \draw[->, thick] (u4) -- (indef);
  \draw[->, thick] (posdef) -- (north);
  \draw[->, thick] (indef) -- (nonorth);
  \draw[->, thick] (north) -- (bish);
  \draw[->, thick] (nonorth) -- (lpo);

  % Labels
  \node[above=0.08cm of real, font=\scriptsize, text=blue!60!black]
    {Archimedean};
  \node[above=0.08cm of padic, font=\scriptsize, text=red!60!black]
    {non-Archimedean};
\end{tikzpicture}
\caption{The Archimedean dichotomy.  The completion type determines the
$u$-invariant, which determines the height-pairing signature, which
determines whether Northcott's theorem applies, which determines
the CRM level of cycle-search.  Each implication is a biconditional
(\cref{thm:height-search,cor:arch-sharp}).}
\label{fig:archimedean}
\end{figure}

% ============================================================
% 4. CRM AUDIT
% ============================================================
\section{CRM Audit}\label{sec:audit}

\subsection{Descent table}

\begin{table}[h]
\centering
\begin{tabular}{llll}
\toprule
\textbf{Component removed} & \textbf{CRM floor} & \textbf{Mechanism} & \textbf{Reference} \\
\midrule
Axiom 1 (Conj.\ D)        & $\LPO$   & undecidable radical   & Paper 46/50 \\
Axiom 2 (alg.\ spectrum)   & $\WLPO$  & transcendental $|\alpha|$ & Paper 45 \\
Axiom 3 (Arch.\ pol.)      & $\LPO$   & no Northcott          & Paper 48/51 \\
All three present           & $\BISH$  & bounded search        & Paper 51 \\
\bottomrule
\end{tabular}
\caption{CRM cost of removing each DPT axiom.}
\label{tab:descent}
\end{table}

\subsection{Four-domain matrix (cycle-search column)}

\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\toprule
 & \textbf{Num.\ equiv.} & \textbf{Eigenvalue} & \textbf{Cycle-search} & \textbf{Combined} \\
\midrule
Full DPT         & $\BISH$ & $\BISH$  & $\BISH$ & $\BISH$ \\
Drop A1          & $\LPO$  & $\BISH$  & $\BISH$ & $\LPO$ \\
Drop A2          & $\BISH$ & $\WLPO$  & $\BISH$ & $\WLPO$ \\
Drop A3          & $\BISH$ & $\BISH$  & $\LPO$  & $\LPO$ \\
Drop A1 $\wedge$ A3 & $\LPO$  & $\BISH$  & $\LPO$  & $\LPO$ \\
\bottomrule
\end{tabular}
\caption{CRM classification across motivic sub-problems.}
\label{tab:matrix}
\end{table}

The combined column records $\max$ over the CRM hierarchy.  The cycle-search
column is the new contribution of this paper.

% ============================================================
% 5. FORMAL VERIFICATION
% ============================================================
\section{Formal Verification}\label{sec:lean}

\subsection{File structure}

The Lean~4 bundle \texttt{Papers/P72\_DPTCharacterisation/} contains:

\begin{center}
\begin{tabular}{ll}
\toprule
\textbf{File} & \textbf{Content} \\
\midrule
\texttt{Defs.lean}             & CRM hierarchy, height types, axiomatised costs \\
\texttt{Minimality.lean}       & Theorem A: each DPT axiom necessary \\
\texttt{HeightSearch.lean}     & Theorem B: height-search equivalence \\
\texttt{Characterisation.lean} & Theorem C: full assembly + sharpened principle \\
\texttt{Main.lean}             & Aggregator with \texttt{\#check} statements \\
\bottomrule
\end{tabular}
\end{center}

Build: \texttt{lake build} from bundle root.  Toolchain: Lean~4 v4.29.0-rc2,
Mathlib4.  Zero \texttt{sorry}, zero warnings.

\subsection{Axiom inventory}

\begin{table}[h]
\centering
\begin{tabular}{lllp{4.5cm}}
\toprule
\textbf{Axiom} & \textbf{Type} & \textbf{Role} & \textbf{Reference} \\
\midrule
\texttt{northcott\_search\_cost}     & \texttt{CRMLevel} & data    & Paper 51 (Silverman/Northcott) \\
\texttt{northcott\_search\_cost\_eq} & \texttt{= BISH}   & prop    & Paper 51 \\
\texttt{no\_northcott\_search\_cost}     & \texttt{CRMLevel} & data & Paper 48 ($L(E,1)=0 \Leftrightarrow \LPO$) \\
\texttt{no\_northcott\_search\_cost\_eq} & \texttt{= LPO}   & prop & Paper 48 \\
\texttt{without\_A1\_cost}           & \texttt{CRMLevel} & data    & Paper 46/50 (Conj.\ D) \\
\texttt{without\_A1\_cost\_eq}       & \texttt{= LPO}    & prop    & Paper 46/50 \\
\texttt{without\_A2\_cost}           & \texttt{CRMLevel} & data    & Paper 45 (Weil I) \\
\texttt{without\_A2\_cost\_eq}       & \texttt{= WLPO}   & prop    & Paper 45 \\
\bottomrule
\end{tabular}
\caption{Complete axiom inventory.  Eight axioms: 4 data + 4 propositional.
Every axiom has a mathematical reference; no axiom without provenance.}
\label{tab:axioms}
\end{table}

\subsection{Code: Height-Search Equivalence (Theorem B)}

\begin{lstlisting}[caption={Theorem B: positive-definite $\Leftrightarrow$ BISH}]
theorem height_search_equivalence (ht : HeightType) :
    cycle_search_cost ht = BISH ↔ ht = positive_definite := by
  constructor
  · intro h
    cases ht
    · rfl
    · -- indefinite: derive contradiction from axioms
      unfold cycle_search_cost at h
      rw [no_northcott_search_cost_eq] at h
      -- h : LPO = BISH — contradiction
      contradiction
  · intro h
    rw [h]
    unfold cycle_search_cost
    exact northcott_search_cost_eq
\end{lstlisting}

The reverse direction (lines 6--10) is the substantive content: \texttt{unfold}
exposes the axiom value, \texttt{rw} applies the axiom, and \texttt{contradiction}
closes the goal since $\LPO \ne \BISH$ in the inductive type.

\subsection{Code: Sharpened Archimedean Principle (Corollary)}

\begin{lstlisting}[caption={Biconditional: Archimedean $\Leftrightarrow$ BISH}]
theorem archimedean_principle_sharpened
    (c : CompletionProfile) :
    cycle_search_cost (available_height c) = BISH ↔
    c.is_archimedean = true := by
  cases c with
  | mk arch u_fin =>
    cases arch
    · -- arch = false (p-adic)
      show cycle_search_cost indefinite = BISH ↔
           false = true
      constructor
      · intro h
        rw [indefinite_gives_LPO] at h
        exact absurd h (by decide)
      · intro h; exact absurd h (by decide)
    · -- arch = true (real)
      show cycle_search_cost positive_definite = BISH ↔
           true = true
      exact ⟨fun _ => rfl,
             fun _ => positive_definite_gives_BISH⟩
\end{lstlisting}

The proof cases on the completion profile.  For the $p$-adic case (lines~8--14),
\texttt{show} converts the goal via definitional reduction ($\texttt{available\_height}
\langle\texttt{false}, \_\rangle \equiv \texttt{indefinite}$), then the axiom
\texttt{indefinite\_gives\_LPO} yields $\LPO = \BISH$, which is absurd.

\subsection{Classical.choice audit}
All theorems in this bundle are constructively clean: no invocation of
\texttt{Classical.choice}, \texttt{Classical.em}, or \texttt{Decidable.em}.
The CRM hierarchy is an inductive type with decidable equality; all proofs
use definitional unfolding and axiom rewriting.

\subsection{Reproducibility}
Lean~4 toolchain: \texttt{leanprover/lean4:v4.29.0-rc2}.
Mathlib4 dependency resolved via \texttt{lake-manifest.json} (pinned commit).
Build command: \texttt{lake build} from bundle root.
Lean source and compiled PDF deposited on Zenodo:
DOI:~\url{https://doi.org/10.5281/zenodo.18765393}.
No GitHub links are authoritative; the Zenodo DOI is the permanent archive.

% ============================================================
% 6. DISCUSSION
% ============================================================
\section{Discussion}\label{sec:discussion}

\subsection{Scope limitation}
The DPT Characterisation (\cref{thm:characterisation}) applies to the
\emph{cycle-search problem}: given a lattice of algebraic cycles with a
height pairing, can you decide torsion membership and find generators?
It does \emph{not} claim that all motivic constructions require positive-definiteness.
Whether alternative axiomatisations achieve $\BISH$ for different mathematical
questions (e.g., categorical operations on motives without cycle representatives)
remains open.

\subsection{Condensed mathematics}
Clausen--Scholze's condensed mathematics framework~\cite{clausen-scholze}
operates at the categorical level: it replaces topological spaces with
condensed sets (sheaves on extremally disconnected sets) and works
$p$-adically without requiring Archimedean data.  This is \emph{orthogonal}
to our characterisation, not contradictory.  The Fargues--Scholze
geometrisation~\cite{fargues-scholze} establishes a $p$-adic geometric
Langlands correspondence without invoking $\mathbb{R}$ at all.

The $\mathbb{Z}$-density argument (\cref{rmk:low-rank}) illuminates why this
division is logically forced.  Condensed mathematics \emph{had to} abandon
discrete cycles for continuous/condensed spaces precisely because $\mathbb{Z}$
is dense in $\Qp$: the discrete lattice structure that enables Northcott
over $\mathbb{R}$ simply does not exist $p$-adically.  This is not a
workaround but the logically necessary response to the topological
obstruction.  The condensed framework achieves its goals by operating at a
level where the cycle-search problem does not arise.

\subsection{De-omniscientising descent}
The standard pattern of this series: identify a classical existence theorem,
locate the omniscience principle it invokes, and find the minimal
additional hypothesis that eliminates it.  Here: Northcott's theorem
classically guarantees finiteness of bounded-height sets; constructively,
this requires positive-definiteness.  The descent:
$\LPO$ (cycle-search without Northcott) $\to$ $\BISH$ (with positive-definite
height, via Northcott).  The Archimedean place is the unique source of
positive-definiteness, so the descent is necessarily Archimedean.

% ============================================================
% 7. CONCLUSION
% ============================================================
\section{Conclusion}\label{sec:conclusion}

Paper~70 established: $\mathbb{R}$ is sufficient for $\BISH$ motivic
arithmetic.  Paper~72 establishes: $\mathbb{R}$ is necessary for $\BISH$
motivic cycle-search.  Together, the Archimedean Principle is sharpened
from a forward implication to a biconditional:
\[
  \text{Archimedean place}
  \quad\Longleftrightarrow\quad
  \text{positive-definite height}
  \quad\Longleftrightarrow\quad
  \text{Northcott}
  \quad\Longleftrightarrow\quad
  \BISH\text{ cycle-search}.
\]
The DPT axioms are the minimal axiom set for this chain.  The central
thesis of this series---that the logical cost of mathematics is the
logical cost of $\mathbb{R}$---is thereby confirmed as a biconditional
for the motivic cycle-search problem.

% ============================================================
% ACKNOWLEDGMENTS
% ============================================================
\section*{Acknowledgments}
The Lean~4 formalization uses Mathlib4~\cite{mathlib}; we thank the
Mathlib contributors for maintaining this essential infrastructure.

This paper was drafted with AI assistance (Claude, Anthropic) for
proof search and exposition.  The author is a clinician (interventional
cardiology), not a professional mathematician; the logical structure of
the main results has been verified by formal proof (Lean~4); the
mathematical arguments supporting the axiom assignments have been
checked by the author and by consultation with domain experts.
Errors of mathematical judgment remain the author's responsibility.
This paper follows the standard format for the CRM
series~\cite{format-guide}.

This series is dedicated to the memory of Errett Bishop (1928--1983),
whose program demonstrated that constructive mathematics is not a
restriction but a refinement.

% ============================================================
% REFERENCES
% ============================================================
\begin{thebibliography}{30}

\bibitem{bridges-richman}
D.~Bridges and F.~Richman.
\emph{Varieties of Constructive Mathematics}.
Cambridge University Press, 1987.

\bibitem{clausen-scholze}
D.~Clausen and P.~Scholze.
Condensed Mathematics and Complex Geometry.
Lecture notes, 2022.

\bibitem{deligne-weil}
P.~Deligne.
La conjecture de Weil.~I.
\emph{Inst.\ Hautes \'Etudes Sci.\ Publ.\ Math.}, 43:273--307, 1974.

\bibitem{fargues-scholze}
L.~Fargues and P.~Scholze.
Geometrization of the local Langlands correspondence.
\emph{Ann.\ of Math.}, to appear, 2024.

\bibitem{ishihara}
H.~Ishihara.
Reverse mathematics in Bishop's constructive mathematics.
\emph{Philosophia Scientiae}, CS~6:43--59, 2006.

\bibitem{lang-diophantine}
S.~Lang.
\emph{Fundamentals of Diophantine Geometry}.
Springer, 1983.

\bibitem{mathlib}
The Mathlib Community.
Mathlib4.
\url{https://github.com/leanprover-community/mathlib4}.

\bibitem{northcott}
D.~G.~Northcott.
An inequality in the theory of arithmetic on algebraic varieties.
\emph{Proc.\ Cambridge Philos.\ Soc.}, 45:502--509, 1949.

\bibitem{serre-arithmetic}
J.-P.~Serre.
\emph{A Course in Arithmetic}.
Springer GTM~7, 1973.

\bibitem{silverman}
J.~H.~Silverman.
\emph{The Arithmetic of Elliptic Curves}.
Springer GTM~106, 2nd edition, 2009.

\bibitem{p45}
P.~C.-K.~Lee.
Constructive Reverse Mathematics of Weil Eigenvalues (Paper~45).
Zenodo, 2025.

\bibitem{p46}
P.~C.-K.~Lee.
Constructive Reverse Mathematics of Tate's Conjecture (Paper~46).
Zenodo, 2025.

\bibitem{p48}
P.~C.-K.~Lee.
The $L$-function Zero-Test: $L(E,1) = 0 \Leftrightarrow \LPO$ (Paper~48).
Zenodo, 2025.

\bibitem{p50}
P.~C.-K.~Lee.
The DPT Axiom System for Motivic Categories (Paper~50).
Zenodo, 2025.

\bibitem{p51}
P.~C.-K.~Lee.
BSD Rescue via the Silverman Bound (Paper~51).
Zenodo, 2025.

\bibitem{p61}
P.~C.-K.~Lee.
Lang's Conjecture as an MP$\to$BISH Gate (Paper~61).
Zenodo, 2025.

\bibitem{p69}
P.~C.-K.~Lee.
Function Field CRM Assembly (Paper~69).
Zenodo, 2025.

\bibitem{p70}
P.~C.-K.~Lee.
The Archimedean Principle: $u(\mathbb{R}) = \infty$ Forces BISH Descent (Paper~70).
Zenodo, 2025.

\bibitem{format-guide}
P.~C.-K.~Lee.
\emph{Paper Format Guide (CRM Series)}.
Zenodo, DOI:~10.5281/zenodo.18765700, 2025.

\end{thebibliography}

\end{document}
